### 2016-01-27 text/MS.Rmd
added original MS version from craig and dustin + suite of citation tools

```{r, results="hide", warnings=FALSE, echo=FALSE, message=FALSE}
source('../R/functions-analyses.R')
cleanbib()
myRefs   <-  listBibs()
exporBibs(myRefs, erase=FALSE)
```
**Temperature effects on mass-scaling exponent of metabolic rates: a manipulative test**

*draft date: `r as.character(Sys.Date())`*

Diego R. Barneche, Craig R. White, Dustin J. Marshall

Centre for Geometric Biology / School of Biological Sciences, Monash University, Clayton, Victoria, Australia 3800

**Abstract**

Organismal metabolic rate often scales allometrically with body mass.  This pattern is observed in both interspecific and ontogenetic (intraspecific) studies, but the causal basis of variation in the allometric scaling of metabolism remains unresolved.  In the present study, we use a manipulative approach to test for an association between ontogenetic metabolic scaling and the rate and form of growth in three phyla of colonial marine organisms. Based on a Dynamic Energy Budget model that predicts the scaling of metabolic rate on the basis of energy allocation to assimilation, maintenance, growth, and maturation, we hypothesise that the scaling exponent of metabolic rate for colonial organisms should be dependent on growth rate for two-dimensional (encrusting) colonies, but not for three-dimensional (arborescent) colonies.  In support of this hypothesis, we found that the scaling exponent of metabolic rate decreased with increasing growth rate for encrusting colonies, but not for arborescent ones. These findings highlight the importance of growth for metabolic scaling during ontogeny, suggest that differences in patterns of growth might explain at least some of the documented variation in the scaling exponent of metabolic rate, and lead to novel and testable predictions for associations between individual metabolism and community structure.

**Introduction**

Metabolic theory has been lauded as having the potential to unite hierarchical levels of biological organisation `r myCite(c('nisbet2000janimecol', 'brown2004', 'vandermeer2006tree'))`, but definitive tests of the mechanistic theories proposed to underlie the putative link between individual metabolic rate ($B_i$) and ecological patterns remain rare `r myCite('kearney2012amnat')`. Examination of the relationship between $B_i$ and individual body mass, $M_i$, provides an opportunity to test the mechanistic basis of metabolic theory, but manipulative experiments that examine this relationship are also rare `r myCite(c('kearney2012amnat', 'glazier2005biolrev', 'white2013jcompphysb'))`. The relationship between $B_i$ and $M_i$ is typically described by a power function of the form $B_i = B_oM_i^{\alpha}. For most animals, the value of $\alpha$ lies between 0.5 and 1, indicating that individual metabolic rate scales sub-linearly with body mass `r myCite(c('schmidtnielsen1984', 'savage2004', 'glazier2005biolrev'))`. Cross-sectional and longitudinal studies of the intra-specific association between $B_i$ and $M_i$ have revealed that the value of $\alpha$ is associated with many biotic and abiotic variables, including pH `r myCite('vaca2010biollett')`, temperature `r myCite('killen2010ele')`, predation `r myCite('glazier2011ecolmongr')`, inbreeding `r myCite('ketola2012jlinsoc')`, and a range of others (for reviews see `r myCite(c('glazier2005biolrev', 'white2013jcompphysb'))`). Similarly, inter-specific studies have shown that the value of $\alpha$ varies spatially and among taxa (e.g. `r myCite(c('lovegrove2000amnat', 'white2012procb'))`), with metabolic level (e.g. `r myCite('glazier2010biolrev')`), and sometimes with body mass (e.g. `r myCite(c('white2009evolution', 'kolokotrones2010nature'))`). While these intra- and inter-specific studies demonstrate clearly that the value of $\alpha$ is not fixed, they each share a common limitation: each study tests for an association between $\alpha$ and a biotic or abiotic variable, using a value of $\alpha$ that is estimated from the association between $B_i$ and $M_i$ for individuals or species of different size. Such correlative studies cannot address the causal association between $B_i$ and $M_i$, because $M_i$ is correlated with a range of other environmental, physiological, and life history variables (see e.g. `r myCite(c('peters1983', 'calder1984book', 'schmidtnielsen1984'))` for comprehensive reviews) and many of these variables are also associated with $B_i$ `r myCite(c('mcnab2002book', 'konarzewski2013jcompphysb', 'white2013jcompphysb'))`.

In contrast to studies of intact individual organisms, colonial animals provide an opportunity for definitive tests of metabolic theory because colony size can be manipulated experimentally and several metabolic theories predict $\alpha$ explicitly for these animals `r myCite(c('nakaya2005procb', 'white2011amnat', 'kearney2012amnat'))`. The fractal geometry model of West *et al.* `r myCite(c('west1997science', 'west1999science'))`, for example, predicts $\alpha$ on the basis of the transport of essential materials through space-filling outward-directed fractal distribution networks. For an $n$-dimensional organism, this model predicts that $\alpha$ should be equal to $n/(n ‑ 1)$, so a value of $\alpha$ = 0.75 is predicted for three-dimensional animals, and a value of $\alpha$ = 0.67 is predicted for two-dimensional ones `r myCite(c('west1997science', 'west1999science'))` see also `r myCite('koontz2009amjbot')`. However, subsequent work has demonstrated that these canonical predictions of the fractal model are dependent on a range of assumptions `r myCite(c('enquist2009nature', 'savage2008ploscompbiol', 'banavar2010pnas', 'kolokotrones2010nature'))`, and modifications of these can yield exponents that vary between 0.5–-1.0 (e.g. `r myCite('price2007pnas')`). For example, the predicted scaling exponent for a two-dimensional organism with a plane-filling network and area-increasing branching is 0.86 `r myCite('price2007pnas')`. Thus, the fractal model predicts an exponent of either 0.67 `r myCite(c('west1997science', 'west1999science'))` or 0.86 `r myCite('price2007pnas')` for two-dimensional animals, depending on the assumptions used. However, the fractal geometry model makes no predictions about the additional effects of biotic and abiotic variables on the value of $\alpha$, unless these variables have effects on the geometry of the distribution network.

The Dynamic Energy Budget (DEB) theory for metabolic organisation predicts patterns of growth and metabolic scaling on the basis of as generalised surface area (source) and volume (sink) relationships, and predicts the values of both $B_o$ and $\alpha$ during ontogeny on the basis of size-dependent variation in energy allocation to assimilation, maintenance, growth, and maturation `r myCite('kooijman2009book')`. Biotic and abiotic variables can therefore influence the predicted values of $\alpha$ through their effects on these processes. In the case of colonial organisms, DEB theory predicts that biotic and abiotic variables that influence growth rate will have different effects on $\alpha$ for two-dimensional encrusting and three-dimensional arborescent (tree-like) colonies, because patterns of growth differ among these forms. For encrusting colonies, the proportion of growing zooids varies as a function of colony circumference, and therefore scales as $M_i^{\sim 0.5}$. In contrast, the number of growing zooids for arborescent colonies scales as $M_i^{\sim 1}$, because a bifurcating tree with $n$ tips has $n-1$ nodes and a mass proportional to $2n-1$ (Figure 1). Given that the DEB predictions for metabolic scaling in colonial organisms depend upon growth rate and the relative contributions of growing and non-growing zooids to whole-colony metabolism, the utility of the DEB model in predicting variation in $\alpha$ can be tested in two ways. Firstly, because the relative contribution of growing and non-growing zooids to whole-colony metabolism varies with growth rate, which is influenced by temperature, a negative relationship between temperature and $\alpha$ is predicted for two-dimensional encrusting colonies `r myCite(c('white2011amnat', 'kearney2012amnat'))`. Since the normalization of the scaling relationship, $B_o$, will increase with temperature due to the thermodynamic effects of temperature on metabolic rate, this prediction will manifest as a negative association between $\alpha$ and $B_o$. Second, the influence of temperature on the scaling exponent of metabolic rate is predicted to differ among growth forms. This prediction arises because the number of growing tips of an arborescent colony scales as $M_i^{\sim 1}$ (Figure 1), so only the elevation of the relationship between whole-colony metabolic rate and colony mass is predicted to be influenced temperature and no association between between $\alpha$ and $B_o$ is predicted.

In the present study we experimentally test these two explicit predictions of the DEB model for metabolic scaling in colonial organisms by examining the effects of colony form (encrusting or arborescent) and temperature (as a proxy for growth rate) on the scaling of metabolic rate for colonial marine organisms from three phyla (Porifera, Bryozoa [Ectoprocta], and Chordata).

**Material and Methods**

To collect our study organisms, roughened acetate sheets secured to the underside of 6 mm thick PVC sheets suspended from floating pontoons at a depth of 1 m were deployed at Manly Boat Harbour, Queensland, Australia (27˚27' S, 153˚11' E) and checked regularly to remove fouling organisms other than our species of interest. After 6 weeks in the field, sheets bearing colonies were returned to the University of Queensland where they were maintained in aerated seawater for up to 48 h prior to measurements. For encrusting colonies, sections of acetate bearing whole colonies were then cut from the sheets and fragmented (see `r myCite('white2011amnat')` for details and tests of potential experimental artefacts). Arborescent colonies were cut directly from the acetate sheets and fragmented.

The rate of oxygen consumption, $\dot{V}\textrm{O}_\textrm{2}$ of colony fragments was measured at 10 or 25 ˚C using a 144-channel PreSens Sensor Dish Reader (SDR, AS-1 Scientific, Wellington, New Zealand) according to standard techniques `r myCite(c('koster2008meps', 'lighton2008book', 'white2011amnat'))`. $\dot{V}\textrm{O}_\textrm{2}$ was measured by placing the colony or fragment in a glass vial containing 0.2$\mu$m filtered sea water and a non-consumptive $\textrm{O}_\textrm{2}$ sensor spot and calculated from the rate of change of $\textrm{O}_\textrm{2}$ saturation ($m_a$, % h^‑1^) as:

$$\dot{V}\textrm{O}_\textrm{2} = ‑1 \left(\frac{m_a – m_b}{100}\right) V~{\alpha}_{\textrm{O}_\textrm{2}}$$, (1)

where $m_b$ is the rate of change of $\textrm{O}_\textrm{2}$ saturation for blank vials containing no bryozoans (% h^-1^), ${\alpha}_{\textrm{O}_\textrm{2}}$ is the oxygen capacitance of air-saturated sea water at 10 ˚C (6.26 ml l^-1^) or 24 ˚C (4.615 ml l^-1^) `r myCite('cameron1986book')` and $V$ is water volume (chambers were 0.005 l in volume and water volume was calculated by subtracting the volume of acetate and animals). At least four blank vials were recorded simultaneously with each 24-well plate to account for microbial oxygen consumption and sensor spots were calibrated with air-saturated seawater (100 % A.S.) and water containing 2% sodium sulfite (0% A.S.). Following measurement of $\dot{V}\textrm{O}_\textrm{2}$, colonies or fragments were blotted dry and weighed to 0.1 mg (Sartorius A 200 S, Sartorius AG, Göttingen, Germany).  All measurements of \dot{V}\textrm{O}_\textrm{2} were made in one of four dark constant temperature cabinets, with temperatures within a run randomised among cabinets.

To test for the effect of growth form and temperature on the relationship between mass and metabolic rate, we used ANCOVA where mass was the continuous variable, temperature and growth form were categorical variables. When we included experimental conveniences (controlled temperature cabinet and plate(cabinet)) as random factors, these factors and their interactions with our factors of interest were not significant. As they explained little additional variance and were not of biological interest, we omitted them from the final model `r myCite('quinn2002book')`. The scaling exponent, $\alpha$, of the relationship between $\dot{V}\textrm{O}_\textrm{2}$ and fragment mass, $M_i$, was then estimated by ordinary least squares (OLS) linear regression of $\textrm{ln}\dot{V}\textrm{O}_\textrm{2} ~ \textrm{ln}M_i$.

**Results**

Three-way interaction was significant; within forms there was an interaction between temperature and logM for encrusting colonies, but not arborescent ones.

**Discussion**

The Dynamic Energy Budget (DEB) theory predicts individual metabolism on the basis of energy allocation to assimilation, maintenance, growth, and maturation `r myCite('kooijman2009book')`. On the basis of this, DEB theory predicts a complex three-way interaction between growth form, temperature, and size `r myCite(c('white2011amnat', 'kearney2012amnat'))`. Thus, DEB not only predicts variation in the scaling exponent of metabolic rate, it does so in a quantitative way. In support of this prediction of DEB, we show that the effect of temperature on the scaling of metabolic rate in colonial organisms depends on the growth form of the colony. The scaling exponent of metabolic rate is temperature-dependent for encrusting colonies (Figure 2A), where the mass of growing units is proportional to colony $M_i^{\sim 0.5}$, but not for arborescent colonies, where the total mass of the growing units is proportional to colony $M_i^{\sim 1}$ (Figure 2B). Such patterns cannot be predicted by models in which the scaling exponent of metabolic rate is assumed to be invariant.

We found that both $\alpha$ and $B_o$ covary with each other and with temperature in encrusting species -- this covariance has also been noted across other species and has some interesting implications for life-history theory. $B_o$ represents the mass-independent metabolic rate of an organism of unit mass, and $\alpha$ represents the power function by which metabolism increases with mass. At higher temperatures, $B_o$ increases and $\alpha$ decreases -- in other words, the metabolic costs of existing, regardless of size, increase with temperature and the metabolic efficiency benefits of increased size increase with temperature. Thus, all else being equal, increases in temperature increase the metabolic benefits of attaining a larger size while decreases in temperature reduce mass-independent metabolic costs as well as reduce the metabolic benefits of attaining a larger size. We would therefore predict that species living at higher temperatures should allocate more resources to attaining larger sizes (as opposed to reproduction etc) while species living at lower temperatures will be under less pressure to allocate resources to size. Of course, there are many extrinsic factors that will affect selection on body size (e.g. predation, competition, etc) and a sizable theory discusses this issue but there are few formal considerations of the implications of temperature of $\alpha$ for life-history theory. Nevertheless, such differentials in the costs and benefits of attaining large size may help to explain the slow pace of life in species living in cold temperatures.

**A little example to consider.**

- Hot temperature, a = 0.6, B = 0.4. Cold temperature, a = 0.4, B = 0.6.  
- Energetic savings of being 2 units of size versus 10 units in size  
- Metabolic cost of being alive  for  2 unit size critter at high temperature is 0.79 or ~40% of total mass  
- Metabolic cost of being alive for 2 unit size critter at low temperature is 0.60 or ~30% of total mass  
- Metabolic cost of being alive for 10 unit size critter at high temperature is 1.5 or ~15% of total mass  
- Metabolic cost of being alive for 10 unit size critter at low temperature is 1.59 or ~16% of total mass  
- So 5 fold increase in size gets you a 2.667 increase in efficiency at high temperatures but same increase in size at low temperature gets you only a 1.85 increase in efficiency. That's a pretty big deal and we should say something about it!  
- Another example, at high temperatures, 10 colonies of 100 units in size will have a cumulative metabolism of 37 while 100 colonies of 10 units in size will have a cumulative metabolism of 150. (same overall mass). At low temperatures with same colony sizes and numbers, they are  63 and 159.  The ratio goes from 4.05 to 2.52. Massive change in metabolism if fragmentation occurs at high temperature, less change if fragmentation occurs at low temperatures.

New paras:  
- Many organisms exhibit multiple growth forms (encrusting and arborescent); disadvantage of arborescent is that there is no increase in efficiency with size.  
- Lack of evidence for WBE. Mass-scaling exponent not universal across the species analysed, and clear interactions with temperature.  
- Winners and losers across latitudinal gradients, and under increased temperatures. Arborsecents are predicted to be less common in warm waters,  
- Pocilliporids fast growing arborescent coral; massive corals are slow growing and something… This provides a mechanistic explanation for why arborescent corals grow fast, but attain smaller sizes than massive corals.  
- Rosa, R., and Seibel, B. A. 2009. Slow pace of life of the Antarctic colossal squid. J. Mar. Biol. Ass. U. K.

**Acknowledgements**

Kearney, people who did the plates, ARC for much money (DP0987626, DP110101776, and whatever grants Dustin wants to add).

**Figures legends**

*Figure 1.* Scaling of the number of the number of tips for a bifurcating arborescent colony with colony mass (expressed relative to the mass of a colony with one tip). Solid line represents isometric scaling of tip number with colony mass.

*Figure 2.* Scaling of rate of oxygen consumption ($\dot{V}\textrm{O}_\textrm{2}, $\mu$l h^-1^) with fragment mass ($M_i$, mg) for size-manipulated fragments of (A) encrusting and (B) arborescent colonial species. Data for *Hippopodina* sp, sponges, *Bugula stolonifera* and *B. neritina* are from the present study; data for *Botrylloides simodensis* and *Hippoporina indica* are from published sources `r myCite(c('nakaya2005procb', 'white2011amnat'))`. Slope and intercept are for the linear ordinary least squares regression relating $\textrm{ln}\dot{V}\textrm{O}_\textrm{2} ~ \textrm{ln}M_i \pm$ SE.

**Figures**

**References**

### 2016-01-27 R/functions-analyses.R

@@ -70,3 +70,44 @@ perSpeciesJagsRun  <-  function(dat) {
 	rownames(jagsOutput)[match(rowNames, paste0('beta[', seq_len(K), ']'))]  <-  colnames(modelMatrix)
 	list(simsMatrix=simsMatrix, jagsOutput=jagsOutput)
 }
+
+#############################
+# DEALING WITH BIB REFERENCES
+#############################
+getIndividualBibs  <-  function(path='~/bibtex_library/') {
+  listRefs     <-  dir(path)[grep('.bib', dir(path))]
+  listRefs[listRefs != 'library.bib']
+}
+
+readBibRefs  <-  function(individualBibs, path='~/bibtex_library/', verbose=TRUE) {
+  lapply(individualBibs, function(x, path, verbose) {
+    if(verbose) {
+      print(x)
+    }
+    read.bib(file.path(path, x))
+  }, path=path, verbose=verbose)
+}
+
+listBibs  <-  function() {
+  listRefs        <-  getIndividualBibs()
+  theRefs         <-  readBibRefs(listRefs)
+  names(theRefs)  <-  gsub('.bib', '', listRefs)
+  theRefs
+}
+
+exporBibs  <-  function(theRefs=listBibs(), libraryPath='../library.bib', erase=TRUE, verbose=TRUE) {
+  if(erase)
+    system(paste('rm -r', libraryPath))
+  if(!file.exists(libraryPath))
+    l_ply(theRefs, function(x, libraryPath, verbose) {
+      if(verbose) {
+        print(x)
+      }
+      write.bibtex(x, file=libraryPath, append=TRUE)
+    }, libraryPath=libraryPath, verbose=verbose)
+}
+
+myCite  <-  function(citationsVec) {
+  paste0('[@', paste0(citationsVec, collapse=';@'),']')
+}
+

### 2016-01-27 paper.R

@@ -1,0 +1,12 @@
+library(knitr)
+library(bibtex)
+library(knitcitations)
+library(plyr)
+
+rm(list=ls())
+###########
+# MAIN TEXT 
+###########
+knit('text/MS.Rmd', output=file.path(getwd(), 'text/MS.md'), quiet=TRUE, encoding = 'utf-8')
+# to word
+system('pandoc -o text/MS.docx text/MS.md -s -S --bibliography library.bib --csl functionalEcology.csl')

### 2016-01-28 R/functions-analyses.R
Plate is now a random effect

@@ -34,10 +34,15 @@ modelJags  <-  function(species) {
 	bug.file[length(bug.file)+1]  <-  	'\t}'
 	bug.file[length(bug.file)+1]  <-  	'\ttauB  ~   dgamma(1.0E-3,1.0E-3)'
 	bug.file[length(bug.file)+1]  <-  	'\tvarB  <-  1/tauB # variance, equivalent to inverse(tauB)'
+	bug.file[length(bug.file)+1]  <-  	'\tfor(i in 1:max(plate)) { '
+	bug.file[length(bug.file)+1]  <-  	'\t\tr[i] ~ dnorm(0, tauR)'
+	bug.file[length(bug.file)+1]  <-  	'\t}'
+	bug.file[length(bug.file)+1]  <-  	'\ttauR  ~   dgamma(1.0E-3,1.0E-3)'
+	bug.file[length(bug.file)+1]  <-  	'\tvarR  <-  1/tauR # variance, equivalent to inverse(tauR)'
 	bug.file[length(bug.file)+1]  <-  	'\n\t#linear regression for metabolic rate'
 	bug.file[length(bug.file)+1]  <-  	'\tfor(i in 1:length(lnRate)) {'
 	bug.file[length(bug.file)+1]  <-  	'\t\tlnRate[i]  ~  dnorm(muB[i], tauB)'
-	bug.file[length(bug.file)+1]  <-  	'\t\tmuB[i]     <- inprod(beta[], jagsModelMatrix[i,])'
+	bug.file[length(bug.file)+1]  <-  	'\t\tmuB[i]     <- r[plate[i]] + inprod(beta[], jagsModelMatrix[i,])'
 	bug.file[length(bug.file)+1]  <-  	'\t}'
 	bug.file[length(bug.file)+1]  <-  	'\n}'
 	write(bug.file, paste0('model_two_way_interaction_', species, '.bug'))
@@ -48,9 +53,9 @@ perSpeciesJagsRun  <-  function(dat) {
 	dat$Plate   <-  as.factor(dat$Plate)
 	if(enoughRuns) {
 		dat$Run      <-  as.factor(dat$Run)
-		modelMatrix  <-  model.matrix(~ Plate + Run + lnMass * Temp, data=dat, contrasts=list(Run='contr.sum', Plate='contr.sum'))
+		modelMatrix  <-  model.matrix(~ Run + lnMass * Temp, data=dat, contrasts=list(Run='contr.sum'))
 	} else {
[…]

### 2016-01-28 analyses.R
added plate as random effect, dropped boltzmann-type analyses, dropped random-effect selection (no longer of interest), and implemented formal DIC model selection between three-way and two-way interaction

@@ -8,52 +8,49 @@ source('paths.R')
 source('dataManipulation.R')
 source('R/functions-analyses.R')
 
-# fit model selection - is run a significant random effect?
-# http://stackoverflow.com/questions/24019807/how-to-compare-a-model-with-no-random-effects-to-a-model-with-a-random-effect-us
-modelLmer1  <-  lmer(lnRate ~ Species * lnMass * Temp + (1 | Run), data=metRates, REML=FALSE)
-modelLm1    <-  lm(lnRate ~ Species * lnMass * Temp, data=metRates)
-AICtab(modelLmer1, modelLm1) # Delta AIC > 2, run is a significant random effect
-
-# refit modelLmer1 using REML=TRUE
-modelLmer  <-  lmer(lnRate ~ Species * lnMass * Temp + (1 | Run), data=metRates, REML=TRUE)
-modelLmer2  <-  lmer(lnRate ~ lnMass * Temp + (1 | Run), data=metRates, REML=TRUE)
-modelLmer3  <-  lmer(lnRate ~ Temp + (1 | Run), data=metRates, REML=TRUE)
-anova(modelLmer, modelLmer2, modelLmer3) # all significant fixed effects
+# refit modelLmer using REML=TRUE and test for significance of fixed effects
+modelLmer1  <-  lmer(lnRate ~ Species * lnMass * Temp + (1 | Run) + (1 | Plate), data=metRates, REML=TRUE)
+modelLmer2  <-  lmer(lnRate ~ Species + lnMass * Temp + (1 | Run) + (1 | Plate), data=metRates, REML=TRUE)
+modelLmer3  <-  lmer(lnRate ~ lnMass * Temp + (1 | Run) + (1 | Plate), data=metRates, REML=TRUE)
+modelLmer4  <-  lmer(lnRate ~ Temp + (1 | Run) + (1 | Plate), data=metRates, REML=TRUE)
+anova(modelLmer1, modelLmer2, modelLmer3, modelLmer4) # all significant fixed effects
 
 # asses confidence intervals for model parameters
-confint.merMod(modelLmer)
+confint.merMod(modelLmer1)
 
-# refit modelLmer using Boltzmann function instead
-boltzLmer       <-  lmer(lnRate ~ Species + lnMass + invKT + (1 | Run), data=metRates)
-
-#############
-# FIT IN JAGS
-#############
-# fitting the Boltzmann function
-set.seed(1)
[…]

### 2016-01-28 R/functions-figures.R
added figure with parameter estimates and posterior density from three-way interaction model

@@ -50,11 +50,6 @@ to.pdf <- function(expr, filename, ...) {
   to.dev(expr, pdf, filename, ...)
 }
 
-linearRescale <- function(x, r.out) {
-  p <- (x - min(x)) / (max(x) - min(x))
-  r.out[[1]] + p * (r.out[[2]] - r.out[[1]])
-}
-
 rounded  <-  function(value, precision=1, change=FALSE) {
   if(change) {
     value  <-  value * -1
@@ -62,94 +57,152 @@ rounded  <-  function(value, precision=1, change=FALSE) {
   sprintf(paste0('%.', precision, 'f'), round(value, precision))
 }
 
-###########################
-# SPECIES AS RANFOM EFFECTS
-###########################
-posteriorsBoltzmann  <-  function() {
-  par(mfrow=c(1, 2), mar=c(5.1,4.1,4.1,3), omi=c(0.5,1,0.5,1), cex=1)
-  theCol  <-  c('Hippopodina sp.'='dodgerblue2', 'Bugula neritina'='darkolivegreen3', 'Sponge'='tomato', 'Bugula stolinifera'='darkgoldenrod')
-
-  # mass
-  plot(NA, xlim=c(-1.5,2.5), ylim=c(0,1.5), xlab='Mass scaling', ylab='Posterior density', las=1)
-  for(i in 1:4) {
-    xDens  <-  density(rowSums(bfit$BUGSoutput$sims.matrix[,c('A', paste0('s[', i, ',1]'))]))
-    polygon(c(xDens$x, min(xDens$x)), c(xDens$y, min(xDens$y)), border=theCol[i], col=make.transparent(theCol[i], 0.3))
-    label(c(0.02, 0.07), rep(1-i*0.1, 2), text=FALSE, type='l', lwd=2, col=theCol[i])
-      label(0.08, 1-i*0.1, round(xDens$x[which.max(xDens$y)[1]], 2), adj=c(0, 0.5), cex=0.8)
+###############
+# PAPER FIGURES
+###############
+figureProp  <-  function(x, side='x') {
+  if(side == 'x') {
+    n  <-  1:2
+  } else {
+    n  <-  3:4
   }
-
-  # temperature
-  par(mar=c(5.1,1.1,4.1,6))
[…]

### 2016-01-28 figures.R

@@ -1,4 +1,6 @@
 library(lme4)
+library(MASS)
+library(RColorBrewer)
 library(extrafont)
 library(fontcm)
 loadfonts()
@@ -9,6 +11,5 @@ source('R/functions-figures.R')
 source('dataManipulation.R')
 load('output/RDatafiles/analyses.RData')
 
-to.pdf(posteriorsBoltzmann(), 'output/figures/posteriorsBoltzmann.pdf', width=11, height=5.5)
-to.pdf(modelComparison(), 'output/figures/modelComparison.pdf', width=11, height=5)
-embed_fonts('output/figures/modelComparison.pdf')
+to.pdf(fig2(), 'output/figures/fig2.pdf', width=8, height=7)
+embed_fonts('output/figures/fig2.pdf')

### 2016-01-28 R/functions-figures.R
fixed to posterior probability + small additions to fig2

@@ -57,6 +57,17 @@ rounded  <-  function(value, precision=1, change=FALSE) {
   sprintf(paste0('%.', precision, 'f'), round(value, precision))
 }
 
+whiteGrid  <-  function() {
+  label(rep(0.2, 2), c(0,1), text=FALSE, type='l', col='white', lwd=0.5)
+  label(rep(0.4, 2), c(0,1), text=FALSE, type='l', col='white', lwd=0.5)
+  label(rep(0.6, 2), c(0,1), text=FALSE, type='l', col='white', lwd=0.5)
+  label(rep(0.8, 2), c(0,1), text=FALSE, type='l', col='white', lwd=0.5)
+  label(c(0,1), rep(0.2, 2), text=FALSE, type='l', col='white', lwd=0.5)
+  label(c(0,1), rep(0.4, 2), text=FALSE, type='l', col='white', lwd=0.5)
+  label(c(0,1), rep(0.6, 2), text=FALSE, type='l', col='white', lwd=0.5)
+  label(c(0,1), rep(0.8, 2), text=FALSE, type='l', col='white', lwd=0.5)
+}
+
 ###############
 # PAPER FIGURES
 ###############
@@ -68,6 +79,12 @@ figureProp  <-  function(x, side='x') {
   }
   usr  <-  par('usr')
   (x-usr[n[1]])/(usr[n[2]]-usr[n[1]])
+}
+
+getProb  <-  function(densObjt) {
+  densObjt$y  <-  (densObjt$x[2:length(densObjt$x)] - densObjt$x[1:(length(densObjt$x)-1)]) * rowMeans(cbind(densObjt$y[2:length(densObjt$y)], densObjt$y[1:(length(densObjt$y)-1)]))
+  densObjt$x  <-  rowMeans(cbind(densObjt$x[2:length(densObjt$x)], densObjt$x[1:(length(densObjt$x)-1)]))
+  densObjt
 }
 
 plotTriad  <-  function(species=species[1], spNames=spNames[1], tmat, y1=TRUE, x1=TRUE, y2=TRUE, x2=TRUE) {
@@ -92,10 +109,8 @@ plotTriad  <-  function(species=species[1], spNames=spNames[1], tmat, y1=TRUE, x
     i25  <-  rowSums(tmat[, c('(Intercept)', paste0('Species', species), paste0('Species', species, ':Temp25'))])
   }
  
-  xDens10    <-  density(s10)
[…]

### 2016-01-29 R/functions-figures.R
fixed interaction estimates, included regression plot in the mix

@@ -87,7 +87,7 @@ getProb  <-  function(densObjt) {
   densObjt
 }
 
-plotTriad  <-  function(species=species[1], spNames=spNames[1], tmat, y1=TRUE, x1=TRUE, y2=TRUE, x2=TRUE) {
+plotTriad  <-  function(species=species, spNames=spNames, tmat, y1=TRUE, x1=TRUE, y2=TRUE, x2=TRUE, y3=TRUE, x3=TRUE) {
   col10     <-  colorRampPalette(rev(brewer.pal(9, 'Blues')))
   col10     <-  col10(32)
   col10[1]  <-  '#FFFFFF'
@@ -98,15 +98,15 @@ plotTriad  <-  function(species=species[1], spNames=spNames[1], tmat, y1=TRUE, x
   # all species together - 3-way interaction
   isReferenceSpecies  <-  species == 'Bryo'
   if(isReferenceSpecies) {
-    s10  <-  tmat[, 'lnMass']
-    s25  <-  rowSums(tmat[, c('lnMass', 'lnMass:Temp25')])
     i10  <-  tmat[, '(Intercept)']
+    s10  <-  tmat[, 'lnMass']
     i25  <-  rowSums(tmat[, c('(Intercept)', 'Temp25')])
+    s25  <-  rowSums(tmat[, c('lnMass', 'lnMass:Temp25')])
   } else {
-    s10  <-  rowSums(tmat[, c('lnMass', paste0('Species', species, ':', 'lnMass'))])
-    s25  <-  rowSums(tmat[, c('lnMass', 'lnMass:Temp25', paste0('Species', species, ':', 'lnMass:Temp25'))])
     i10  <-  rowSums(tmat[, c('(Intercept)', paste0('Species', species))])
-    i25  <-  rowSums(tmat[, c('(Intercept)', paste0('Species', species), paste0('Species', species, ':Temp25'))])
+    s10  <-  rowSums(tmat[, c('lnMass', paste0('Species', species, ':', 'lnMass'))])
+    i25  <-  rowSums(tmat[, c('(Intercept)', paste0('Species', species), 'Temp25', paste0('Species', species, ':Temp25'))])
[…]

### 2016-01-29 figures.R

@@ -8,7 +8,6 @@ loadfonts()
 rm(list=ls())
 source('paths.R')
 source('R/functions-figures.R')
-source('dataManipulation.R')
 load('output/RDatafiles/analyses.RData')
 
 to.pdf(fig2(), 'output/figures/fig2.pdf', width=8, height=7)

### 2016-01-29 tables.R
added tables.R

@@ -1,0 +1,13 @@
+library(plyr)
+rm(list=ls())
+source('paths.R')
+source('dataManipulation.R')
+source('R/functions-figures.R')
+
+table1  <-  ddply(metRates, .(Species, Temp), function(x) {
+	data.frame(minMass = rounded(min(exp(x$lnMass))),
+			   maxMass = rounded(max(exp(x$lnMass))),
+			   n         = nrow(x), stringsAsFactors=FALSE)
+})
+
+write.csv(table1, 'output/data/table1.csv', row.names=FALSE)

### 2016-01-29 analyses.R
added original dataset to analyses output

@@ -66,5 +66,5 @@ for(i in 1:4) {
 }
 plot(comparedOutputs$lme4Coefs, comparedOutputs$jagsCoefs)
 
-rm(list=ls()[!(ls() %in% c('modelLmer1', 'tfit', 'tjagsModelMatrix', 'tjagsout', 'tfit2', 'tjagsModelMatrix2', 'tjagsout2', 'perSpeciesLmFits', 'perSpeciesJagsFits'))])
+rm(list=ls()[!(ls() %in% c('metRates', 'modelLmer1', 'tfit', 'tjagsModelMatrix', 'tjagsout', 'tfit2', 'tjagsModelMatrix2', 'tjagsout2', 'perSpeciesLmFits', 'perSpeciesJagsFits'))])
 save.image('output/RDatafiles/analyses.RData')

### 2016-01-30 analyses.R
removed species-by-species models, boltzmann model as well

@@ -1,14 +1,12 @@
-library(plyr)
 library(R2jags)
 library(lme4)
-library(bbmle)
 
 rm(list=ls())
 source('paths.R')
 source('dataManipulation.R')
 source('R/functions-analyses.R')
 
-# refit modelLmer using REML=TRUE and test for significance of fixed effects
+# test for significance of fixed effects
 modelLmer1  <-  lmer(lnRate ~ Species * lnMass * Temp + (1 | Run) + (1 | Plate), data=metRates, REML=TRUE)
 modelLmer2  <-  lmer(lnRate ~ Species + lnMass * Temp + (1 | Run) + (1 | Plate), data=metRates, REML=TRUE)
 modelLmer3  <-  lmer(lnRate ~ lnMass * Temp + (1 | Run) + (1 | Plate), data=metRates, REML=TRUE)
@@ -19,8 +17,8 @@ anova(modelLmer1, modelLmer2, modelLmer3, modelLmer4) # all significant fixed ef
 confint.merMod(modelLmer1)
 
 # fitting a three-way interaction function using run and plate as random effects on intercept
-# first create model matrix for three-way interaction - same order as modelLmer
-tjagsModelMatrix  <-  model.matrix(modelLmer)
+# first create model matrix for three-way interaction - same order as modelLmer1
+tjagsModelMatrix  <-  model.matrix(modelLmer1)
 K                 <-  ncol(tjagsModelMatrix)
 set.seed(1)
 tJags       <-  list('lnRate' = metRates$lnRate,
@@ -28,15 +26,15 @@ tJags       <-  list('lnRate' = metRates$lnRate,
 					 'K'     =  K,
 					 'run'   =  as.numeric(as.factor(metRates$Run)),
 					 'plate' =  as.numeric(as.factor(metRates$Plate)))
-tfit        <-  jags(data=tJags, parameters.to.save=c('varB', 'varR', 'varP', paste0('r[', sort(unique(as.numeric(as.factor(metRates$Run)))), ']'), paste0('p[', sort(unique(as.numeric(as.factor(metRates$Plate)))), ']'), paste0('beta[', seq_len(K), ']')), model.file='model_three_way_interaction.bug', n.chains=3, n.iter=5e5, DIC=TRUE, n.thin=250)
[…]

### 2016-01-30 R/functions-analyses.R
removed species-by-species model functions, added tempDelta for paper numbers

@@ -11,71 +11,6 @@ rmFcts  <-  function(filename) {
   rm(list=ls(.GlobalEnv)[ls(.GlobalEnv) %in% fct_files], envir=.GlobalEnv)
 }
 
-############
-# MODEL FITS
-############
-perSpeciesLmRun  <-  function(dat) {
-	enoughRuns  <-  length(unique(dat$Run)) > 1
-	dat$Plate   <-  as.factor(dat$Plate)
-	if(enoughRuns) {
-		dat$Run    <-  as.factor(dat$Run)		
-		lm(lnRate ~ Plate + Run + lnMass * Temp, data=dat, contrasts=list(Run='contr.sum', Plate='contr.sum'))
-	} else {
-		lm(lnRate ~ Plate + lnMass * Temp, data=dat, contrasts=list(Plate='contr.sum'))
-	}
-}
-
-modelJags  <-  function(species) {
-	bug.file                      <-  vector(mode='character')
-	bug.file[length(bug.file)+1]  <-  	'model {'
-	bug.file[length(bug.file)+1]  <-  	'\n\t# priors for fixed effects in metabolic rate analysis'
-	bug.file[length(bug.file)+1]  <-  	'\tfor(i in 1:K) { '
-	bug.file[length(bug.file)+1]  <-  	'\t\tbeta[i] ~ dnorm(0, 0.0001)'
-	bug.file[length(bug.file)+1]  <-  	'\t}'
-	bug.file[length(bug.file)+1]  <-  	'\ttauB  ~   dgamma(1.0E-3,1.0E-3)'
-	bug.file[length(bug.file)+1]  <-  	'\tvarB  <-  1/tauB # variance, equivalent to inverse(tauB)'
-	bug.file[length(bug.file)+1]  <-  	'\tfor(i in 1:max(plate)) { '
-	bug.file[length(bug.file)+1]  <-  	'\t\tr[i] ~ dnorm(0, tauR)'
-	bug.file[length(bug.file)+1]  <-  	'\t}'
-	bug.file[length(bug.file)+1]  <-  	'\ttauR  ~   dgamma(1.0E-3,1.0E-3)'
-	bug.file[length(bug.file)+1]  <-  	'\tvarR  <-  1/tauR # variance, equivalent to inverse(tauR)'
-	bug.file[length(bug.file)+1]  <-  	'\n\t#linear regression for metabolic rate'
-	bug.file[length(bug.file)+1]  <-  	'\tfor(i in 1:length(lnRate)) {'
[…]

### 2016-01-30 numbers.R
new numbers for paper

@@ -1,14 +1,17 @@
+library(plyr)
 library(lme4)
 
 rm(list=ls())
 source('paths.R')
+source('R/functions-analyses.R')
 load('output/RDatafiles/analyses.RData')
 
-# compare q10's between both models
-r10  <-  exp(fixef(modelLmer)['(Intercept)'] + fixef(modelLmer)['lnMass']*1)
-r25  <-  exp((fixef(modelLmer)['(Intercept)'] + fixef(modelLmer)['Temp25']) + fixef(modelLmer)['lnMass']*1)
-q10  <-  (r25 / r10) ^ (10/(25-10))
+# model selection result
+deltaDF   <-  anova(modelLmer1, modelLmer2)['Chi Df'][2,]
+deltaDIC  <-  tfit2$BUGSoutput$DIC - tfit$BUGSoutput$DIC
 
-r10b  <-  exp(fixef(boltzLmer)['(Intercept)'] + fixef(boltzLmer)['lnMass']*1 + (fixef(boltzLmer)['invKT'] / 8.62e-5 * (1 / 293.15 - 1 / 283.15)))
-r25b  <-  exp(fixef(boltzLmer)['(Intercept)'] + fixef(boltzLmer)['lnMass']*1 + (fixef(boltzLmer)['invKT'] / 8.62e-5 * (1 / 293.15 - 1 / 298.15)))
-q10b  <-  (r25b / r10b) ^ (10/(25-10))
+# compare deltas between temperatures for each species
+species     <-  unique(metRates$Species)
+tempDeltas  <-  ldply(species, getTempDeltas)
+
+save.image('output/RDatafiles/numbers.RData')

### 2016-01-30 R/functions-figures.R
renamed fig2 to fig1, changed some axes labels

@@ -125,7 +125,7 @@ plotTriad  <-  function(species=species, spNames=spNames, tmat, y1=TRUE, x1=TRUE
   }
 
   if(y2) {
-    y2  <-  substitute('Metabolic normalisation, '*italic('B'['o']))
+    y2  <-  substitute('ln '*italic('B'['o'])*' (units?)')
   } else {
     y2  <-  ''
   }
@@ -137,13 +137,13 @@ plotTriad  <-  function(species=species, spNames=spNames, tmat, y1=TRUE, x1=TRUE
   }
 
   if(y3) {
-    y3  <-  substitute('Metabolic rates, '*italic('B'['i']))
+    y3  <-  substitute('ln Rates, '*italic('B'['i'])*' (units?)')
   } else {
     y3  <-  ''
   }
 
   if(x3) {
-    x3  <-  'ln Mass (mg)'
+    x3  <-  substitute('ln Mass, '*italic('M'['i'])*' (mg)')
   } else {
     x3  <-  ''
   }
@@ -220,7 +220,7 @@ plotTriad  <-  function(species=species, spNames=spNames, tmat, y1=TRUE, x1=TRUE
   label(0.86, 0.06, rounded(mean(i25), 2), adj=c(1, 0.5), col='tomato', cex=0.8)
 }
 
-fig2  <-  function() {
+fig1  <-  function() {
   tmat  <-  tfit$BUGSoutput$sims.matrix
   tmat  <-  tmat[,paste0('beta[', seq_along(fixef(modelLmer1)), ']')]
   colnames(tmat)  <-  names(fixef(modelLmer1))

### 2016-01-30 figures.R

@@ -10,5 +10,5 @@ source('paths.R')
 source('R/functions-figures.R')
 load('output/RDatafiles/analyses.RData')
 
-to.pdf(fig2(), 'output/figures/fig2.pdf', width=8, height=7)
-embed_fonts('output/figures/fig2.pdf')
+to.pdf(fig1(), 'output/figures/fig1.pdf', width=8, height=7)
+embed_fonts('output/figures/fig1.pdf')

### 2016-01-30 text/MS.Rmd
new version of MS all the way up to half results, tables and figures added

@@
 ```{r, results="hide", warnings=FALSE, echo=FALSE, message=FALSE}
+load('../output/RDatafiles/analyses.RData')
+load('../output/RDatafiles/numbers.RData')
 source('../R/functions-analyses.R')
+source('../R/functions-figures.R')
 cleanbib()
 myRefs   <-  listBibs()
-exporBibs(myRefs, erase=FALSE)
+exporBibs(myRefs, erase=TRUE)
 ```
@@
 
-Metabolic theory has been lauded as having the potential to unite hierarchical levels of biological organisation `r myCite(c('nisbet2000janimecol', 'brown2004', 'vandermeer2006tree'))`, but definitive tests of the mechanistic theories proposed to underlie the putative link between individual metabolic rate ($B_i$) and ecological patterns remain rare `r myCite('kearney2012amnat')`.
-Examination of the relationship between $B_i$ and individual body mass, $M_i$, provides an opportunity to test the mechanistic basis of metabolic theory, but manipulative experiments that examine this relationship are also rare `r myCite(c('kearney2012amnat', 'glazier2005biolrev', 'white2013jcompphysb'))`.
-The relationship between $B_i$ and $M_i$ is typically described by a power function of the form $B_i = B_oM_i^{\alpha}.
-For most animals, the value of $\alpha$ lies between 0.5 and 1, indicating that individual metabolic rate scales sub-linearly with body mass `r myCite(c('schmidtnielsen1984', 'savage2004', 'glazier2005biolrev'))`.
-Cross-sectional and longitudinal studies of the intra-specific association between $B_i$ and $M_i$ have revealed that the value of $\alpha$ is associated with many biotic and abiotic variables, including pH `r myCite('vaca2010biollett')`, temperature `r myCite('killen2010ele')`, predation `r myCite('glazier2011ecolmongr')`, inbreeding `r myCite('ketola2012jlinsoc')`, and a range of others (for reviews see `r myCite(c('glazier2005biolrev', 'white2013jcompphysb'))`).
-Similarly, inter-specific studies have shown that the value of $\alpha$ varies spatially and among taxa (e.g.
-`r myCite(c('lovegrove2000amnat', 'white2012procb'))`), with metabolic level (e.g.
+Metabolic theories have been lauded as having the potential to unite hierarchical levels of biological organisation, from individuals to ecosystems `r myCite(c('nisbet2000janimecol', 'brown2004', 'vandermeer2006tree'))`.
+Although definitive tests of their underlying assumptions remain rare `r myCite(c('kearney2012amnat', 'newberry2015ploscomputbiol'))`, myriad studies examined their predictions with regards to the relationship between individual level metabolic rates, $B_i$, and individual body mass, $M_i$ `r myCite(c('kleiber1932', 'kleiber1961', 'peters1983', 'schmidtnielsen1984', 'glazier2005biolrev', 'delong2010', 'glazier2010biolrev'))`.
+The relationship between $B_i$ and $M_i$ is typically described by a power function of the form $B_i = B_oM_i^{\alpha}$, where $B_o$ is a normalisation constant independent of mass and $\alpha$ is a dimensionless scaling exponent.
+For most animals, individual metabolic rate scales sub-linearly with body mass, with $\alpha$ estimates falling between 0.5 and 1, and averaging around 0.75 `r myCite(c('kleiber1961', 'savage2004', 'glazier2005biolrev', 'delong2010'))`.
+Metabolic rates are also influenced by temperature.
+Two prevalent competing metabolic theories, the fractal geometry model `r myCite(c('west1997science', 'west1999science'))`, and  the Dynamic Energy Budget (DEB) theory for metabolic organisation `r myCite('kooijman2009book')`, provide opposing views on the role of temperature on metabolic rates and the relationship between the mass scaling of metabolic rates and temperature: the former assumes that temperature affects the metabolic normalisation only `r myCite('gillooly2001')`, i.e.
+mass-scaling and temperature dependence are independent; contrarily, the latter predicts changing relationships between mass scaling and temperature dependence through the effects of temperature on assimilation, maintenance, growth, and maturation `r myCite('kooijman2009book')`.
+
+Empirical studies of the intra-specific association between $B_i$ and $M_i$ have revealed that the value of $\alpha$ is associated with many biotic and abiotic variables, including pH `r myCite('vaca2010biollett')`, temperature `r myCite('killen2010ele')`, predation `r myCite('glazier2011ecolmongr')`, inbreeding `r myCite('ketola2012jlinsoc')`, and a range of others (for reviews see `r myCite(c('glazier2005biolrev', 'white2013jcompphysb'))`).
+Similarly, inter-specific studies have shown that the value of $\alpha$ varies among taxa and/or locations (e.g.
+`r myCite(c('lovegrove2000amnat', 'white2012procb', 'barneche2014ele'))`), with metabolic level (e.g.
 `r myCite('glazier2010biolrev')`), and sometimes with body mass (e.g.
@@
 
-In contrast to studies of intact individual organisms, colonial animals provide an opportunity for definitive tests of metabolic theory because colony size can be manipulated experimentally and several metabolic theories predict $\alpha$ explicitly for these animals `r myCite(c('nakaya2005procb', 'white2011amnat', 'kearney2012amnat'))`.
-The fractal geometry model of West *et al.* `r myCite(c('west1997science', 'west1999science'))`, for example, predicts $\alpha$ on the basis of the transport of essential materials through space-filling outward-directed fractal distribution networks.
-For an $n$-dimensional organism, this model predicts that $\alpha$ should be equal to $n/(n ‑ 1)$, so a value of $\alpha$ = 0.75 is predicted for three-dimensional animals, and a value of $\alpha$ = 0.67 is predicted for two-dimensional ones `r myCite(c('west1997science', 'west1999science'))` see also `r myCite('koontz2009amjbot')`.
-However, subsequent work has demonstrated that these canonical predictions of the fractal model are dependent on a range of assumptions `r myCite(c('enquist2009nature', 'savage2008ploscompbiol', 'banavar2010pnas', 'kolokotrones2010nature'))`, and modifications of these can yield exponents that vary between 0.5–-1.0 (e.g.
-`r myCite('price2007pnas')`).
-For example, the predicted scaling exponent for a two-dimensional organism with a plane-filling network and area-increasing branching is 0.86 `r myCite('price2007pnas')`.
-Thus, the fractal model predicts an exponent of either 0.67 `r myCite(c('west1997science', 'west1999science'))` or 0.86 `r myCite('price2007pnas')` for two-dimensional animals, depending on the assumptions used.
-However, the fractal geometry model makes no predictions about the additional effects of biotic and abiotic variables on the value of $\alpha$, unless these variables have effects on the geometry of the distribution network.
+In contrast to studies that make use of intact individual organisms, colonial animals provide an opportunity for definitive tests of metabolic theories.
+Colony size can be manipulated experimentally, hence allowing for the potential to remove the confounding effects of body size and life-history traits (e.g.
+`r myCite('white2011amnat')`).
+Furthermore, different colonial species may grow in different dimensions, and the mass-scaling exponent $\alpha$ is explicitly predicted to change with number of body dimensions `r myCite(c('nakaya2005procb', 'white2011amnat', 'kearney2012amnat'))`.
+For example, the fractal geometry model `r myCite(c('west1997science', 'west1999science'))` predicts that $\alpha = n/(n + 1)$ for a $n$-dimensional organism, hence a value of $\alpha$ = 0.75 is predicted for three-dimensional animals while a value of $\alpha$ = 0.67 is predicted for two-dimensional animals (but see `r myCite(c('enquist2009nature', 'savage2008ploscompbiol', 'banavar2010pnas', 'kolokotrones2010nature', 'price2007pnas', 'koontz2009amjbot'))`).
+On the other hand, DEB predicts that metabolic scaling in colonial organisms are dependent upon growth rate (which is in turn influenced by temperature) and the relative contributions of growing and non-growing zooids to whole-colony metabolism.
 
-The Dynamic Energy Budget (DEB) theory for metabolic organisation predicts patterns of growth and metabolic scaling on the basis of as generalised surface area (source) and volume (sink) relationships, and predicts the values of both $B_o$ and $\alpha$ during ontogeny on the basis of size-dependent variation in energy allocation to assimilation, maintenance, growth, and maturation `r myCite('kooijman2009book')`.
-Biotic and abiotic variables can therefore influence the predicted values of $\alpha$ through their effects on these processes.
-In the case of colonial organisms, DEB theory predicts that biotic and abiotic variables that influence growth rate will have different effects on $\alpha$ for two-dimensional encrusting and three-dimensional arborescent (tree-like) colonies, because patterns of growth differ among these forms.
-For encrusting colonies, the proportion of growing zooids varies as a function of colony circumference, and therefore scales as $M_i^{\sim 0.5}$.
-In contrast, the number of growing zooids for arborescent colonies scales as $M_i^{\sim 1}$, because a bifurcating tree with $n$ tips has $n-1$ nodes and a mass proportional to $2n-1$ (Figure 1).
-Given that the DEB predictions for metabolic scaling in colonial organisms depend upon growth rate and the relative contributions of growing and non-growing zooids to whole-colony metabolism, the utility of the DEB model in predicting variation in $\alpha$ can be tested in two ways.
-Firstly, because the relative contribution of growing and non-growing zooids to whole-colony metabolism varies with growth rate, which is influenced by temperature, a negative relationship between temperature and $\alpha$ is predicted for two-dimensional encrusting colonies `r myCite(c('white2011amnat', 'kearney2012amnat'))`.
-Since the normalization of the scaling relationship, $B_o$, will increase with temperature due to the thermodynamic effects of temperature on metabolic rate, this prediction will manifest as a negative association between $\alpha$ and $B_o$.
-Second, the influence of temperature on the scaling exponent of metabolic rate is predicted to differ among growth forms.
-This prediction arises because the number of growing tips of an arborescent colony scales as $M_i^{\sim 1}$ (Figure 1), so only the elevation of the relationship between whole-colony metabolic rate and colony mass is predicted to be influenced temperature and no association between between $\alpha$ and $B_o$ is predicted.
-
-In the present study we experimentally test these two explicit predictions of the DEB model for metabolic scaling in colonial organisms by examining the effects of colony form (encrusting or arborescent) and temperature (as a proxy for growth rate) on the scaling of metabolic rate for colonial marine organisms from three phyla (Porifera, Bryozoa [Ectoprocta], and Chordata).
+In the present study we experimentally manipulate size in different colonial marine organisms from two phyla (Bryozoa [Ectoprocta] and Porifera) in order to test whether the mass scaling and temperature dependence of metabolic rates interact `r myCite(c('white2011amnat', 'kearney2012amnat'))`.
+In doing so, we explicitly test if these interactions are general or species specific.
 
@@
 
-To collect our study organisms, roughened acetate sheets secured to the underside of 6 mm thick PVC sheets suspended from floating pontoons at a depth of 1 m were deployed at Manly Boat Harbour, Queensland, Australia (27˚27' S, 153˚11' E) and checked regularly to remove fouling organisms other than our species of interest.
+*1. Animal collections*  
+To collect our study organisms, roughened acetate sheets secured to the underside of 6 mm thick PVC sheets suspended from floating pontoons at a depth of 1 m were deployed at Manly Boat Harbour, Queensland, Australia (27˚27' S, 153˚11' E) and checked regularly to remove fouling organisms other than our species of interest: *Bugula neritina* and *B.
+stolonifera* (arborescent bryozoan), *Hippopodina* sp.
+(encrusting bryozoan), and an unidentified species of sponge belonging to the Microcionidae family `r myCite('hart2009ecology')` (Table 1).
 After 6 weeks in the field, sheets bearing colonies were returned to the University of Queensland where they were maintained in aerated seawater for up to 48 h prior to measurements.
@@
 
-The rate of oxygen consumption, $\dot{V}\textrm{O}_\textrm{2}$ of colony fragments was measured at 10 or 25 ˚C using a 144-channel PreSens Sensor Dish Reader (SDR, AS-1 Scientific, Wellington, New Zealand) according to standard techniques `r myCite(c('koster2008meps', 'lighton2008book', 'white2011amnat'))`.
+*2. Respiration measurements*  
+Individual metabolic rates (**provide units**, *n* = `r nrow(metRates)`), $B_i$, were measured as rate of oxygen consumption, $\dot{V}\textrm{O}_\textrm{2}$ of colony fragments was measured at 10 or 25 ˚C using a 144-channel PreSens Sensor Dish Reader (SDR, AS-1 Scientific, Wellington, New Zealand) according to standard techniques `r myCite(c('koster2008meps', 'lighton2008book', 'white2011amnat'))`.
 $\dot{V}\textrm{O}_\textrm{2}$ was measured by placing the colony or fragment in a glass vial containing 0.2$\mu$m filtered sea water and a non-consumptive $\textrm{O}_\textrm{2}$ sensor spot and calculated from the rate of change of $\textrm{O}_\textrm{2}$ saturation ($m_a$, % h^‑1^) as:
 
-$$\dot{V}\textrm{O}_\textrm{2} = ‑1 \left(\frac{m_a – m_b}{100}\right) V~{\alpha}_{\textrm{O}_\textrm{2}}$$, (1)
+$$\dot{V}\textrm{O}_\textrm{2} = ‑1 \left(\frac{m_a – m_b}{100}\right) V~{\beta}_{\textrm{O}_\textrm{2}}$$, (1)
 
-where $m_b$ is the rate of change of $\textrm{O}_\textrm{2}$ saturation for blank vials containing no bryozoans (% h^-1^), ${\alpha}_{\textrm{O}_\textrm{2}}$ is the oxygen capacitance of air-saturated sea water at 10 ˚C (6.26 ml l^-1^) or 24 ˚C (4.615 ml l^-1^) `r myCite('cameron1986book')` and $V$ is water volume (chambers were 0.005 l in volume and water volume was calculated by subtracting the volume of acetate and animals).
+where $m_b$ is the rate of change of $\textrm{O}_\textrm{2}$ saturation for blank vials containing no bryozoans (% h^-1^), ${\beta}_{\textrm{O}_\textrm{2}}$ is the oxygen capacitance of air-saturated sea water at 10 ˚C (6.26 ml l^-1^) or 24 ˚C (4.615 ml l^-1^) [**double-check these numbers.
+Shouldn't the second be 25 ˚C?**] `r myCite('cameron1986book')` and $V$ is water volume (chambers were 0.005 l in volume and water volume was calculated by subtracting the volume of acetate and animals).
 At least four blank vials were recorded simultaneously with each 24-well plate to account for microbial oxygen consumption and sensor spots were calibrated with air-saturated seawater (100 % A.S.) and water containing 2% sodium sulfite (0% A.S.).
 Following measurement of $\dot{V}\textrm{O}_\textrm{2}$, colonies or fragments were blotted dry and weighed to 0.1 mg (Sartorius A 200 S, Sartorius AG, Göttingen, Germany).
-All measurements of \dot{V}\textrm{O}_\textrm{2} were made in one of four dark constant temperature cabinets, with temperatures within a run randomised among cabinets.
+All measurements of $\dot{V}\textrm{O}_\textrm{2}$ were made in one of four dark constant temperature cabinets, with temperatures within a run randomised among cabinets.
 
-To test for the effect of growth form and temperature on the relationship between mass and metabolic rate, we used ANCOVA where mass was the continuous variable, temperature and growth form were categorical variables.
-When we included experimental conveniences (controlled temperature cabinet and plate(cabinet)) as random factors, these factors and their interactions with our factors of interest were not significant.
-As they explained little additional variance and were not of biological interest, we omitted them from the final model `r myCite('quinn2002book')`.
-The scaling exponent, $\alpha$, of the relationship between $\dot{V}\textrm{O}_\textrm{2}$ and fragment mass, $M_i$, was then estimated by ordinary least squares (OLS) linear regression of $\textrm{ln}\dot{V}\textrm{O}_\textrm{2} ~ \textrm{ln}M_i$.
+*3. Statistical analyses*  
+To test for the interacting effects of body mass and temperature on metabolic rate across species, we fitted a three-way interaction linear model using log-transformed data
+
+$$\textrm{ln}B_i \sim \textrm{ln}M_i * Temperature * Species$$, (2)
+
+where mass (mg), $\textrm{ln}M_i$, is a continuous variable, while temperature (two levels: 10 ˚C and 25 ˚C) and species (four levels corresponding to the four species) are categorical variables.
+With this model, we allow each species to have a different mass-scaling slope that may change with temperature (4 $\times$ 2 = 8 parameters); it also accounts for species- and temperature-specific deviations on the metabolic normalisation (i.e.
+statistical model intercept; 4 $\times$ 2 = 8 parameters).
+We also included experimental conveniences (plate reader, $\Delta P_r$, and date, $\Delta D$) as random effects on the metabolic normalisation to help control for natural experimental variability (3 additional hyper parameters corresponding to the standard deviation of $\Delta P_r$, $\Delta D$ and random-effects residual).
+
+We then performed a model selection by comparing the statistical model in equation 2 with a simpler two-way interaction model of similar structure,
+
+$$\textrm{ln}B_i \sim \textrm{ln}M_i * Temperature + Species$$. (3)
+
+In this model, different from the model in equation 2, the mass scaling slope is no longer allowed to vary across different species, but is still allowed to vary between different temperatures (1 $\times$ 2 = 2 parameters).
+The metabolic normalisation, however, is still allowed to vary across species (4 $\times$ 1 = 4 parameters) and between temperatures (1 additional parameters), and random effects were maintained (i.e.
+3 hyper parameters).
+Equation 3 represents a statistical model nested in the model represented by equation 2, and therefore their performance can be directly assessed `r myCite('zuur2009book')`.
+Such comparison allowed us to formally test our main hypothesis of significant mass-temperature interactions across different species.
+
+Both models were fitted in a Bayesian framework by calling *JAGS* version 3.4.0 from the R package *R2jags* version 0.05-03 `r myCite('r2jags')` in order to derive posterior distributions and associated 95% credible intervals (CIs) for the fitted parameters.
+Random effects were assumed to be normally distributed, with means of 0.
+Fitted parameters were assigned priors that were vague (i.e.
+locally uniform over the region supported by the likelihood) `r myCite('kruschke2014book')`.
+The posterior distributions of model parameters were estimated using Markov chain Monte Carlo (MCMC) methods by constructing three chains of 500,000 steps each.
+Chains were thinned using a 250-step interval, so a total of 6,000 steps were retained to estimate posterior distributions (i.e.
+3 $\times$ 500,000/250 = 6,000).
+Models were compared adopting Deviance information criterion (DIC), which are commonly used in Bayesian model selection and are analogously similar to the Akaike information criterion (AIC), as this method also penalises for additional model parameters `r myCite('hooten2015ecology')`.
+We adopted a conservative approach, and models were considered significantly different only if DIC units differed by > 10 units `r myCite('spiegelhalter2002jrss')`.
+All data and code necessary to reproduce this paper, its analyses, tables and figures can be obtained on GitHub https://github.com/dbarneche/MTRBrEs.
 
@@
 
-Three-way interaction was significant; within forms there was an interaction between temperature and logM for encrusting colonies, but not arborescent ones.
+Our model selection approach indicated that the three-way interaction model (equation 2) performed significantly better than the two-way interaction model (equation 3) ($\Delta$d.f.
+= `r deltaDF`; $\Delta$DIC = `r rounded(deltaDIC)`), thus implying that different species present different interactions between mass-scaling and temperature dependence.
+Particularly, the best model yields no interaction between mass scaling and temperature dependence for *Bugula stolonifera* and the Microcionidae sponge (large overlap between 95% credible  intervals).
+On the other hand, *Bugula neritina* and *Hippopodina* sp.
+present opposing interactions (Figure 1).
+For example, parameter estimates indicate that the difference in metabolic rates between 25 ˚C and 10 ˚C for *Hippopodina* sp.
+decreases `r rounded(tempDeltas[1,'tempDelta'])`-fold moving from `r round(exp(2))` to `r round(exp(5))` mg.
+On the other hand, rates increase `r rounded(1/tempDeltas[3,'tempDelta'])`-fold along the same body-mass gradient for *Bugula neritina*.
+
+**Compare model estimates with WBE and DEB predictions?**
 
@@
 
+**Tables**
+
+*Table 1.* Mass ranges and sampling efforts for the four species of Bryozoa and Porifera experimentally manipulated in the present study at two extreme temperatures.
+
 **Figures legends**
 
-*Figure 1.* Scaling of the number of the number of tips for a bifurcating arborescent colony with colony mass (expressed relative to the mass of a colony with one tip).
-Solid line represents isometric scaling of tip number with colony mass.
-
-*Figure 2.* Scaling of rate of oxygen consumption ($\dot{V}\textrm{O}_\textrm{2}, $\mu$l h^-1^) with fragment mass ($M_i$, mg) for size-manipulated fragments of (A) encrusting and (B) arborescent colonial species.
-Data for *Hippopodina* sp, sponges, *Bugula stolonifera* and *B.
-neritina* are from the present study; data for *Botrylloides simodensis* and *Hippoporina indica* are from published sources `r myCite(c('nakaya2005procb', 'white2011amnat'))`.
-Slope and intercept are for the linear ordinary least squares regression relating $\textrm{ln}\dot{V}\textrm{O}_\textrm{2} ~ \textrm{ln}M_i \pm$ SE.
+*Figure 1.* Estimates from the three-way interaction model as obtained by Bayesian methods in *JAGS*.
+Results for each species are represented in groups of four plots, with blue and red colours representing actual data and parameter estimates at 10 ˚C and 25 ˚C, respectively.
+*Top-left*: posterior distributions of mass-scaling exponents, with horizontal bars representing 95% credible intervals.
+*Bottom-right*: posterior distributions of metabolic normalisations, with vertical bars representing 95% credible intervals.
+*Bottom-left*: covariance between mass-scaling exponents and metabolic normalisations for each of the 6,000 MCMC iterations (translucent circles), with overlaid density heat-map.
+*Top-right*: regressions of individual metabolic rate, $\textrm{ln}B_i$, as functions of individual body mass, $\textrm{ln}M_i$.
+Regression fits were plotted using the average from the posterior distributions.
+In  probability posterior plots (top-left and bottom-right), smooth probability curves were approximated from the posterior density curves using very small bins (see electronic supplementary material).
 

### 2016-02-01 R/functions-analyses.R
added new numbers for results

@@ -54,24 +54,75 @@ myCite  <-  function(citationsVec) {
 ###############
 # PAPER NUMBERS
 ###############
-getTempDeltas   <-  function(species) {
-	tmat  <-  tfit$BUGSoutput$sims.matrix
-	tmat  <-  tmat[, paste0('beta[', seq_along(fixef(modelLmer1)), ']')]
-	colnames(tmat)  <-  names(fixef(modelLmer1))
-
-	isReferenceSpecies  <-  species == 'Bryo'
-	if(isReferenceSpecies) {
-		i10  <-  tmat[, '(Intercept)']
-		s10  <-  tmat[, 'lnMass']
-		i25  <-  rowSums(tmat[, c('(Intercept)', 'Temp25')])
-		s25  <-  rowSums(tmat[, c('lnMass', 'lnMass:Temp25')])
-	} else {
-		i10  <-  rowSums(tmat[, c('(Intercept)', paste0('Species', species))])
-		s10  <-  rowSums(tmat[, c('lnMass', paste0('Species', species, ':', 'lnMass'))])
-		i25  <-  rowSums(tmat[, c('(Intercept)', paste0('Species', species), 'Temp25', paste0('Species', species, ':Temp25'))])
-		s25  <-  rowSums(tmat[, c('lnMass', paste0('Species', species, ':', 'lnMass'), 'lnMass:Temp25', paste0('Species', species, ':', 'lnMass:Temp25'))])
-	}
-	
-	# deltas between temperatures
-	data.frame(species=species, tempDelta=(exp(mean(i25) + mean(s25)*2) / exp(mean(i10) + mean(s10)*2)) / (exp(mean(i25) + mean(s25)*5) / exp(mean(i10) + mean(s10)*5)), stringsAsFactors=FALSE)
+cleanJagsSummary  <-  function(jagsSummary=tfit$BUGSoutput$summary, modelMatrix=modelLmer1) {
+  jagsMat            <-  jagsSummary[paste0('beta[', seq_along(fixef(modelMatrix)), ']'), ]
+  rownames(jagsMat)  <-  names(fixef(modelMatrix))
+  jagsMat
 }
+
+cleanJagsMatrix  <-  function(jagsMatrix=tfit$BUGSoutput$sims.matrix, modelMatrix=modelLmer1) {
+  jagsMat            <-  jagsMatrix[, paste0('beta[', seq_along(fixef(modelMatrix)), ']')]
[…]

### 2016-02-01 numbers.R

@@ -10,8 +10,16 @@ load('output/RDatafiles/analyses.RData')
 deltaDF   <-  anova(modelLmer1, modelLmer2)['Chi Df'][2,]
 deltaDIC  <-  tfit2$BUGSoutput$DIC - tfit$BUGSoutput$DIC
 
-# compare deltas between temperatures for each species
-species     <-  unique(metRates$Species)
-tempDeltas  <-  ldply(species, getTempDeltas)
+# get average estimates and 95% CI for each parameter
+species           <-  unique(metRates$Species)
+averageEstimates  <-  ldply(species, getAverageEstimates)
+ci95Estimates     <-  ldply(species, get95CIEstimates)
 
+# compare deltas between temperatures for each species
+tempDeltas  <-  ddply(averageEstimates, .(species), getTempDeltas)
+
+# q10-equivalents [ q10 == (r25 / r10) ^ (10/(25-10)) ] for Microcionidae and Bugula stolonifera
+q10andErs  <-  ddply(averageEstimates[averageEstimates$species %in% c('Sponge', 'Stolonifera'), ], .(species), getq10andEr)
+
+rm(list=ls()[!(ls() %in% c('deltaDF', 'deltaDIC', 'species', 'averageEstimates', 'ci95Estimates', 'tempDeltas', 'q10andErs'))])
 save.image('output/RDatafiles/numbers.RData')

### 2016-02-01 analyses.R
new potential models to fit in jags for model selection

@@ -6,12 +6,16 @@ source('paths.R')
 source('dataManipulation.R')
 source('R/functions-analyses.R')
 
-# test for significance of fixed effects
+# quickly test for significance of fixed effects using lme4
 modelLmer1  <-  lmer(lnRate ~ Species * lnMass * Temp + (1 | Run) + (1 | Plate), data=metRates, REML=TRUE)
 modelLmer2  <-  lmer(lnRate ~ Species + lnMass * Temp + (1 | Run) + (1 | Plate), data=metRates, REML=TRUE)
 modelLmer3  <-  lmer(lnRate ~ lnMass * Temp + (1 | Run) + (1 | Plate), data=metRates, REML=TRUE)
 modelLmer4  <-  lmer(lnRate ~ Temp + (1 | Run) + (1 | Plate), data=metRates, REML=TRUE)
+modelLmer5  <-  lmer(lnRate ~ Species * Temp + lnMass + (1 | Run) + (1 | Plate), data=metRates, REML=TRUE)
+modelLmer6  <-  lmer(lnRate ~ Species * lnMass + Temp + (1 | Run) + (1 | Plate), data=metRates, REML=TRUE)
 anova(modelLmer1, modelLmer2, modelLmer3, modelLmer4) # all significant fixed effects
+anova(modelLmer1, modelLmer5) # three-way significantly better
+anova(modelLmer1, modelLmer6) # three-way significantly better
 
 # asses confidence intervals for model parameters
 confint.merMod(modelLmer1)

### 2016-02-01 text/MS.Rmd
version 2 of manuscript, with full summary, results, and half discussion

@@
 myRefs   <-  listBibs()
-exporBibs(myRefs, erase=TRUE)
+exporBibs(myRefs, erase=FALSE)
 ```
@@
 
-**Abstract**
+**Summary**
 
-Organismal metabolic rate often scales allometrically with body mass.
-This pattern is observed in both interspecific and ontogenetic (intraspecific) studies, but the causal basis of variation in the allometric scaling of metabolism remains unresolved.
-In the present study, we use a manipulative approach to test for an association between ontogenetic metabolic scaling and the rate and form of growth in three phyla of colonial marine organisms.
-Based on a Dynamic Energy Budget model that predicts the scaling of metabolic rate on the basis of energy allocation to assimilation, maintenance, growth, and maturation, we hypothesise that the scaling exponent of metabolic rate for colonial organisms should be dependent on growth rate for two-dimensional (encrusting) colonies, but not for three-dimensional (arborescent) colonies.
-In support of this hypothesis, we found that the scaling exponent of metabolic rate decreased with increasing growth rate for encrusting colonies, but not for arborescent ones.
-These findings highlight the importance of growth for metabolic scaling during ontogeny, suggest that differences in patterns of growth might explain at least some of the documented variation in the scaling exponent of metabolic rate, and lead to novel and testable predictions for associations between individual metabolism and community structure.
+1.
+Organismal metabolic rate often scales allometrically with body mass, both inter and intra-specifically.
+Metabolic rates are also known to increase with temperature, but most theoretical tests of mass scaling and temperature dependence have treated both phenomena independently.
+2.
+In the present study, we use a manipulative approach to test for an association between metabolic mass scaling and temperature dependence in four species belonging to two phyla of colonial marine invertebrates.
+3.
+We show that mass-scaling exponents and temperature dependencies interact in different ways for different species.
+In so doing, we discuss the potentials of our results to multi-level scaling theory, showing that some species might be winners while others might be losers under different temperature scenarios.
+4.
+These findings highlight the need for new metabolic theory that formally incorporate interactions between mass-scaling and temperature dependence in ectothermic organisms, in order to better explain documented variation in the mass-scaling exponent of metabolic rate, and lead to novel and testable predictions for associations between individual metabolism and community structure.
+
+**Key-words:** energetics, scaling, thermal limits, latitude, metabolic theory
 
@@
 *3. Statistical analyses*  
+**Things to possibly improve -- since we use a Bayesian approach, why let parameters be estimated from data (i.e.
+vague priors)?
+Why not adopt strong priors?
+Maybe letting it estimate is better to check whether parameters follow one or another theory?**
+**Things to possibly improve -- Include more models for model selection?
+E.g.
+two-way models with species*temp interaction or species*mass?
+lme4 already indicates that they are not better, I'm pretty sure the Bayesian DIC will follow results along the same lines.
+We'd be covering our bases better beforehand before a reviewer asks for this**
+
 To test for the interacting effects of body mass and temperature on metabolic rate across species, we fitted a three-way interaction linear model using log-transformed data
@@
 = `r deltaDF`; $\Delta$DIC = `r rounded(deltaDIC)`), thus implying that different species present different interactions between mass-scaling and temperature dependence.
-Particularly, the best model yields no interaction between mass scaling and temperature dependence for *Bugula stolonifera* and the Microcionidae sponge (large overlap between 95% credible  intervals).
-On the other hand, *Bugula neritina* and *Hippopodina* sp.
-present opposing interactions (Figure 1).
-For example, parameter estimates indicate that the difference in metabolic rates between 25 ˚C and 10 ˚C for *Hippopodina* sp.
-decreases `r rounded(tempDeltas[1,'tempDelta'])`-fold moving from `r round(exp(2))` to `r round(exp(5))` mg.
-On the other hand, rates increase `r rounded(1/tempDeltas[3,'tempDelta'])`-fold along the same body-mass gradient for *Bugula neritina*.
+Particularly, the best model yields no interaction between mass scaling and temperature dependence for *Bugula stolonifera* and the Microcionidae sponge (large overlap between 95% credible intervals for the mass-scaling slopes, $\alpha$) (Figure 1).
+Still, while the mass-scaling slopes are statistically the same between temperatures for these species, there is evidence for higher metabolic normalisations (i.e.
+intercepts), $B_o$, at higher temperatures (though not statistically significant for Microcionidae).
+For example, metabolic rates increase `r rounded(exp(averageEstimates$lnB25[averageEstimates$species == 'Stolonifera'])/exp(averageEstimates$lnB10[averageEstimates$species == 'Stolonifera']))`-fold between 10 ˚C and 25 ˚C for *B.
+stolonifera*, yielding a $Q_{10}$ of `r rounded(q10andErs$q10[q10andErs$species == 'Stolonifera'])`.
+It is worth noting that this $Q_{10}$ value is equivalent to an activation energy of `r rounded(q10andErs$er[q10andErs$species == 'Stolonifera'], 2)` eV (i.e.
+$E_r = \frac{k \times 283.15 \times 298.15 \times 1.5 \times \textrm{ln}Q_{10}}{(298.15 - 283.15)}$, where $E_r$ is an activation energy and $k$ is the Boltzmann constant, 8.62 $\times$ 10^-5^ eV K^-1^), which is consistent with the average 06 -- 0.7 eV reported in the Metabolic Theory of Ecology literature `r myCite('yvondurocher2012a')`.
+Different from the above species, the best model indicates that *Bugula neritina* and *Hippopodina* sp.
+present opposing interactions between mass-scaling and temperature dependence (Figure 1).
+For example, average posterior parameter estimates indicate that the difference in metabolic rates between 25 ˚C and 10 ˚C for *Hippopodina* sp.
+decreases, on average, `r rounded(tempDeltas[1,'tempDelta'])`-fold moving from `r round(exp(2))` to `r round(exp(5))` mg.
+For *Bugula neritina*, however, rates increase `r rounded(1/tempDeltas[3,'tempDelta'])`-fold along the same body-mass gradient.
 
-**Compare model estimates with WBE and DEB predictions?**
+Posterior distributions for mass-scaling exponents show varied consistency with predictions from the two metabolic theories.
+For example, *Hippopodina* sp.
+shows exponents that encompass both 0.67 (**DEB prediction??**; Fractal geometry model prediction for 2-D organisms; `r myCite('koontz2009amjbot')`) and 0.75 (Fractal geometry model prediction for 3-D organisms `r myCite(c('west1997science', 'west1999science'))`) at 10 ˚C (95% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's10_97.5'], 2)`), but that are significantly shallower than 0.75 at 25 ˚C (95% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's25_97.5'], 2)`).
+On the other hand, *Bugula neritina* presents scaling exponents that overlap both 0.67 and 0.75 at 25 ˚C (95% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's25_97.5'], 2)`), but are significantly shallower than these values at 10 ˚C (95% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's10_97.5'], 2)`).
+Microcionidae presents exponents that only overlap 0.67 at 10 ˚C, but not at 25 ˚C (95% CI at 10 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's10_97.5'], 2)`; 95% CI at 25 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's25_97.5'], 2)`).
+Finally, *Bugula stolonifera* presents values that overlap both 0.67 and 0.75 at both temperatures (95% CI at 10 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's10_97.5'], 2)`; 95% CI at 25 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's25_97.5'], 2)`).
 
@@
 
-The Dynamic Energy Budget (DEB) theory predicts individual metabolism on the basis of energy allocation to assimilation, maintenance, growth, and maturation `r myCite('kooijman2009book')`.
-On the basis of this, DEB theory predicts a complex three-way interaction between growth form, temperature, and size `r myCite(c('white2011amnat', 'kearney2012amnat'))`.
-Thus, DEB not only predicts variation in the scaling exponent of metabolic rate, it does so in a quantitative way.
-In support of this prediction of DEB, we show that the effect of temperature on the scaling of metabolic rate in colonial organisms depends on the growth form of the colony.
-The scaling exponent of metabolic rate is temperature-dependent for encrusting colonies (Figure 2A), where the mass of growing units is proportional to colony $M_i^{\sim 0.5}$, but not for arborescent colonies, where the total mass of the growing units is proportional to colony $M_i^{\sim 1}$ (Figure 2B).
-Such patterns cannot be predicted by models in which the scaling exponent of metabolic rate is assumed to be invariant.
+Here, we show experimentally that the mass-scaling of metabolic rates changes between two extreme temperatures, and in different ways across different species of marine invertebrates.
+We build on recent empirical evidence documenting similar changes (e.g.
+`r myCite('killen2010ele')`), yielding further support to the need for new metabolic theories (or reformulation of current ones) that formally incorporate interactive effects between mass scaling and temperature dependence of biological rates.
+*Currently, the Dynamic Energy Budget (DEB) theory predicts individual metabolism on the basis of energy allocation to assimilation, maintenance, growth, and maturation `r myCite('kooijman2009book')`, yielding a complex three-way interaction between growth form, temperature, and size `r myCite(c('white2011amnat', 'kearney2012amnat'))`.* **If we want to keep this comment on DEB, we will have to formally present how and what predictions DEB makes for changes in mass-scaling between different temperatures**.
+Our results, however, do not indicate consistent mass-temperature interactions across growth forms.
+Possibly, another important axis involved, which has not been considered here, is the acclimation capacity of species upon acute temperature changes (e.g.
+`r myCite('rezende2014functecol')`).
+Our findings are constrained to acute-related changes, and therefore do not consider temporal acclimation or even adaptation to changes in environmental temperature.
 
-We found that both $\alpha$ and $B_o$ covary with each other and with temperature in encrusting species -- this covariance has also been noted across other species and has some interesting implications for life-history theory.
-$B_o$ represents the mass-independent metabolic rate of an organism of unit mass, and $\alpha$ represents the power function by which metabolism increases with mass.
-At higher temperatures, $B_o$ increases and $\alpha$ decreases -- in other words, the metabolic costs of existing, regardless of size, increase with temperature and the metabolic efficiency benefits of increased size increase with temperature.
-Thus, all else being equal, increases in temperature increase the metabolic benefits of attaining a larger size while decreases in temperature reduce mass-independent metabolic costs as well as reduce the metabolic benefits of attaining a larger size.
-We would therefore predict that species living at higher temperatures should allocate more resources to attaining larger sizes (as opposed to reproduction etc) while species living at lower temperatures will be under less pressure to allocate resources to size.
-Of course, there are many extrinsic factors that will affect selection on body size (e.g.
-predation, competition, etc) and a sizable theory discusses this issue but there are few formal considerations of the implications of temperature of $\alpha$ for life-history theory.
-Nevertheless, such differentials in the costs and benefits of attaining large size may help to explain the slow pace of life in species living in cold temperatures.
+The fact that $B_o$ and $\alpha$ covary in different ways depending on species and temperature (Figure 1) suggests that the metabolic costs of maintenance will vary for different species at different environmental temperatures.
+For example, *Bugula neritina* presented a shallower mass-scaling exponent at colder temperatures, possibly indicating that large cold-environment individuals are more efficient in spending energy than counterparts in warm environments.
+Thus, for this species, all else being equal, increases in temperature decrease the metabolic benefits of attaining a larger size while decreases in temperature reduce the metabolic normalisation.
+On the other hand, our results indicate that *Hippopodina* sp.
+presents a reduction in metabolic benefits of attaining a larger size at colder temperatures.
+Thus, for this encrusting species, one could predict that individuals living at higher temperatures should allocate more resources to attaining larger sizes (as opposed to reproduction, for instance) while individuals living at lower temperatures will be under less pressure to allocate resources to size.
+Importantly, though, we note that there are many extrinsic factors that may affect selection on body size (e.g.
+predation, competition, etc).
+Yet, such differentials in the costs and benefits of attaining large size may help explain why in species living in cold temperatures present a slower pace of life (e.g.
+`r myCite('rosa2010jmba')`).
+Moreover, they may represent necessary improvements on current attempts to scale up energy from individuals to higher levels of organisation (e.g.
+`r myCite(c('allen2005','yvondurocher2012b','barneche2014ele'))`).
 
-**A little example to consider.**
+Importantly, some of our results are also consistent with existing theoretical predictions. 
+- Temperature dependence and mass-scaling of *B. stolonifera* consistent with MTE. Why this species only?
 
-- Hot temperature, a = 0.6, B = 0.4. Cold temperature, a = 0.4, B = 0.6.  
-- Energetic savings of being 2 units of size versus 10 units in size  
-- Metabolic cost of being alive  for  2 unit size critter at high temperature is 0.79 or ~40% of total mass  
-- Metabolic cost of being alive for 2 unit size critter at low temperature is 0.60 or ~30% of total mass  
-- Metabolic cost of being alive for 10 unit size critter at high temperature is 1.5 or ~15% of total mass  
-- Metabolic cost of being alive for 10 unit size critter at low temperature is 1.59 or ~16% of total mass  
-- So 5 fold increase in size gets you a 2.667 increase in efficiency at high temperatures but same increase in size at low temperature gets you only a 1.85 increase in efficiency.
-That's a pretty big deal and we should say something about it!
-- Another example, at high temperatures, 10 colonies of 100 units in size will have a cumulative metabolism of 37 while 100 colonies of 10 units in size will have a cumulative metabolism of 150.
-(same overall mass).
-At low temperatures with same colony sizes and numbers, they are  63 and 159.
-The ratio goes from 4.05 to 2.52.
-Massive change in metabolism if fragmentation occurs at high temperature, less change if fragmentation occurs at low temperatures.
-
-New paras:  
-- Many organisms exhibit multiple growth forms (encrusting and arborescent); disadvantage of arborescent is that there is no increase in efficiency with size.  
-- Lack of evidence for WBE. Mass-scaling exponent not universal across the species analysed, and clear interactions with temperature.  
-- Winners and losers across latitudinal gradients, and under increased temperatures. Arborsecents are predicted to be less common in warm waters,  
-- Pocilliporids fast growing arborescent coral; massive corals are slow growing and something… This provides a mechanistic explanation for why arborescent corals grow fast, but attain smaller sizes than massive corals.
-- Rosa, R., and Seibel, B. A. 2009. Slow pace of life of the Antarctic colossal squid. J. Mar. Biol. Ass. U. K.
+Conclusions paragraph
 
@@
 Results for each species are represented in groups of four plots, with blue and red colours representing actual data and parameter estimates at 10 ˚C and 25 ˚C, respectively.
-*Top-left*: posterior distributions of mass-scaling exponents, with horizontal bars representing 95% credible intervals.
-*Bottom-right*: posterior distributions of metabolic normalisations, with vertical bars representing 95% credible intervals.
+*Top-left* and *bottom-right*: posterior probabilities of mass-scaling exponents and metabolic normalisations (respectively), with bars over curves representing 95% credible intervals, and numbers representing average estimates.
 *Bottom-left*: covariance between mass-scaling exponents and metabolic normalisations for each of the 6,000 MCMC iterations (translucent circles), with overlaid density heat-map.
@@
 Regression fits were plotted using the average from the posterior distributions.
-In  probability posterior plots (top-left and bottom-right), smooth probability curves were approximated from the posterior density curves using very small bins (see electronic supplementary material).
+In probability posterior plots (*top-left* and *bottom-right*), smooth probability curves (*n* = 511 values) were approximated from posterior density curves (*n* = 512 values); differences between adjacent $x$ values were multiplied by the average of adjacent $y$ values (i.e.
+$(x_{i+1} - x_{i}) \times ((y_{i+1} - y_{i})/2)$) thus insuring that the integral of the probability curve $\approx$ 1.
 

### 2016-02-01 text/MS.Rmd
few minor corrections and additions to text

@@
 In this model, different from the model in equation 2, the mass scaling slope is no longer allowed to vary across different species, but is still allowed to vary between different temperatures (1 $\times$ 2 = 2 parameters).
-The metabolic normalisation, however, is still allowed to vary across species (4 $\times$ 1 = 4 parameters) and between temperatures (1 additional parameters), and random effects were maintained (i.e.
+The metabolic normalisation, however, is still allowed to vary across species (4 $\times$ 1 = 4 parameters) and between temperatures (1 additional parameter), and random effects were maintained (i.e.
 3 hyper parameters).
@@
 Thus, for this encrusting species, one could predict that individuals living at higher temperatures should allocate more resources to attaining larger sizes (as opposed to reproduction, for instance) while individuals living at lower temperatures will be under less pressure to allocate resources to size.
-Importantly, though, we note that there are many extrinsic factors that may affect selection on body size (e.g.
-predation, competition, etc).
-Yet, such differentials in the costs and benefits of attaining large size may help explain why in species living in cold temperatures present a slower pace of life (e.g.
+Importantly, though, we note that there are many extrinsic factors that may affect abundance and selection on body size (e.g.
+predation, competition, etc; `r myCite('barneche2016procb')`).
+Yet, such differentials in the costs and benefits of attaining large size may help explain why some species living in cold temperatures present a slower pace of life (e.g.
 `r myCite('rosa2010jmba')`).
 Moreover, they may represent necessary improvements on current attempts to scale up energy from individuals to higher levels of organisation (e.g.
-`r myCite(c('allen2005','yvondurocher2012b','barneche2014ele'))`).
+`r myCite(c('allen2005','yvondurocher2012b','barneche2014ele','barneche2016procb'))`).
 
@@
 Regression fits were plotted using the average from the posterior distributions.
-In probability posterior plots (*top-left* and *bottom-right*), smooth probability curves (*n* = 511 values) were approximated from posterior density curves (*n* = 512 values); differences between adjacent $x$ values were multiplied by the average of adjacent $y$ values (i.e.
-$(x_{i+1} - x_{i}) \times ((y_{i+1} - y_{i})/2)$) thus insuring that the integral of the probability curve $\approx$ 1.
+In probability posterior plots (*top-left* and *bottom-right*), smooth probability curves (*n* = 511 values) were approximated from posterior probability density curves (*n* = 512 values); differences between adjacent $x$ values (i.e.
+parameter estimates) were multiplied by the average of adjacent $y$ values (i.e.
+respective probability density values; $(x_{i+1} - x_{i}) \times ((y_{i+1} - y_{i})/2)$) thus insuring that the integral of the probability curve $\approx$ 1.
 

### 2016-02-08 text/MS.Rmd
implemented LOO model selection and updated text, tables, figures, numbers and functions

@@
 myRefs   <-  listBibs()
-exporBibs(myRefs, erase=FALSE)
+exporBibs(myRefs, erase=TRUE)
 ```
@@
 
-Metabolic theories have been lauded as having the potential to unite hierarchical levels of biological organisation, from individuals to ecosystems `r myCite(c('nisbet2000janimecol', 'brown2004', 'vandermeer2006tree'))`.
+Metabolic theories provide a potential to unite hierarchical levels of biological organisation, from individuals to ecosystems `r myCite(c('nisbet2000janimecol', 'brown2004', 'vandermeer2006tree'))`.
 Although definitive tests of their underlying assumptions remain rare `r myCite(c('kearney2012amnat', 'newberry2015ploscomputbiol'))`, myriad studies examined their predictions with regards to the relationship between individual level metabolic rates, $B_i$, and individual body mass, $M_i$ `r myCite(c('kleiber1932', 'kleiber1961', 'peters1983', 'schmidtnielsen1984', 'glazier2005biolrev', 'delong2010', 'glazier2010biolrev'))`.
 The relationship between $B_i$ and $M_i$ is typically described by a power function of the form $B_i = B_oM_i^{\alpha}$, where $B_o$ is a normalisation constant independent of mass and $\alpha$ is a dimensionless scaling exponent.
-For most animals, individual metabolic rate scales sub-linearly with body mass, with $\alpha$ estimates falling between 0.5 and 1, and averaging around 0.75 `r myCite(c('kleiber1961', 'savage2004', 'glazier2005biolrev', 'delong2010'))`.
-Metabolic rates are also influenced by temperature.
-Two prevalent competing metabolic theories, the fractal geometry model `r myCite(c('west1997science', 'west1999science'))`, and  the Dynamic Energy Budget (DEB) theory for metabolic organisation `r myCite('kooijman2009book')`, provide opposing views on the role of temperature on metabolic rates and the relationship between the mass scaling of metabolic rates and temperature: the former assumes that temperature affects the metabolic normalisation only `r myCite('gillooly2001')`, i.e.
-mass-scaling and temperature dependence are independent; contrarily, the latter predicts changing relationships between mass scaling and temperature dependence through the effects of temperature on assimilation, maintenance, growth, and maturation `r myCite('kooijman2009book')`.
-
-Empirical studies of the intra-specific association between $B_i$ and $M_i$ have revealed that the value of $\alpha$ is associated with many biotic and abiotic variables, including pH `r myCite('vaca2010biollett')`, temperature `r myCite('killen2010ele')`, predation `r myCite('glazier2011ecolmongr')`, inbreeding `r myCite('ketola2012jlinsoc')`, and a range of others (for reviews see `r myCite(c('glazier2005biolrev', 'white2013jcompphysb'))`).
-Similarly, inter-specific studies have shown that the value of $\alpha$ varies among taxa and/or locations (e.g.
-`r myCite(c('lovegrove2000amnat', 'white2012procb', 'barneche2014ele'))`), with metabolic level (e.g.
-`r myCite('glazier2010biolrev')`), and sometimes with body mass (e.g.
-`r myCite(c('white2009evolution', 'kolokotrones2010nature'))`).
-While these intra- and inter-specific studies demonstrate clearly that the value of $\alpha$ is not fixed, they each share a common limitation: each study tests for an association between $\alpha$ and a biotic or abiotic variable, using a value of $\alpha$ that is estimated from the association between $B_i$ and $M_i$ for individuals or species of different size.
+Numerous studies have demonstrated that, for most animals, individual metabolic rate scales sub-linearly with body mass, with $\alpha$ estimates varying with many biotic and abiotic variables (e.g.
+`r myCite(c('lovegrove2000amnat', 'white2009evolution', 'killen2010ele', 'kolokotrones2010nature', 'vaca2010biollett', 'glazier2011ecolmongr', 'ketola2012jlinsoc', 'white2012procb', 'white2013jcompphysb', 'barneche2014ele'))`), generally falling between 0.5 and 1, and averaging around 0.75.
+While these studies demonstrate clearly that the value of $\alpha$ is not fixed, they each share a common limitation: each study tests for an association between $\alpha$ and a biotic or abiotic variable, using a value of $\alpha$ that is estimated from the association between $B_i$ and $M_i$ for individuals or species of different size.
 Such correlative studies cannot address the causal association between $B_i$ and $M_i$, because $M_i$ is correlated with a range of other environmental, physiological, and life history variables (see e.g.
@@
 
-In contrast to studies that make use of intact individual organisms, colonial animals provide an opportunity for definitive tests of metabolic theories.
+In contrast to studies that make use of intact individual organisms, colonial animals provide an opportunity for more definitive tests of metabolic theories.
 Colony size can be manipulated experimentally, hence allowing for the potential to remove the confounding effects of body size and life-history traits (e.g.
@@
 For example, the fractal geometry model `r myCite(c('west1997science', 'west1999science'))` predicts that $\alpha = n/(n + 1)$ for a $n$-dimensional organism, hence a value of $\alpha$ = 0.75 is predicted for three-dimensional animals while a value of $\alpha$ = 0.67 is predicted for two-dimensional animals (but see `r myCite(c('enquist2009nature', 'savage2008ploscompbiol', 'banavar2010pnas', 'kolokotrones2010nature', 'price2007pnas', 'koontz2009amjbot'))`).
-On the other hand, DEB predicts that metabolic scaling in colonial organisms are dependent upon growth rate (which is in turn influenced by temperature) and the relative contributions of growing and non-growing zooids to whole-colony metabolism.
+Moreover, experimental size manipulation in colonial organisms allows for tests of how ecological processes affect metabolic rates.
+For instance, it has been demonstrated that mimicking partial predation through size manipulation yields different responses in growth and reproductive onset in different species of encrusting and arborescent bryozoans `r myCite(c('bone2005jembe', 'hart2009ecologyb'))`.
+However, to our knowledge, few studies have explicitly investigated the effects of partial predation on individual metabolic rates (e.g.
+`r myCite('white2011amnat')`).
+
+Metabolic rates are also influenced by temperature, typically following an exponential function `r myCite(c('gillooly2001', 'kooijman2009book'))`.
+Existing metabolic theories assume nevertheless different paths through which temperature affects metabolic rates.
+For example, the fractal geometry model `r myCite(c('west1997science', 'west1999science'))` assumes that temperature affects the metabolic normalisation only `r myCite(c('gillooly2001', 'brown2004'))`, i.e.
+mass-scaling and temperature dependence of metabolic rates are fundamentally independent; contrarily, the Dynamic Energy Budget (DEB) theory for metabolic organisation `r myCite('kooijman2009book')` predicts changing relationships between mass scaling and temperature dependence through the effects of temperature on assimilation, maintenance, growth, and maturation `r myCite('kooijman2009book')`.
+Whether temperature dependence and mass scaling truly interact remains a contentious issue, though recent empirical evidence indicates that they do, at least in fishes `r myCite('killen2010ele')`.
+Testing such interaction is of great ecological significance because different mass-scaling exponents under different temperatures would potentially indicate contrasting life-history trajectories `r myCite('sebens1987annurevecolsys')` for a given species under different environments **(REFS??)**.
 
 In the present study we experimentally manipulate size in different colonial marine organisms from two phyla (Bryozoa [Ectoprocta] and Porifera) in order to test whether the mass scaling and temperature dependence of metabolic rates interact `r myCite(c('white2011amnat', 'kearney2012amnat'))`.
-In doing so, we explicitly test if these interactions are general or species specific.
+In doing so, we explicitly test if these interactions are general or species specific using a Bayesian model selection procedure.
+Finally, we discuss the relevance of our results to the life history of different species living under different environments.
 
@@
 Individual metabolic rates (**provide units**, *n* = `r nrow(metRates)`), $B_i$, were measured as rate of oxygen consumption, $\dot{V}\textrm{O}_\textrm{2}$ of colony fragments was measured at 10 or 25 ˚C using a 144-channel PreSens Sensor Dish Reader (SDR, AS-1 Scientific, Wellington, New Zealand) according to standard techniques `r myCite(c('koster2008meps', 'lighton2008book', 'white2011amnat'))`.
-$\dot{V}\textrm{O}_\textrm{2}$ was measured by placing the colony or fragment in a glass vial containing 0.2$\mu$m filtered sea water and a non-consumptive $\textrm{O}_\textrm{2}$ sensor spot and calculated from the rate of change of $\textrm{O}_\textrm{2}$ saturation ($m_a$, % h^‑1^) as:
+$\dot{V}\textrm{O}_\textrm{2}$ was measured by placing the colony or fragment in a glass vial containing 0.2$\mu$m filtered sea water and a non-consumptive $\textrm{O}_\textrm{2}$ sensor spot and calculated from the rate of change of $\textrm{O}_\textrm{2}$ saturation ($m_a$, \% h^‑1^) as:
 
@@
 
-where $m_b$ is the rate of change of $\textrm{O}_\textrm{2}$ saturation for blank vials containing no bryozoans (% h^-1^), ${\beta}_{\textrm{O}_\textrm{2}}$ is the oxygen capacitance of air-saturated sea water at 10 ˚C (6.26 ml l^-1^) or 24 ˚C (4.615 ml l^-1^) [**double-check these numbers.
+where $m_b$ is the rate of change of $\textrm{O}_\textrm{2}$ saturation for blank vials containing no bryozoans (\% h^-1^), ${\beta}_{\textrm{O}_\textrm{2}}$ is the oxygen capacitance of air-saturated sea water at 10 ˚C (6.26 ml l^-1^) or 24 ˚C (4.615 ml l^-1^) [**double-check these numbers.
 Shouldn't the second be 25 ˚C?**] `r myCite('cameron1986book')` and $V$ is water volume (chambers were 0.005 l in volume and water volume was calculated by subtracting the volume of acetate and animals).
-At least four blank vials were recorded simultaneously with each 24-well plate to account for microbial oxygen consumption and sensor spots were calibrated with air-saturated seawater (100 % A.S.) and water containing 2% sodium sulfite (0% A.S.).
+At least four blank vials were recorded simultaneously with each 24-well plate to account for microbial oxygen consumption and sensor spots were calibrated with air-saturated seawater (100\% A.S.) and water containing 2\% sodium sulfite (0\% A.S.).
 Following measurement of $\dot{V}\textrm{O}_\textrm{2}$, colonies or fragments were blotted dry and weighed to 0.1 mg (Sartorius A 200 S, Sartorius AG, Göttingen, Germany).
@@
 *3. Statistical analyses*  
-**Things to possibly improve -- since we use a Bayesian approach, why let parameters be estimated from data (i.e.
-vague priors)?
-Why not adopt strong priors?
-Maybe letting it estimate is better to check whether parameters follow one or another theory?**
-**Things to possibly improve -- Include more models for model selection?
-E.g.
-two-way models with species*temp interaction or species*mass?
-lme4 already indicates that they are not better, I'm pretty sure the Bayesian DIC will follow results along the same lines.
-We'd be covering our bases better beforehand before a reviewer asks for this**
-
 To test for the interacting effects of body mass and temperature on metabolic rate across species, we fitted a three-way interaction linear model using log-transformed data
@@
 where mass (mg), $\textrm{ln}M_i$, is a continuous variable, while temperature (two levels: 10 ˚C and 25 ˚C) and species (four levels corresponding to the four species) are categorical variables.
-With this model, we allow each species to have a different mass-scaling slope that may change with temperature (4 $\times$ 2 = 8 parameters); it also accounts for species- and temperature-specific deviations on the metabolic normalisation (i.e.
-statistical model intercept; 4 $\times$ 2 = 8 parameters).
-We also included experimental conveniences (plate reader, $\Delta P_r$, and date, $\Delta D$) as random effects on the metabolic normalisation to help control for natural experimental variability (3 additional hyper parameters corresponding to the standard deviation of $\Delta P_r$, $\Delta D$ and random-effects residual).
+With this model, we allow each species to have a different mass-scaling slope that may change with temperature; we also account for species- and temperature-specific deviations on the metabolic normalisation (i.e.
+statistical model).
+We also included date of experiment, $\Delta D$ as a random effect on the metabolic normalisation to help control for natural experimental variability.
+We then performed a Bayesian model selection procedure (see below) by comparing the above full model with a series of nested two- and one-way interaction models in order to formally test for mass-temperature interactions in metabolic rates among different species.
 
-We then performed a model selection by comparing the statistical model in equation 2 with a simpler two-way interaction model of similar structure,
-
-$$\textrm{ln}B_i \sim \textrm{ln}M_i * Temperature + Species$$. (3)
-
-In this model, different from the model in equation 2, the mass scaling slope is no longer allowed to vary across different species, but is still allowed to vary between different temperatures (1 $\times$ 2 = 2 parameters).
-The metabolic normalisation, however, is still allowed to vary across species (4 $\times$ 1 = 4 parameters) and between temperatures (1 additional parameter), and random effects were maintained (i.e.
-3 hyper parameters).
-Equation 3 represents a statistical model nested in the model represented by equation 2, and therefore their performance can be directly assessed `r myCite('zuur2009book')`.
-Such comparison allowed us to formally test our main hypothesis of significant mass-temperature interactions across different species.
-
-Both models were fitted in a Bayesian framework by calling *JAGS* version 3.4.0 from the R package *R2jags* version 0.05-03 `r myCite('r2jags')` in order to derive posterior distributions and associated 95% credible intervals (CIs) for the fitted parameters.
+Both models were fitted in a Bayesian framework by calling *JAGS* version 3.4.0 from the R package *R2jags* version 0.05-03 `r myCite('r2jags')` in order to derive posterior distributions and associated 95\% credible intervals (CIs) for the fitted parameters.
 Random effects were assumed to be normally distributed, with means of 0.
@@
 locally uniform over the region supported by the likelihood) `r myCite('kruschke2014book')`.
-The posterior distributions of model parameters were estimated using Markov chain Monte Carlo (MCMC) methods by constructing three chains of 500,000 steps each.
-Chains were thinned using a 250-step interval, so a total of 6,000 steps were retained to estimate posterior distributions (i.e.
-3 $\times$ 500,000/250 = 6,000).
-Models were compared adopting Deviance information criterion (DIC), which are commonly used in Bayesian model selection and are analogously similar to the Akaike information criterion (AIC), as this method also penalises for additional model parameters `r myCite('hooten2015ecology')`.
-We adopted a conservative approach, and models were considered significantly different only if DIC units differed by > 10 units `r myCite('spiegelhalter2002jrss')`.
+The posterior distributions of model parameters were estimated using Markov chain Monte Carlo (MCMC) methods by constructing five chains of 1,000,000 steps each, including 500,000-step burn-in periods.
+Chains were thinned using a 250-step interval, so a total of 10,000 steps were retained to estimate posterior distributions (i.e.
+5 $\times$ (1,000,000 - 500,000)/250 = 10,000).
+Each nested model was compared against the full model using leave-one-out cross-validation (LOO), which, similar to widely applicable information criterion (WAIC), is a fully Bayesian model selection procedure for estimating pointwise out-of-sample prediction accuracy using the log-likelihood evaluated at the posterior simulations of the parameter values `r myCite('hooten2015ecology')`.
+Model significance was assessed by calculating a *z*-score from the average difference $\pm$ SE between expected log pointwise predictive density of both models, and models were considered significantly different at $\alpha \le$ 0.05.
+Model comparison was implemented using the R package *loo* version 0.1.3 `r myCite('loo')`.
 All data and code necessary to reproduce this paper, its analyses, tables and figures can be obtained on GitHub https://github.com/dbarneche/MTRBrEs.
@@
 
-Our model selection approach indicated that the three-way interaction model (equation 2) performed significantly better than the two-way interaction model (equation 3) ($\Delta$d.f.
-= `r deltaDF`; $\Delta$DIC = `r rounded(deltaDIC)`), thus implying that different species present different interactions between mass-scaling and temperature dependence.
-Particularly, the best model yields no interaction between mass scaling and temperature dependence for *Bugula stolonifera* and the Microcionidae sponge (large overlap between 95% credible intervals for the mass-scaling slopes, $\alpha$) (Figure 1).
+Our model selection approach indicated that the three-way interaction model (equation 2) performed significantly better than all nested two- and one-way interaction models (Table 2), thus implying that different species present different interactions between mass-scaling and temperature dependence.
+Particularly, the best model yields no interaction between mass scaling and temperature dependence for *Bugula stolonifera* and the Microcionidae sponge (large overlap between 95\% credible intervals for the mass-scaling slopes, $\alpha$) (Figure 1).
 Still, while the mass-scaling slopes are statistically the same between temperatures for these species, there is evidence for higher metabolic normalisations (i.e.
@@
 It is worth noting that this $Q_{10}$ value is equivalent to an activation energy of `r rounded(q10andErs$er[q10andErs$species == 'Stolonifera'], 2)` eV (i.e.
-$E_r = \frac{k \times 283.15 \times 298.15 \times 1.5 \times \textrm{ln}Q_{10}}{(298.15 - 283.15)}$, where $E_r$ is an activation energy and $k$ is the Boltzmann constant, 8.62 $\times$ 10^-5^ eV K^-1^), which is consistent with the average 06 -- 0.7 eV reported in the Metabolic Theory of Ecology literature `r myCite('yvondurocher2012a')`.
+$E_r = \frac{k \times 283.15 \times 298.15 \times 1.5 \times \textrm{ln}Q_{10}}{(298.15 - 283.15)}$, where $E_r$ is an activation energy and $k$ is the Boltzmann constant, 8.62 $\times$ 10^-5^ eV K^-1^), which is consistent with the average 06 -- 0.7 eV reported in the Metabolic Theory of Ecology literature `r myCite(c('gillooly2001','yvondurocher2012a'))`.
 Different from the above species, the best model indicates that *Bugula neritina* and *Hippopodina* sp.
@@
 For example, *Hippopodina* sp.
-shows exponents that encompass both 0.67 (**DEB prediction??**; Fractal geometry model prediction for 2-D organisms; `r myCite('koontz2009amjbot')`) and 0.75 (Fractal geometry model prediction for 3-D organisms `r myCite(c('west1997science', 'west1999science'))`) at 10 ˚C (95% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's10_97.5'], 2)`), but that are significantly shallower than 0.75 at 25 ˚C (95% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's25_97.5'], 2)`).
-On the other hand, *Bugula neritina* presents scaling exponents that overlap both 0.67 and 0.75 at 25 ˚C (95% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's25_97.5'], 2)`), but are significantly shallower than these values at 10 ˚C (95% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's10_97.5'], 2)`).
-Microcionidae presents exponents that only overlap 0.67 at 10 ˚C, but not at 25 ˚C (95% CI at 10 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's10_97.5'], 2)`; 95% CI at 25 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's25_97.5'], 2)`).
-Finally, *Bugula stolonifera* presents values that overlap both 0.67 and 0.75 at both temperatures (95% CI at 10 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's10_97.5'], 2)`; 95% CI at 25 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's25_97.5'], 2)`).
+shows exponents that encompass both 0.67 (Fractal geometry model prediction for 2-D organisms; `r myCite('koontz2009amjbot')`) and 0.75 (Fractal geometry model prediction for 3-D organisms; `r myCite(c('west1997science', 'west1999science'))`) at 10 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's10_97.5'], 2)`), but that are significantly shallower than 0.75 at 25 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's25_97.5'], 2)`).
+On the other hand, *Bugula neritina* presents scaling exponents that overlap both 0.67 and 0.75 at 25 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's25_97.5'], 2)`), but are significantly shallower than these values at 10 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's10_97.5'], 2)`).
+Microcionidae presents exponents that only overlap 0.67 at 10 ˚C, but not at 25 ˚C (95\% CI at 10 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's10_97.5'], 2)`; 95\% CI at 25 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's25_97.5'], 2)`).
+Finally, *Bugula stolonifera* presents values that overlap both 0.67 and 0.75 at both temperatures (95\% CI at 10 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's10_97.5'], 2)`; 95\% CI at 25 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's25_97.5'], 2)`).
 
@@
 `r myCite('killen2010ele')`), yielding further support to the need for new metabolic theories (or reformulation of current ones) that formally incorporate interactive effects between mass scaling and temperature dependence of biological rates.
-*Currently, the Dynamic Energy Budget (DEB) theory predicts individual metabolism on the basis of energy allocation to assimilation, maintenance, growth, and maturation `r myCite('kooijman2009book')`, yielding a complex three-way interaction between growth form, temperature, and size `r myCite(c('white2011amnat', 'kearney2012amnat'))`.* **If we want to keep this comment on DEB, we will have to formally present how and what predictions DEB makes for changes in mass-scaling between different temperatures**.
 Our results, however, do not indicate consistent mass-temperature interactions across growth forms.
@@
 
-Importantly, some of our results are also consistent with existing theoretical predictions. 
-- Temperature dependence and mass-scaling of *B. stolonifera* consistent with MTE. Why this species only?
+In the wild, partial predation is a common phenomenon experienced by colonial organisms, which must present different defence strategies `r myCite('sebens1987annurevecolsys')`.
+Our findings suggest that, for some species, predation might modulate the energetic demands of individuals of similar body mass in different environments (assuming that the size-manipulation approach herein adopted may serve as a surrogate for natural field predation; `r myCite('bone2005jembe')`), thus affecting life-history trajectories.
+For instance, at 25 ˚C, a 400-gram individual of *Bugula stolonifera* will spend `r rounded(metabolicDiff$metabolicDiff[metabolicDiff$species == 'Stolonifera'], 2)`-fold more energy than a similar-sized individual at 10 ˚C.
+Conversely, due to the estimated positive interaction between mass-scaling and temperature, a similar-sized individual of *B.
+neritina* will spend `r rounded(metabolicDiff$metabolicDiff[metabolicDiff$species == 'Bugula'], 2)`-fold more energy at 25 ˚C than at 10 ˚C.
+These differences, for instance, may help determine which species is a better competitor in different environments, as individual metabolic rates will also affect growth rates and time of reproduction onset `r myCite(c('sebens1987annurevecolsys', 'hart2009ecologyb'))`.
+It is worth noting that while we do not explicitly control for size manipulation in our study (i.e.
+evaluate the same model for non-manipulated individuals), previous evidence suggests that there is no effect of fragmentation on energetic demands of colonial organisms `r myCite('white2011amnat')`.
 
-Conclusions paragraph
+Here we have demonstrated that the effects of temperature interact in various ways with the mass-scaling of metabolic rates in different species of colonial marine invertebrates encompassing two distinct phyla.
+By adopting a size-manipulation approach, we achieved a more definitive test of mass-scaling of metabolic rates for cancelling the confounding effects of size and age, and by mimicking predation, which constitutes a fundamental ecological process in the wild.
+Mass-scaling exponents vary for different organisms, and are not consistent across growth forms nor phylogeny.
+These results suggest that more accurate predictions of abundance and biomass production at higher levels of organisation, from populations to ecosystems, will required a species-specific understanding of the interacting effects of mass scaling and temperature dependence, perhaps leading to improvements of current metabolic theories, or even the development of new ones.
 
@@
 
-**Figures legends**
+*Table 2.* The best candidate model was chosen based on a selection procedure with nested models where non-significant fixed effects were successively removed based on leave-one-out cross-validation.
+Model significance was assessed by calculating a *z*-score from the average difference $\pm$ SE between expected log pointwise predictive density of both models ($\widehat{elpd_{diff}}$).
+
+**Figures**
 
@@
 Results for each species are represented in groups of four plots, with blue and red colours representing actual data and parameter estimates at 10 ˚C and 25 ˚C, respectively.
-*Top-left* and *bottom-right*: posterior probabilities of mass-scaling exponents and metabolic normalisations (respectively), with bars over curves representing 95% credible intervals, and numbers representing average estimates.
-*Bottom-left*: covariance between mass-scaling exponents and metabolic normalisations for each of the 6,000 MCMC iterations (translucent circles), with overlaid density heat-map.
+*Top-left* and *bottom-right*: posterior probabilities of mass-scaling exponents and metabolic normalisations (respectively), with bars over curves representing 95\% credible intervals, and numbers representing average estimates.
+*Bottom-left*: covariance between mass-scaling exponents and metabolic normalisations for each of the 10,000 MCMC iterations (translucent circles), with overlaid density heat-map.
 *Top-right*: regressions of individual metabolic rate, $\textrm{ln}B_i$, as functions of individual body mass, $\textrm{ln}M_i$.
@@
 
-**Figures**
-
 **References**

### 2016-02-08 R/functions-analyses.R

@@ -52,17 +52,92 @@ myCite  <-  function(citationsVec) {
 }
 
 ###############
+# DATA ANALYSIS
+###############
+jagsModel  <-  function(run=TRUE, outputFileName) {
+  jmod                                <-  vector(mode='character')
+  jmod[(length(jmod) + 1)]  <-  'model {'
+  jmod[(length(jmod) + 1)]  <-  '\t# priors for fixed effects in metabolic rate analysis'
+  jmod[(length(jmod) + 1)]  <-  '\tfor (i in 1:K) {'
+  jmod[(length(jmod) + 1)]  <-  '\t\tbeta[i] ~ dnorm(0, 0.0001)'
+  jmod[(length(jmod) + 1)]  <-  '\t}'
+  jmod[(length(jmod) + 1)]  <-  '\ttauB  ~   dgamma(1.0E-3,1.0E-3)'
+  jmod[(length(jmod) + 1)]  <-  '\tvarB  <-  1/tauB # variance, equivalent to inverse(tauB)'
+  jmod[(length(jmod) + 1)]  <-  '\n\t#linear regression for metabolic rate'
+  jmod[(length(jmod) + 1)]  <-  '\tfor(i in 1:length(lnRate)) {'
+  jmod[(length(jmod) + 1)]  <-  '\t\tlnRate[i]  ~  dnorm(muB[i], tauB)'
+  jmod[(length(jmod) + 1)]  <-  '\t\tmuB[i]  <-  inprod(beta[], tjagsModelMatrix[i,])'
+  jmod[(length(jmod) + 1)]  <-  '\t\tlogLik[i]  <- - log(tauB) + log(2*pi)  + pow(lnRate[i]-muB[i],2) * tauB'
+  jmod[(length(jmod) + 1)]  <-  '\t}'
+  
+  if(run) {
+      jmod                      <-  sub('muB[i]  <-  ', 'muB[i]  <-  r[runJ[i]] + ', jmod, fixed=TRUE)
+      jmod[(length(jmod) + 1)]  <-  '\n\t# priors for random effects at the run level'
+      jmod[(length(jmod) + 1)]  <-  '\tfor(i in 1:max(runJ)) {'
+      jmod[(length(jmod) + 1)]  <-  '\t\tr[i] ~ dnorm(0, tauR)'
+      jmod[(length(jmod) + 1)]  <-  '\t}'
+      jmod[(length(jmod) + 1)]  <-  '\ttauR  ~   dgamma(1.0E-3,1.0E-3)'
[…]

### 2016-02-08 R/functions-figures.R

@@ -87,7 +87,7 @@ getProb  <-  function(densObjt) {
   densObjt
 }
 
-plotTriad  <-  function(species=species, spNames=spNames, tmat, y1=TRUE, x1=TRUE, y2=TRUE, x2=TRUE, y3=TRUE, x3=TRUE) {
+plotTriad  <-  function(species=species, spNames=spNames, tmat, x1=TRUE, y1=TRUE, x2=TRUE, y2=TRUE, x3=TRUE, y3=TRUE, x4=TRUE, y4=TRUE) {
   col10     <-  colorRampPalette(rev(brewer.pal(9, 'Blues')))
   col10     <-  col10(32)
   col10[1]  <-  '#FFFFFF'
@@ -112,50 +112,62 @@ plotTriad  <-  function(species=species, spNames=spNames, tmat, y1=TRUE, x1=TRUE
   xDens10    <-  getProb(density(s10))
   xDens25    <-  getProb(density(s25))
   
-  if(y1) {
-    y1  <-  substitute('Posterior probability '%*%' 10'^{-3})
-  } else {
-    y1  <-  ''
-  }
-
   if(x1) {
-    x1  <-  substitute('Mass scaling exponent, '*italic(alpha))
+    x1l  <-  substitute('Mass scaling exponent, '*italic(alpha))
   } else {
-    x1  <-  ''
+    x1l  <-  ''
   }
 
-  if(y2) {
-    y2  <-  substitute('ln '*italic('B'['o'])*' (units?)')
+  if(y1) {
+    y1l  <-  substitute('Posterior prob. '%*%' 10'^{-3})
   } else {
-    y2  <-  ''
+    y1l  <-  ''
   }
 
   if(x2) {
-    x2  <-  substitute('Posterior probability '%*%' 10'^{-3})
+    x2l  <-  substitute('ln Mass, '*italic('M'['i'])*' (mg)')
   } else {
-    x2  <-  ''
+    x2l  <-  ''
   }
 
-  if(y3) {
-    y3  <-  substitute('ln Rates, '*italic('B'['i'])*' (units?)')
+  if(y2) {
+    y2l  <-  substitute('ln Rates, '*italic('B'['i'])*' (units?)')
   } else {
-    y3  <-  ''
+    y2l  <-  ''
   }
 
[…]

### 2016-02-08 analyses.R

@@ -1,58 +1,49 @@
 library(R2jags)
-library(lme4)
+library(loo)
+library(plyr)
 
 rm(list=ls())
 source('paths.R')
 source('dataManipulation.R')
 source('R/functions-analyses.R')
 
-# quickly test for significance of fixed effects using lme4
-modelLmer1  <-  lmer(lnRate ~ Species * lnMass * Temp + (1 | Run) + (1 | Plate), data=metRates, REML=TRUE)
-modelLmer2  <-  lmer(lnRate ~ Species + lnMass * Temp + (1 | Run) + (1 | Plate), data=metRates, REML=TRUE)
-modelLmer3  <-  lmer(lnRate ~ lnMass * Temp + (1 | Run) + (1 | Plate), data=metRates, REML=TRUE)
-modelLmer4  <-  lmer(lnRate ~ Temp + (1 | Run) + (1 | Plate), data=metRates, REML=TRUE)
-modelLmer5  <-  lmer(lnRate ~ Species * Temp + lnMass + (1 | Run) + (1 | Plate), data=metRates, REML=TRUE)
-modelLmer6  <-  lmer(lnRate ~ Species * lnMass + Temp + (1 | Run) + (1 | Plate), data=metRates, REML=TRUE)
-anova(modelLmer1, modelLmer2, modelLmer3, modelLmer4) # all significant fixed effects
-anova(modelLmer1, modelLmer5) # three-way significantly better
-anova(modelLmer1, modelLmer6) # three-way significantly better
+# run models in lm in order to obtain model matrices of contrasts for jags
+lmModel1  <-  lm(lnRate ~ Species + lnMass + Temp + Species:lnMass + Species:Temp + Temp:lnMass + Temp:lnMass:Species, data=metRates)
+lmModel2  <-  lm(lnRate ~ Species + lnMass + Temp + Species:lnMass + Species:Temp + Temp:lnMass, data=metRates)
+lmModel3  <-  lm(lnRate ~ Species + lnMass + Temp + Species:lnMass + Species:Temp, data=metRates)
+lmModel4  <-  lm(lnRate ~ Species + lnMass + Temp + Species:lnMass + Temp:lnMass, data=metRates)
+lmModel5  <-  lm(lnRate ~ Species + lnMass + Temp + Species:Temp + Temp:lnMass, data=metRates)
[…]

### 2016-02-08 numbers.R

@@ -6,10 +6,6 @@ source('paths.R')
 source('R/functions-analyses.R')
 load('output/RDatafiles/analyses.RData')
 
-# model selection result
-deltaDF   <-  anova(modelLmer1, modelLmer2)['Chi Df'][2,]
-deltaDIC  <-  tfit2$BUGSoutput$DIC - tfit$BUGSoutput$DIC
-
 # get average estimates and 95% CI for each parameter
 species           <-  unique(metRates$Species)
 averageEstimates  <-  ldply(species, getAverageEstimates)
@@ -18,8 +14,11 @@ ci95Estimates     <-  ldply(species, get95CIEstimates)
 # compare deltas between temperatures for each species
 tempDeltas  <-  ddply(averageEstimates, .(species), getTempDeltas)
 
+# metabolic difference (in n-fold) between different temperatures for similar-sized animals
+metabolicDiff  <-  ddply(averageEstimates, .(species), getMetabolicDiff, lnMass=log(400))
+
 # q10-equivalents [ q10 == (r25 / r10) ^ (10/(25-10)) ] for Microcionidae and Bugula stolonifera
 q10andErs  <-  ddply(averageEstimates[averageEstimates$species %in% c('Sponge', 'Stolonifera'), ], .(species), getq10andEr)
 
-rm(list=ls()[!(ls() %in% c('deltaDF', 'deltaDIC', 'species', 'averageEstimates', 'ci95Estimates', 'tempDeltas', 'q10andErs'))])
+rm(list=ls()[!(ls() %in% c('species', 'averageEstimates', 'ci95Estimates', 'tempDeltas', 'metabolicDiff', 'q10andErs'))])
 save.image('output/RDatafiles/numbers.RData')

### 2016-02-08 tables.R

@@ -2,12 +2,26 @@ library(plyr)
 rm(list=ls())
 source('paths.R')
 source('dataManipulation.R')
+source('R/functions-analyses.R')
 source('R/functions-figures.R')
+load('output/RDatafiles/analyses.RData')
 
+# table 1
 table1  <-  ddply(metRates, .(Species, Temp), function(x) {
 	data.frame(minMass = rounded(min(exp(x$lnMass))),
 			   maxMass = rounded(max(exp(x$lnMass))),
 			   n         = nrow(x), stringsAsFactors=FALSE)
 })
-
 write.csv(table1, 'output/data/table1.csv', row.names=FALSE)
+
+# table 2
+# modelComparisonPVals  <-  function(data, modelList, verbose=TRUE) {
+#FixEffComps  <-  ddply(nested, .(complexModel, nestedModel), modelComparisonPVals, modelList=fixedSelec)
+
+table2  <-  ldply(2:9, function(x) {
+	modDiff  <-  compare(fixedSelec[[1]]$loo, fixedSelec[[x]]$loo)
+	data.frame(elpdDiff  =  rounded(modDiff[['elpd_diff']], 2),
+			   se        =  rounded(modDiff[['se']], 2),
+			   zPVal     =  cleanPvals(2*pnorm(-abs((modDiff[['elpd_diff']] - 0) / modDiff[['se']]))))
+})
+write.csv(table2, 'output/data/table2.csv', row.names=FALSE)

### 2016-03-17 R/functions-figures.R
updated units for metabolic rates in fig1

@@ -131,7 +131,7 @@ plotTriad  <-  function(species=species, spNames=spNames, tmat, x1=TRUE, y1=TRUE
   }
 
   if(y2) {
-    y2l  <-  substitute('ln Rates, '*italic('B'['i'])*' (units?)')
+    y2l  <-  substitute('ln Rates, '*italic('B'['i'])*' ('*mu*'l O'[2]*' h'^{-1}*')')
   } else {
     y2l  <-  ''
   }
@@ -143,7 +143,7 @@ plotTriad  <-  function(species=species, spNames=spNames, tmat, x1=TRUE, y1=TRUE
   }
 
   if(y3) {
-    y3l  <-  substitute('ln '*italic('B'['o'])*' (units?)')
+    y3l  <-  substitute('ln '*italic('B'['o']))
   } else {
     y3l  <-  ''
   }
@@ -155,7 +155,7 @@ plotTriad  <-  function(species=species, spNames=spNames, tmat, x1=TRUE, y1=TRUE
   }
 
   if(y4) {
-    y4l  <-  substitute('ln '*italic('B'['o'])*' (units?)')
+    y4l  <-  substitute('ln '*italic('B'['o']))
   } else {
     y4l  <-  ''
   }

### 2016-03-17 tables.R
updated tables with new results

@@ -1,4 +1,5 @@
 library(plyr)
+library(loo)
 rm(list=ls())
 source('paths.R')
 source('dataManipulation.R')
@@ -15,9 +16,6 @@ table1  <-  ddply(metRates, .(Species, Temp), function(x) {
 write.csv(table1, 'output/data/table1.csv', row.names=FALSE)
 
 # table 2
-# modelComparisonPVals  <-  function(data, modelList, verbose=TRUE) {
-#FixEffComps  <-  ddply(nested, .(complexModel, nestedModel), modelComparisonPVals, modelList=fixedSelec)
-
 table2  <-  ldply(2:9, function(x) {
 	modDiff  <-  compare(fixedSelec[[1]]$loo, fixedSelec[[x]]$loo)
 	data.frame(elpdDiff  =  rounded(modDiff[['elpd_diff']], 2),

### 2016-03-17 text/MS.Rmd
new version of MS with corrected dataset

@@
 ```
-**Temperature effects on mass-scaling exponent of metabolic rates: a manipulative test**
+**Temperature effects on mass-scaling exponents in colonial animals: a manipulative test**
 
@@
 1.
-Organismal metabolic rate often scales allometrically with body mass, both inter and intra-specifically.
-Metabolic rates are also known to increase with temperature, but most theoretical tests of mass scaling and temperature dependence have treated both phenomena independently.
+Metabolic theory posits that size and temperature are fundamental drivers of ecological patterns and processes, via their respective associations with metabolic rates.
+Metabolic rates are well known to increase allometrically with body mass and approximately exponentially with temperature, but most applications of metabolic theory treat these effects as independent.
 2.
-In the present study, we use a manipulative approach to test for an association between metabolic mass scaling and temperature dependence in four species belonging to two phyla of colonial marine invertebrates.
+Almost all studies of intraspecific scaling of unitary organisms must rely on ontogenetic differences in size to examine its relationship with metabolic rate, thereby potentially confounding age, nutrition or growth rate with size –- all of which can affect metabolic rate.
+In the present study, we use a manipulative approach to test for an association between metabolic mass scaling and temperature dependence in four species of colonial marine invertebrates belonging to two phyla.
 3.
-We show that mass-scaling exponents and temperature dependencies interact in different ways for different species.
-In so doing, we discuss the potentials of our results to multi-level scaling theory, showing that some species might be winners while others might be losers under different temperature scenarios.
+While increased temperature increased metabolic rate, we found no consistent effect of temperature or growth form on metabolic scaling; rather the effects were species-specific.
+We show that mass-scaling exponents and temperature dependencies interact in ways that are unanticipated by extant theory.
 4.
-These findings highlight the need for new metabolic theory that formally incorporate interactions between mass-scaling and temperature dependence in ectothermic organisms, in order to better explain documented variation in the mass-scaling exponent of metabolic rate, and lead to novel and testable predictions for associations between individual metabolism and community structure.
+The allometric scaling of metabolic rate that we observe in the present study contrasts with the isometric scaling of water clearance and feeding rates observed in some colonial marine invertebrates.
+Our findings suggest that the often observed competitive superiority of colonial organisms over unitary ones may arise because energy availability for growth and or reproduction (the difference between energy intake and expenditure) increases more strongly with size in colonial organisms.
 
@@
 
-Metabolic theories provide a potential to unite hierarchical levels of biological organisation, from individuals to ecosystems `r myCite(c('nisbet2000janimecol', 'brown2004', 'vandermeer2006tree'))`.
+Metabolic theories seek to unite hierarchical levels of biological organisation, from individuals to ecosystems `r myCite(c('nisbet2000janimecol', 'brown2004', 'vandermeer2006tree'))`.
 Although definitive tests of their underlying assumptions remain rare `r myCite(c('kearney2012amnat', 'newberry2015ploscomputbiol'))`, myriad studies examined their predictions with regards to the relationship between individual level metabolic rates, $B_i$, and individual body mass, $M_i$ `r myCite(c('kleiber1932', 'kleiber1961', 'peters1983', 'schmidtnielsen1984', 'glazier2005biolrev', 'delong2010', 'glazier2010biolrev'))`.
@@
 Numerous studies have demonstrated that, for most animals, individual metabolic rate scales sub-linearly with body mass, with $\alpha$ estimates varying with many biotic and abiotic variables (e.g.
-`r myCite(c('lovegrove2000amnat', 'white2009evolution', 'killen2010ele', 'kolokotrones2010nature', 'vaca2010biollett', 'glazier2011ecolmongr', 'ketola2012jlinsoc', 'white2012procb', 'white2013jcompphysb', 'barneche2014ele'))`), generally falling between 0.5 and 1, and averaging around 0.75.
+`r myCite(c('lovegrove2000amnat', 'white2009evolution', 'killen2010ele', 'kolokotrones2010nature', 'vaca2010biollett', 'glazier2011ecolmongr', 'ketola2012jlinsoc', 'white2012procb', 'white2013jcompphysb', 'barneche2014ele'))`), generally falling between 0.5 and 1.
 While these studies demonstrate clearly that the value of $\alpha$ is not fixed, they each share a common limitation: each study tests for an association between $\alpha$ and a biotic or abiotic variable, using a value of $\alpha$ that is estimated from the association between $B_i$ and $M_i$ for individuals or species of different size.
@@
 Furthermore, different colonial species may grow in different dimensions, and the mass-scaling exponent $\alpha$ is explicitly predicted to change with number of body dimensions `r myCite(c('nakaya2005procb', 'white2011amnat', 'kearney2012amnat'))`.
-For example, the fractal geometry model `r myCite(c('west1997science', 'west1999science'))` predicts that $\alpha = n/(n + 1)$ for a $n$-dimensional organism, hence a value of $\alpha$ = 0.75 is predicted for three-dimensional animals while a value of $\alpha$ = 0.67 is predicted for two-dimensional animals (but see `r myCite(c('enquist2009nature', 'savage2008ploscompbiol', 'banavar2010pnas', 'kolokotrones2010nature', 'price2007pnas', 'koontz2009amjbot'))`).
-Moreover, experimental size manipulation in colonial organisms allows for tests of how ecological processes affect metabolic rates.
-For instance, it has been demonstrated that mimicking partial predation through size manipulation yields different responses in growth and reproductive onset in different species of encrusting and arborescent bryozoans `r myCite(c('bone2005jembe', 'hart2009ecologyb'))`.
-However, to our knowledge, few studies have explicitly investigated the effects of partial predation on individual metabolic rates (e.g.
-`r myCite('white2011amnat')`).
+For example, the fractal geometry model `r myCite(c('west1997science', 'west1999science'))` predicts that $\alpha = n/(n + 1)$ for a $n$-dimensional organism, hence a value of $\alpha$ = 0.75 is predicted for three-dimensional animals while a value of $\alpha$ = 0.67 is predicted for two-dimensional animals (`r myCite(c('enquist2009nature', 'savage2008ploscompbiol', 'banavar2010pnas', 'kolokotrones2010nature', 'koontz2009amjbot'))`; but see `r myCite('price2007pnas')` for additional predictions of $\alpha$ upon modifications of canonical assumptions).
+Moreover, experimental size manipulation in colonial organisms allows for tests of how ecological processes such as predation or disturbance affect metabolic rates.
+For instance, it has been demonstrated that mimicking the effects of partial predation of physical disturbance through experimental size manipulation yields different responses in growth and reproductive onset in different species of encrusting and arborescent bryozoans `r myCite(c('bone2005jembe', 'hart2009ecologyb'))`.
+Yet, the metabolic consequences of these size manipulations remains unclear.
 
-Metabolic rates are also influenced by temperature, typically following an exponential function `r myCite(c('gillooly2001', 'kooijman2009book'))`.
+Metabolic rates are influenced by temperature, typically following an exponential function `r myCite(c('gillooly2001', 'kooijman2009book'))`.
 Existing metabolic theories assume nevertheless different paths through which temperature affects metabolic rates.
 For example, the fractal geometry model `r myCite(c('west1997science', 'west1999science'))` assumes that temperature affects the metabolic normalisation only `r myCite(c('gillooly2001', 'brown2004'))`, i.e.
-mass-scaling and temperature dependence of metabolic rates are fundamentally independent; contrarily, the Dynamic Energy Budget (DEB) theory for metabolic organisation `r myCite('kooijman2009book')` predicts changing relationships between mass scaling and temperature dependence through the effects of temperature on assimilation, maintenance, growth, and maturation `r myCite('kooijman2009book')`.
-Whether temperature dependence and mass scaling truly interact remains a contentious issue, though recent empirical evidence indicates that they do, at least in fishes `r myCite('killen2010ele')`.
-Testing such interaction is of great ecological significance because different mass-scaling exponents under different temperatures would potentially indicate contrasting life-history trajectories `r myCite('sebens1987annurevecolsys')` for a given species under different environments **(REFS??)**.
+mass-scaling and temperature dependence of metabolic rates are fundamentally independent.
+In contrast, the Dynamic Energy Budget (DEB) theory for metabolic organisation `r myCite('kooijman2009book')` predicts changing relationships between mass scaling and temperature dependence through the effects of temperature on assimilation, maintenance, growth, and maturation `r myCite('kooijman2009book')`.
+Whether temperature dependence and mass scaling interact remains a contentious issue, though recent empirical evidence indicates that they do, at least in fishes `r myCite('killen2010ele')`.
+Testing such interactions matters because different mass-scaling exponents under different temperatures yield different metabolic efficiencies associated with different body sizes.
+For example, a body size that maximizes scope for reproduction under one temperature may be suboptimal in a cooler temperature or vice versa `r myCite('sebens1987annurevecolsys')`.
 
@@
 To collect our study organisms, roughened acetate sheets secured to the underside of 6 mm thick PVC sheets suspended from floating pontoons at a depth of 1 m were deployed at Manly Boat Harbour, Queensland, Australia (27˚27' S, 153˚11' E) and checked regularly to remove fouling organisms other than our species of interest: *Bugula neritina* and *B.
-stolonifera* (arborescent bryozoan), *Hippopodina* sp.
-(encrusting bryozoan), and an unidentified species of sponge belonging to the Microcionidae family `r myCite('hart2009ecology')` (Table 1).
+stolonifera* (arborescent bryozoans that have a three dimensional tree-like growth form), *Hippopodina* sp.
+(an encrusting bryozoan that grows mostly in two dimensions as flat disc), and an unidentified species of sponge belonging to the Microcionidae family (`r myCite('hart2009ecology')`; again a flat species that grows mostly in two dimensions) (Table 1).
 After 6 weeks in the field, sheets bearing colonies were returned to the University of Queensland where they were maintained in aerated seawater for up to 48 h prior to measurements.
-For encrusting colonies, sections of acetate bearing whole colonies were then cut from the sheets and fragmented (see `r myCite('white2011amnat')` for details and tests of potential experimental artefacts).
+For encrusting colonies, sections of acetate bearing whole colonies were then cut from the sheets and fragmented (see e.g.
+`r myCite(c('nakaya2005procb', 'white2011amnat'))` for details and tests of potential experimental artefacts).
 Arborescent colonies were cut directly from the acetate sheets and fragmented.
@@
 *2. Respiration measurements*  
-Individual metabolic rates (**provide units**, *n* = `r nrow(metRates)`), $B_i$, were measured as rate of oxygen consumption, $\dot{V}\textrm{O}_\textrm{2}$ of colony fragments was measured at 10 or 25 ˚C using a 144-channel PreSens Sensor Dish Reader (SDR, AS-1 Scientific, Wellington, New Zealand) according to standard techniques `r myCite(c('koster2008meps', 'lighton2008book', 'white2011amnat'))`.
+Individual metabolic rates (converted to $\mu$l $\textrm{O}_\textrm{2}$ h^-1^, *n* = `r nrow(metRates)`), $B_i$, were measured as rate of oxygen consumption, $\dot{V}\textrm{O}_\textrm{2}$ of colony fragments was measured at nominal temperatures of 10 or 25 ˚C using a 144-channel PreSens Sensor Dish Reader (SDR, AS-1 Scientific, Wellington, New Zealand) according to standard techniques `r myCite(c('koster2008meps', 'lighton2008book', 'white2011amnat'))`.
 $\dot{V}\textrm{O}_\textrm{2}$ was measured by placing the colony or fragment in a glass vial containing 0.2$\mu$m filtered sea water and a non-consumptive $\textrm{O}_\textrm{2}$ sensor spot and calculated from the rate of change of $\textrm{O}_\textrm{2}$ saturation ($m_a$, \% h^‑1^) as:
@@
 
-where $m_b$ is the rate of change of $\textrm{O}_\textrm{2}$ saturation for blank vials containing no bryozoans (\% h^-1^), ${\beta}_{\textrm{O}_\textrm{2}}$ is the oxygen capacitance of air-saturated sea water at 10 ˚C (6.26 ml l^-1^) or 24 ˚C (4.615 ml l^-1^) [**double-check these numbers.
-Shouldn't the second be 25 ˚C?**] `r myCite('cameron1986book')` and $V$ is water volume (chambers were 0.005 l in volume and water volume was calculated by subtracting the volume of acetate and animals).
+where $m_b$ is the rate of change of $\textrm{O}_\textrm{2}$ saturation for blank vials containing no bryozoans (\% h^-1^), ${\beta}_{\textrm{O}_\textrm{2}}$ is the oxygen capacitance of air-saturated sea water at 10 ˚C (6.4 ml l^-1^) or 25 ˚C (4.74 ml l^-1^) `r myCite('cameron1986book')` and $V$ is water volume (chambers were 0.005 l in volume and water volume was calculated by subtracting the volume of acetate and animals).
 At least four blank vials were recorded simultaneously with each 24-well plate to account for microbial oxygen consumption and sensor spots were calibrated with air-saturated seawater (100\% A.S.) and water containing 2\% sodium sulfite (0\% A.S.).
@@
 
-Both models were fitted in a Bayesian framework by calling *JAGS* version 3.4.0 from the R package *R2jags* version 0.05-03 `r myCite('r2jags')` in order to derive posterior distributions and associated 95\% credible intervals (CIs) for the fitted parameters.
+All models were fitted in a Bayesian framework by calling *JAGS* version 3.4.0 from the R package *R2jags* version 0.05-03 `r myCite('r2jags')` in order to derive posterior distributions and associated 95\% credible intervals (CIs) for the fitted parameters.
 Random effects were assumed to be normally distributed, with means of 0.
@@
 For example, metabolic rates increase `r rounded(exp(averageEstimates$lnB25[averageEstimates$species == 'Stolonifera'])/exp(averageEstimates$lnB10[averageEstimates$species == 'Stolonifera']))`-fold between 10 ˚C and 25 ˚C for *B.
-stolonifera*, yielding a $Q_{10}$ of `r rounded(q10andErs$q10[q10andErs$species == 'Stolonifera'])`.
-It is worth noting that this $Q_{10}$ value is equivalent to an activation energy of `r rounded(q10andErs$er[q10andErs$species == 'Stolonifera'], 2)` eV (i.e.
-$E_r = \frac{k \times 283.15 \times 298.15 \times 1.5 \times \textrm{ln}Q_{10}}{(298.15 - 283.15)}$, where $E_r$ is an activation energy and $k$ is the Boltzmann constant, 8.62 $\times$ 10^-5^ eV K^-1^), which is consistent with the average 06 -- 0.7 eV reported in the Metabolic Theory of Ecology literature `r myCite(c('gillooly2001','yvondurocher2012a'))`.
-Different from the above species, the best model indicates that *Bugula neritina* and *Hippopodina* sp.
-present opposing interactions between mass-scaling and temperature dependence (Figure 1).
-For example, average posterior parameter estimates indicate that the difference in metabolic rates between 25 ˚C and 10 ˚C for *Hippopodina* sp.
+stolonifera*, yielding a $Q_{10}$ of `r rounded(q10andErs$q10[q10andErs$species == 'Stolonifera'])`, which is equivalent to an activation energy of `r rounded(q10andErs$er[q10andErs$species == 'Stolonifera'], 2)` eV `r myCite(c('gillooly2001','yvondurocher2012a'))`.
+The best model indicates that the difference in metabolic rates between 25 ˚C and 10 ˚C for *Hippopodina* sp.
 decreases, on average, `r rounded(tempDeltas[1,'tempDelta'])`-fold moving from `r round(exp(2))` to `r round(exp(5))` mg.
@@
 
-Posterior distributions for mass-scaling exponents show varied consistency with predictions from the two metabolic theories.
-For example, *Hippopodina* sp.
-shows exponents that encompass both 0.67 (Fractal geometry model prediction for 2-D organisms; `r myCite('koontz2009amjbot')`) and 0.75 (Fractal geometry model prediction for 3-D organisms; `r myCite(c('west1997science', 'west1999science'))`) at 10 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's10_97.5'], 2)`), but that are significantly shallower than 0.75 at 25 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's25_97.5'], 2)`).
-On the other hand, *Bugula neritina* presents scaling exponents that overlap both 0.67 and 0.75 at 25 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's25_97.5'], 2)`), but are significantly shallower than these values at 10 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's10_97.5'], 2)`).
+Average posterior estimates of mass-scaling exponents are inconsistent among the studied species and between growth forms (2D and 3D).
+For example, for organisms that exhibit 2D growth, *Hippopodina* sp.
+shows exponents that encompass both 0.67 (Fractal geometry model prediction for 2D organisms; `r myCite('koontz2009amjbot')`) and 0.75 (Fractal geometry model prediction for 3D organisms; `r myCite(c('west1997science', 'west1999science'))`) at 10 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's10_97.5'], 2)`), but that are significantly shallower than 0.75 at 25 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's25_97.5'], 2)`).
 Microcionidae presents exponents that only overlap 0.67 at 10 ˚C, but not at 25 ˚C (95\% CI at 10 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's10_97.5'], 2)`; 95\% CI at 25 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's25_97.5'], 2)`).
+On the other hand, for 3D organisms, *Bugula neritina* presents scaling exponents that overlap both 0.67 and 0.75 at 25 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's25_97.5'], 2)`), but are significantly shallower at 10 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's10_97.5'], 2)`).
 Finally, *Bugula stolonifera* presents values that overlap both 0.67 and 0.75 at both temperatures (95\% CI at 10 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's10_97.5'], 2)`; 95\% CI at 25 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's25_97.5'], 2)`).
@@
 
-Here, we show experimentally that the mass-scaling of metabolic rates changes between two extreme temperatures, and in different ways across different species of marine invertebrates.
+Here, we show experimentally that the mass-scaling of metabolic rates changes between two temperatures, and in different ways across different species of marine invertebrates.
 We build on recent empirical evidence documenting similar changes (e.g.
 `r myCite('killen2010ele')`), yielding further support to the need for new metabolic theories (or reformulation of current ones) that formally incorporate interactive effects between mass scaling and temperature dependence of biological rates.
-Our results, however, do not indicate consistent mass-temperature interactions across growth forms.
-Possibly, another important axis involved, which has not been considered here, is the acclimation capacity of species upon acute temperature changes (e.g.
-`r myCite('rezende2014functecol')`).
-Our findings are constrained to acute-related changes, and therefore do not consider temporal acclimation or even adaptation to changes in environmental temperature.
+Our results, however, do not indicate consistent mass-temperature interactions across growth forms -- we found idiosyncratic effects of temperature on scaling exponents across closely related species with the same growth form.
+Possibly, another important axis involved, which has not been considered here, is the acclimation capacity of species upon acute temperature changes.
+Our experimental set up, however, acclimated organisms for periods (48 h) that seem to be long enough to buffer these effects `r myCite('rezende2014functecol')`.
 
-The fact that $B_o$ and $\alpha$ covary in different ways depending on species and temperature (Figure 1) suggests that the metabolic costs of maintenance will vary for different species at different environmental temperatures.
-For example, *Bugula neritina* presented a shallower mass-scaling exponent at colder temperatures, possibly indicating that large cold-environment individuals are more efficient in spending energy than counterparts in warm environments.
-Thus, for this species, all else being equal, increases in temperature decrease the metabolic benefits of attaining a larger size while decreases in temperature reduce the metabolic normalisation.
+That metabolic rate scales at an exponent of less than 1 across all four species (with the exception of *Hippopodina* sp.
+measured at 10 ˚C) has interesting implications for our understanding of the energetics of growth for different colonial animals.
+For instance, evidence suggest that water clearance and feeding rates scale either isometrically or allometrically with colony size in sponges `r myCite(c('riisgard1993meps', 'kowalke2000jembe', 'mcmurray2014aquatbiol'))`, and these rates may follow the same mass scaling of metabolic rates (see e.g.
+`r myCite('thomassen1995meps')`).
+In contrast, most studies indicate that feeding rates scale isometrically or superlinearly in both encrusting and arborescent bryozoans (e.g.
+`r myCite(c('okamura1984jembe', 'okamura1985jembe', 'pratt2005meps'))`).
+Our results and those from other studies `r myCite('hartikainen2014jeb')` indicate that metabolic rates scale allometrically in arborescent 3D bryozoans.
+Consequently, while every unit increase in size results in proportional increase in feeding rate, every unit increase in size results in a less-than-proportional increase in metabolic rate.
+Thus, as colonies increase in size, their capacity to capture food increases more quickly than the rate at which energy is expended.
+
+This differential between ingestion rates and metabolic costs may explain the observed exponential growth and reproduction observed in some colonial marine invertebrates (e.g.
+`r myCite(c('marshall2013evolution', 'padilla1996limn', 'winston2010icb'))`) -- because the ratio of intake to expenditure increases with size, larger colonies can allocate proportionately more energy to growth and/or reproduction.
+In other words, for these colonial marine invertebrates, larger colonies are more efficient at allocating food intake to somatic growth and/or reproduction than smaller colonies.
+Isometric energy gains and allometric energy expenditure in colonial marine invertebrates may explain why they are often competitively dominant, outgrowing and excluding unitary organisms (which generally present allometric scaling of both feeding and metabolic rates, see e.g.
+`r myCite(c('kooijman2009book', 'kearney2012amnat'))`) in a range of systems `r myCite(c('buss1980pnas', 'buss1990tree'))`.
+
+While the mass scaling exponent was unaffected by temperature for two species (*Bugula stolonifera* and Microcionidae), it decreased and increased for one encrusting (*Hippopodina* sp.) and one arborescent species (*Bugula neritina*), respectively.
+These results imply that energy use may vary for the same species in different environments, possibly affecting their ultimate colonisation success and abundance.
+For example, the fact that *Bugula neritina* presented a shallower mass-scaling exponent at colder temperatures possibly indicates that large cold-environment individuals are more efficient in spending energy than counterparts in warm environments.
 On the other hand, our results indicate that *Hippopodina* sp.
 presents a reduction in metabolic benefits of attaining a larger size at colder temperatures.
-Thus, for this encrusting species, one could predict that individuals living at higher temperatures should allocate more resources to attaining larger sizes (as opposed to reproduction, for instance) while individuals living at lower temperatures will be under less pressure to allocate resources to size.
 Importantly, though, we note that there are many extrinsic factors that may affect abundance and selection on body size (e.g.
-predation, competition, etc; `r myCite('barneche2016procb')`).
+predation, competition).
 Yet, such differentials in the costs and benefits of attaining large size may help explain why some species living in cold temperatures present a slower pace of life (e.g.
 `r myCite('rosa2010jmba')`).
-Moreover, they may represent necessary improvements on current attempts to scale up energy from individuals to higher levels of organisation (e.g.
+Moreover, they might improve current attempts to scale up energy from individuals to higher levels of organisation (e.g.
 `r myCite(c('allen2005','yvondurocher2012b','barneche2014ele','barneche2016procb'))`).
 
-In the wild, partial predation is a common phenomenon experienced by colonial organisms, which must present different defence strategies `r myCite('sebens1987annurevecolsys')`.
-Our findings suggest that, for some species, predation might modulate the energetic demands of individuals of similar body mass in different environments (assuming that the size-manipulation approach herein adopted may serve as a surrogate for natural field predation; `r myCite('bone2005jembe')`), thus affecting life-history trajectories.
+In the wild, partial predation and physical disturbance are common sources of partial mortality in colonial organisms `r myCite('sebens1987annurevecolsys')`.
+Our findings suggest that, for some species, size reductions mediated by predation or disturbance might modulate the energetic demands of individuals of similar body mass in different environments, thus affecting life-history trajectories.
 For instance, at 25 ˚C, a 400-gram individual of *Bugula stolonifera* will spend `r rounded(metabolicDiff$metabolicDiff[metabolicDiff$species == 'Stolonifera'], 2)`-fold more energy than a similar-sized individual at 10 ˚C.
@@
 These differences, for instance, may help determine which species is a better competitor in different environments, as individual metabolic rates will also affect growth rates and time of reproduction onset `r myCite(c('sebens1987annurevecolsys', 'hart2009ecologyb'))`.
-It is worth noting that while we do not explicitly control for size manipulation in our study (i.e.
-evaluate the same model for non-manipulated individuals), previous evidence suggests that there is no effect of fragmentation on energetic demands of colonial organisms `r myCite('white2011amnat')`.
 
@@
 By adopting a size-manipulation approach, we achieved a more definitive test of mass-scaling of metabolic rates for cancelling the confounding effects of size and age, and by mimicking predation, which constitutes a fundamental ecological process in the wild.
-Mass-scaling exponents vary for different organisms, and are not consistent across growth forms nor phylogeny.
-These results suggest that more accurate predictions of abundance and biomass production at higher levels of organisation, from populations to ecosystems, will required a species-specific understanding of the interacting effects of mass scaling and temperature dependence, perhaps leading to improvements of current metabolic theories, or even the development of new ones.
+Mass-scaling exponents vary for different species, and are not consistent across growth forms nor phylogeny.
+Finally, our study suggests that the often observed competitive superiority of colonial organisms over unitary ones may arise because energy availability for growth and or reproduction increases more strongly with size in colonial organisms.
 
@@
 
-Kearney, people who did the plates, ARC for much money (DP0987626, DP110101776, and whatever grants Dustin wants to add).
+We thank M.
+R.
+Kearney for **what reason?**, field and laboratory assistants **names?**, the Australian Research Council for financial support (grants DP0987626, DP110101776, FT130101493 to C.
+R.
+White), and **whatever grants Dustin wants to add**.
 

### 2016-03-17 text/coverLetter.Rmd
added cover letter

Diego R. Barneche  
School of Biological Sciences  
Monash University  
Melbourne 3800 VIC  
Australia  
barnechedr@gmail.com

March XX, 2016

To the Editors of Functional Ecology,

Please find attached our manuscript entitled "Temperature effects on mass-scaling exponents in colonial animals: a manipulative test" for consideration as a Standard Paper in Functional Ecology.

In this article, we use a manipulative approach to test for an association between metabolic mass scaling and temperature dependence in four species of colonial marine invertebrates belonging to two phyla. By adopting a size-manipulation approach, we achieved a more definitive test of mass-scaling of metabolic rates for cancelling the confounding effects of size and age, and by mimicking predation, which constitutes a fundamental ecological process in the wild. Our results indicate that mass scaling and temperature dependence of metabolic rates interact in different ways across the studies species, and these interactions are not consistent across growth form nor phyla.

Our study should be of general interest to readers of Functional Ecology because our findings suggest that different species might have different competitive advantages at different environmental temperatures. We suggest that the often observed competitive superiority of some colonial organisms over unitary ones may arise due to the differential scaling between metabolic and feeding rates -- energy availability for growth and or reproduction increases more strongly with size in colonial organisms compared to unitary ones.

Below, we list some potential reviewers, contact details, and a summary of their areas of expertise. Please do not hesitate to contact me if you have any questions.

Thank you for your consideration.  
Sincerely,

Diego R. Barneche

Suggested Referees:

Dr. Douglas S. Glazier (Department of Biology, Juniata College, Huntingdon, Pennsylvania 16652 USA; E-mail: glazier@juniata.edu).  
Expertise: metabolic scaling

Dr. Shaun S. Killen (Division of Environmental & Evolutionary Biology, Faculty of Biomedical & Life Sciences, Graham Kerr Building, University of Glasgow, Glasgow G12 8QQ, UK; E-mail: s.killen@bio.gla.ac.uk).  
Expertise: metabolic scaling

Dr. Beth Okamura (Department of Life Sciences, Natural History Museum, London SW7 5BD, UK; E-mail: b.okamura@nhm.ac.uk).  
Expertise: Bryozoa life history and scaling

### 2016-03-17 paper.R

@@ -10,3 +10,10 @@ rm(list=ls())
 knit('text/MS.Rmd', output=file.path(getwd(), 'text/MS.md'), quiet=TRUE, encoding = 'utf-8')
 # to word
 system('pandoc -o text/MS.docx text/MS.md -s -S --bibliography library.bib --csl functionalEcology.csl')
+
+##############
+# COVER LETTER 
+##############
+knit('text/coverLetter.Rmd', output=file.path(getwd(), 'text/coverLetter.md'), quiet=TRUE, encoding = 'utf-8')
+# to word
+system('pandoc -o text/coverLetter.docx text/coverLetter.md')

### 2016-03-17 text/MS.Rmd
added new paragraph about population-level energy use

@@
 Metabolic theory posits that size and temperature are fundamental drivers of ecological patterns and processes, via their respective associations with metabolic rates.
-Metabolic rates are well known to increase allometrically with body mass and approximately exponentially with temperature, but most applications of metabolic theory treat these effects as independent.
+Metabolic rates tend to increase allometrically with body mass and approximately exponentially with temperature, but most applications of metabolic theory treat these effects as independent.
 2.
-Almost all studies of intraspecific scaling of unitary organisms must rely on ontogenetic differences in size to examine its relationship with metabolic rate, thereby potentially confounding age, nutrition or growth rate with size –- all of which can affect metabolic rate.
+Most studies of intraspecific scaling of unitary organisms must rely on ontogenetic differences in size to examine its relationship with metabolic rate, thereby potentially confounding age, nutrition or growth rate with size, all of which can affect metabolic rate.
 In the present study, we use a manipulative approach to test for an association between metabolic mass scaling and temperature dependence in four species of colonial marine invertebrates belonging to two phyla.
@@
 `r myCite(c('marshall2013evolution', 'padilla1996limn', 'winston2010icb'))`) -- because the ratio of intake to expenditure increases with size, larger colonies can allocate proportionately more energy to growth and/or reproduction.
-In other words, for these colonial marine invertebrates, larger colonies are more efficient at allocating food intake to somatic growth and/or reproduction than smaller colonies.
+In other words, for these colonial marine invertebrates, larger colonies are more efficient at garnering food and allocating it to somatic growth and/or reproduction than smaller colonies.
 Isometric energy gains and allometric energy expenditure in colonial marine invertebrates may explain why they are often competitively dominant, outgrowing and excluding unitary organisms (which generally present allometric scaling of both feeding and metabolic rates, see e.g.
@@
 
+At the population level, disturbance and predation can also create different size distributions.
+Our finding of allometric scaling of metabolic rate implies non-intuitive outcomes of such shifts in mean size of colonies in a population.
+For example, at 25 ˚C a population of 10 colonies of *Bugula neritina* that each weigh 10 mg would have a cumulative metabolic demand of `r rounded(10 * exp(averageEstimates$lnB25[averageEstimates$species == 'Bugula'] + log(10)*averageEstimates$s25[averageEstimates$species == 'Bugula']), 2)` $\mu$l $\textrm{O}_\textrm{2}$ h^-1^.
+If a disturbance event halved each colony with minimal loss of biomass, then a population of those (now) 20 colonies, each weighing 5 mg, would have cumulative metabolic demand of `r rounded(20 * exp(averageEstimates$lnB25[averageEstimates$species == 'Bugula'] + log(5)*averageEstimates$s25[averageEstimates$species == 'Bugula']), 2)` $\mu$l $\textrm{O}_\textrm{2}$ h^-1^ according to their new size.
+Thus, disturbance events change not only the size distribution of colonies but also the energy requirements of the population even if total standing biomass remains unchanged `r myCite('barneche2014ele')`.
+
 Here we have demonstrated that the effects of temperature interact in various ways with the mass-scaling of metabolic rates in different species of colonial marine invertebrates encompassing two distinct phyla.
-By adopting a size-manipulation approach, we achieved a more definitive test of mass-scaling of metabolic rates for cancelling the confounding effects of size and age, and by mimicking predation, which constitutes a fundamental ecological process in the wild.
+By adopting a size-manipulation approach, we achieved a more definitive test of mass-scaling of metabolic rates for cancelling the confounding effects of size and age.
 Mass-scaling exponents vary for different species, and are not consistent across growth forms nor phylogeny.
@@
 
-We thank M.
-R.
-Kearney for **what reason?**, field and laboratory assistants **names?**, the Australian Research Council for financial support (grants DP0987626, DP110101776, FT130101493 to C.
-R.
-White), and **whatever grants Dustin wants to add**.
+We thank the Australian Research Council for financial support (grants DP0987626, DP110101776, FT130101493 to C. R. White, and DP110103529 to DJM).
 

### 2016-03-18 text/MS.Rmd
first version submitted to Functional Ecology

@@
 
-Diego R. Barneche, Craig R. White, Dustin J. Marshall
+Diego R. Barneche\*, Craig R. White, Dustin J. Marshall
 
 Centre for Geometric Biology / School of Biological Sciences, Monash University, Clayton, Victoria, Australia 3800  
+\*Corresponding author: barnechedr@gmail.com
 
@@
 Our model selection approach indicated that the three-way interaction model (equation 2) performed significantly better than all nested two- and one-way interaction models (Table 2), thus implying that different species present different interactions between mass-scaling and temperature dependence.
-Particularly, the best model yields no interaction between mass scaling and temperature dependence for *Bugula stolonifera* and the Microcionidae sponge (large overlap between 95\% credible intervals for the mass-scaling slopes, $\alpha$) (Figure 1).
+Particularly, the best model yields no interaction between mass scaling and temperature dependence for *Bugula stolonifera* and the Microcionidae sponge (large overlap between 95\% credible intervals for the mass-scaling slopes, $\alpha$) (Fig.
+1).
 Still, while the mass-scaling slopes are statistically the same between temperatures for these species, there is evidence for higher metabolic normalisations (i.e.
@@
 Our finding of allometric scaling of metabolic rate implies non-intuitive outcomes of such shifts in mean size of colonies in a population.
-For example, at 25 ˚C a population of 10 colonies of *Bugula neritina* that each weigh 10 mg would have a cumulative metabolic demand of `r rounded(10 * exp(averageEstimates$lnB25[averageEstimates$species == 'Bugula'] + log(10)*averageEstimates$s25[averageEstimates$species == 'Bugula']), 2)` $\mu$l $\textrm{O}_\textrm{2}$ h^-1^.
-If a disturbance event halved each colony with minimal loss of biomass, then a population of those (now) 20 colonies, each weighing 5 mg, would have cumulative metabolic demand of `r rounded(20 * exp(averageEstimates$lnB25[averageEstimates$species == 'Bugula'] + log(5)*averageEstimates$s25[averageEstimates$species == 'Bugula']), 2)` $\mu$l $\textrm{O}_\textrm{2}$ h^-1^ according to their new size.
+For example, at 25 ˚C a population of 10 colonies of Microcionidae that each weigh 400 mg would have a cumulative metabolic demand of ~`r rounded(10 * exp(averageEstimates$lnB25[averageEstimates$species == 'Sponge'] + log(400)*averageEstimates$s25[averageEstimates$species == 'Sponge']), 0)` $\mu$l $\textrm{O}_\textrm{2}$ h^-1^.
+If a disturbance event halved each colony with minimal loss of biomass, then a population of those (now) 20 colonies, each weighing 200 mg, would have cumulative metabolic demand that is `r rounded((round(20 * exp(averageEstimates$lnB25[averageEstimates$species == 'Sponge'] + log(200)*averageEstimates$s25[averageEstimates$species == 'Sponge']))) / (round(10 * exp(averageEstimates$lnB25[averageEstimates$species == 'Sponge'] + log(400)*averageEstimates$s25[averageEstimates$species == 'Sponge']))), 1)`-fold higher (i.e.
+~`r rounded(20 * exp(averageEstimates$lnB25[averageEstimates$species == 'Sponge'] + log(200)*averageEstimates$s25[averageEstimates$species == 'Sponge']), 0)` $\mu$l $\textrm{O}_\textrm{2}$ h^-1^), according to their new size.
 Thus, disturbance events change not only the size distribution of colonies but also the energy requirements of the population even if total standing biomass remains unchanged `r myCite('barneche2014ele')`.
@@
 
-We thank the Australian Research Council for financial support (grants DP0987626, DP110101776, FT130101493 to C. R. White, and DP110103529 to DJM).
+We thank the Australian Research Council for financial support (grants DP0987626, DP110101776, FT130101493 to C.
+R.
+White, and DP110103529 to D.
+J.
+Marshall).
+The authors declare no conflict of interest.
+
+**Data Accessibility**
+
+The dataset associated with this manuscript will be archived on DRYAD should the manuscript be accepted.
+All data and code necessary to reproduce this paper, its analyses, tables and figures can also be obtained on GitHub https://github.com/dbarneche/MTRBrEs.
 

### 2016-03-18 text/coverLetter.Rmd
first version submitted to Functional Ecology

@@
-Diego R. Barneche  
-School of Biological Sciences  
-Monash University  
-Melbourne 3800 VIC  
-Australia  
-barnechedr@gmail.com
-
-March XX, 2016
+March 18 2016
 
@@
 In this article, we use a manipulative approach to test for an association between metabolic mass scaling and temperature dependence in four species of colonial marine invertebrates belonging to two phyla.
-By adopting a size-manipulation approach, we achieved a more definitive test of mass-scaling of metabolic rates for cancelling the confounding effects of size and age, and by mimicking predation, which constitutes a fundamental ecological process in the wild.
+By adopting a size-manipulation approach, we achieved a more definitive test of mass-scaling of metabolic rates for cancelling the confounding effects of size and age, and by fragmenting colonies, a natural source of size variation in modular organisms in the wild.
 Our results indicate that mass scaling and temperature dependence of metabolic rates interact in different ways across the studies species, and these interactions are not consistent across growth form nor phyla.
@@
 Diego R. Barneche  
+(on behalf of all coauthors)  
+School of Biological Sciences  
+Monash University  
+Melbourne 3800 VIC  
+Australia  
+barnechedr@gmail.com
 

### 2016-05-19 text/MS.Rmd
MS v10

@@
 1.
-Metabolic theory posits that size and temperature are fundamental drivers of ecological patterns and processes, via their respective associations with metabolic rates.
-Metabolic rates tend to increase allometrically with body mass and approximately exponentially with temperature, but most applications of metabolic theory treat these effects as independent.
+Metabolic theories posit that body size and temperature are fundamental drivers of ecological patterns and processes, from population to ecosystem levels, because they determine metabolic rates at the individual level.
+Individual-level metabolic rates tend to scale sub-linearly with body mass and approximately exponentially with temperature, but whether these effects act independently remains a contentious issue.
 2.
 Most studies of intraspecific scaling of unitary organisms must rely on ontogenetic differences in size to examine its relationship with metabolic rate, thereby potentially confounding age, nutrition or growth rate with size, all of which can affect metabolic rate.
-In the present study, we use a manipulative approach to test for an association between metabolic mass scaling and temperature dependence in four species of colonial marine invertebrates belonging to two phyla.
+In the present study, we use a manipulative approach to test for an association (or lack thereof) between metabolic mass scaling and temperature dependence in four species of colonial marine invertebrates belonging to two phyla.
+Experimental size manipulation in colonial organisms is of particular relevance to test how ecological processes such as predation and disturbance affect individual physiology and consequently energy flux at population and community levels.
 3.
-While increased temperature increased metabolic rate, we found no consistent effect of temperature or growth form on metabolic scaling; rather the effects were species-specific.
-We show that mass-scaling exponents and temperature dependencies interact in ways that are unanticipated by extant theory.
+While increased temperature increased metabolic rate in the four species analysed, we found no consistent effect of temperature or growth form on the scaling of metabolic rate with body mass.
+Two species exhibited independent effects of mass and temperature on metabolic rate, whereas the other two species exhibited decreased and increased mass-scaling exponents with increasing temperature.
+Importantly, three out of four species consistently presented sub-linear allometric scaling of metabolic rates regardless of experimental temperature.
+We explore the ecological consequences of how these different metabolic scaling potentially affect energy fluxes at population and community levels.
 4.
 The allometric scaling of metabolic rate that we observe in the present study contrasts with the isometric scaling of water clearance and feeding rates observed in some colonial marine invertebrates.
-Our findings suggest that the often observed competitive superiority of colonial organisms over unitary ones may arise because energy availability for growth and or reproduction (the difference between energy intake and expenditure) increases more strongly with size in colonial organisms.
+Our findings suggest that the often observed competitive superiority of colonial over unitary organisms may arise because energy availability for growth and or reproduction (the difference between energy intake and expenditure) increases more strongly with size in colonial organisms.
 
@@
 
-Metabolic theories seek to unite hierarchical levels of biological organisation, from individuals to ecosystems `r myCite(c('nisbet2000janimecol', 'brown2004', 'vandermeer2006tree'))`.
-Although definitive tests of their underlying assumptions remain rare `r myCite(c('kearney2012amnat', 'newberry2015ploscomputbiol'))`, myriad studies examined their predictions with regards to the relationship between individual level metabolic rates, $B_i$, and individual body mass, $M_i$ `r myCite(c('kleiber1932', 'kleiber1961', 'peters1983', 'schmidtnielsen1984', 'glazier2005biolrev', 'delong2010', 'glazier2010biolrev'))`.
-The relationship between $B_i$ and $M_i$ is typically described by a power function of the form $B_i = B_oM_i^{\alpha}$, where $B_o$ is a normalisation constant independent of mass and $\alpha$ is a dimensionless scaling exponent.
-Numerous studies have demonstrated that, for most animals, individual metabolic rate scales sub-linearly with body mass, with $\alpha$ estimates varying with many biotic and abiotic variables (e.g.
-`r myCite(c('lovegrove2000amnat', 'white2009evolution', 'killen2010ele', 'kolokotrones2010nature', 'vaca2010biollett', 'glazier2011ecolmongr', 'ketola2012jlinsoc', 'white2012procb', 'white2013jcompphysb', 'barneche2014ele'))`), generally falling between 0.5 and 1.
-While these studies demonstrate clearly that the value of $\alpha$ is not fixed, they each share a common limitation: each study tests for an association between $\alpha$ and a biotic or abiotic variable, using a value of $\alpha$ that is estimated from the association between $B_i$ and $M_i$ for individuals or species of different size.
-Such correlative studies cannot address the causal association between $B_i$ and $M_i$, because $M_i$ is correlated with a range of other environmental, physiological, and life history variables (see e.g.
+Metabolic theories seek to unite hierarchical levels of biological organisation based on the scaling of energetic demands (i.e.
+metabolic rates), from individuals to ecosystems `r myCite(c('nisbet2000janimecol', 'brown2004', 'vandermeer2006tree'))`.
+The role of individual body mass, $M_i$, as a determinant of individual-level metabolic rate, $B_i$, is key to this goal.
+Myriad studies have examined the relationship between $B_i$ and $M_i$ `r myCite(c('kleiber1932', 'kleiber1961', 'peters1983', 'schmidtnielsen1984', 'glazier2005biolrev', 'delong2010', 'glazier2010biolrev'))`, typically revealing a power function of the form $B_i = B_oM_i^{\alpha}$, where $B_o$ is a normalisation constant independent of mass and $\alpha$ is a dimensionless scaling exponent.
+For most animals, $B_i$ scales sub-linearly (or negatively allometric, sensu `r myCite('huxley1932')`) with $M_i$, i.e.
+with $\alpha$ < 1, generally falling between 0.5 and 1.
+Importantly, $\alpha$ estimates can also vary considerably among taxa and with many biotic and abiotic variables (e.g.
+`r myCite(c('lovegrove2000amnat', 'moses2008amnat', 'white2009evolution', 'delong2010', 'killen2010ele', 'kolokotrones2010nature', 'vaca2010biollett', 'glazier2011ecolmongr', 'ketola2012jlinsoc', 'white2012procb', 'white2013jcompphysb', 'barneche2014ele', 'hirst2014ele'))`).
+While these studies demonstrate clearly that the value of $\alpha$ is not fixed, most share a common limitation: they rely on pre-existing differences in mass.
+Such correlative studies cannot address the causal association between $B_i$ and $M_i$ properly, because $M_i$ is correlated with a range of other environmental, physiological, and life history variables (see e.g.
 `r myCite(c('peters1983', 'calder1984book', 'schmidtnielsen1984'))` for comprehensive reviews) and many of these variables are also associated with $B_i$ `r myCite(c('mcnab2002book', 'konarzewski2013jcompphysb', 'white2013jcompphysb'))`.
 
-In contrast to studies that make use of intact individual organisms, colonial animals provide an opportunity for more definitive tests of metabolic theories.
-Colony size can be manipulated experimentally, hence allowing for the potential to remove the confounding effects of body size and life-history traits (e.g.
-`r myCite('white2011amnat')`).
+In contrast to studies that make use of intact individual organisms, colonial animals provide an opportunity for more definitive tests of the association between $B_i$ and $M_i$.
+Colony size can be manipulated experimentally, hence allowing for the potential to remove the confounding effects of life-history traits that are also correlated with body size (e.g.
+`r myCite(c('pratt2005meps', 'hart2009ecologyb', 'white2011amnat'))`).
 Furthermore, different colonial species may grow in different dimensions, and the mass-scaling exponent $\alpha$ is explicitly predicted to change with number of body dimensions `r myCite(c('nakaya2005procb', 'white2011amnat', 'kearney2012amnat'))`.
 For example, the fractal geometry model `r myCite(c('west1997science', 'west1999science'))` predicts that $\alpha = n/(n + 1)$ for a $n$-dimensional organism, hence a value of $\alpha$ = 0.75 is predicted for three-dimensional animals while a value of $\alpha$ = 0.67 is predicted for two-dimensional animals (`r myCite(c('enquist2009nature', 'savage2008ploscompbiol', 'banavar2010pnas', 'kolokotrones2010nature', 'koontz2009amjbot'))`; but see `r myCite('price2007pnas')` for additional predictions of $\alpha$ upon modifications of canonical assumptions).
-Moreover, experimental size manipulation in colonial organisms allows for tests of how ecological processes such as predation or disturbance affect metabolic rates.
-For instance, it has been demonstrated that mimicking the effects of partial predation of physical disturbance through experimental size manipulation yields different responses in growth and reproductive onset in different species of encrusting and arborescent bryozoans `r myCite(c('bone2005jembe', 'hart2009ecologyb'))`.
-Yet, the metabolic consequences of these size manipulations remains unclear.
+Currently, however, it is uncertain how colonial organisms may conform to these predictions because the original fractal geometry model considers resource-distribution networks within unitary organisms (or individual zooids) and not colonies that might (or might not) distribute resources in fundamentally different ways.
+Assessing the scaling of metabolic rate with body mass in colonial organisms is therefore a fundamental step in assessing how this mode of life conforms to general predictions, as well as providing an excellent opportunity for empirically manipulating mass.
 
-Metabolic rates are influenced by temperature, typically following an exponential function `r myCite(c('gillooly2001', 'kooijman2009book'))`.
-Existing metabolic theories assume nevertheless different paths through which temperature affects metabolic rates.
-For example, the fractal geometry model `r myCite(c('west1997science', 'west1999science'))` assumes that temperature affects the metabolic normalisation only `r myCite(c('gillooly2001', 'brown2004'))`, i.e.
-mass-scaling and temperature dependence of metabolic rates are fundamentally independent.
-In contrast, the Dynamic Energy Budget (DEB) theory for metabolic organisation `r myCite('kooijman2009book')` predicts changing relationships between mass scaling and temperature dependence through the effects of temperature on assimilation, maintenance, growth, and maturation `r myCite('kooijman2009book')`.
-Whether temperature dependence and mass scaling interact remains a contentious issue, though recent empirical evidence indicates that they do, at least in fishes `r myCite('killen2010ele')`.
-Testing such interactions matters because different mass-scaling exponents under different temperatures yield different metabolic efficiencies associated with different body sizes.
+Experimental size manipulation in colonial organisms also allows for tests of how ecological processes such as predation or disturbance affect individual physiology.
+For instance, it has been demonstrated that mimicking the effects of partial predation or physical disturbance through experimental size manipulation yields different responses in growth and reproductive onset in one arborescent (no effects on growth, delayed reproduction and reduced fecundity; `r myCite('bone2005jembe')`) and one encrusting bryozoan (reduced growth, with reproductive onset and fecundity dependent on the age of remaining zooids; `r myCite('hart2009ecologyb')`).
+On the other hand, size manipulations on one encrusting bryozoan `r myCite('white2011amnat')` and one colonial ascidian `r myCite('nakaya2005procb')` only affected the size of individuals, such that the scaling relationships between metabolic rate and colony size did not differ between size-manipulated and unmanipulated colonies.
+Sub-linear allometric scaling, in turn, implies that mass-specific metabolic rates decrease with size, and therefore changes in size structure will influence total population- and community- level respiration flux (if total standing biomass is held fixed; `r myCite(c('allen2005', 'barneche2014ele'))`).
+Meanwhile, resource ingestion rates appear to scale linearly with size, in some colonial organisms at least (e.g.
+`r myCite('pratt2005meps')`), suggesting that these two key processes -- energy ingestion and expenditure -- scale very differently with colony size.
+These different scaling relationships suggest that larger colonies have an energetic advantage over smaller colonies, although this remains to be tested.
+
+Metabolic rates are also influenced by temperature, typically following an exponential function `r myCite(c('gillooly2001', 'kooijman2009book'))`.
+Whether the temperature dependence and mass scaling of metabolic rates interact remains a contentious issue, though recent empirical evidence indicates that they do, and in various ways (e.g.
+`r myCite(c('killen2010ele', 'glazier2005biolrev', 'glazier2010biolrev', 'price2012ele', 'glazier2014systems'))`).
+Testing for interactions between mass and temperature on metabolic rates matters because different mass-scaling exponents under different temperatures yield different metabolic efficiencies associated with different body sizes.
 For example, a body size that maximizes scope for reproduction under one temperature may be suboptimal in a cooler temperature or vice versa `r myCite('sebens1987annurevecolsys')`.
-
 In the present study we experimentally manipulate size in different colonial marine organisms from two phyla (Bryozoa [Ectoprocta] and Porifera) in order to test whether the mass scaling and temperature dependence of metabolic rates interact `r myCite(c('white2011amnat', 'kearney2012amnat'))`.
-In doing so, we explicitly test if these interactions are general or species specific using a Bayesian model selection procedure.
-Finally, we discuss the relevance of our results to the life history of different species living under different environments.
+In doing so, we explicitly test if these interactions are general or species specific using a model selection procedure.
 
@@
 To collect our study organisms, roughened acetate sheets secured to the underside of 6 mm thick PVC sheets suspended from floating pontoons at a depth of 1 m were deployed at Manly Boat Harbour, Queensland, Australia (27˚27' S, 153˚11' E) and checked regularly to remove fouling organisms other than our species of interest: *Bugula neritina* and *B.
-stolonifera* (arborescent bryozoans that have a three dimensional tree-like growth form), *Hippopodina* sp.
-(an encrusting bryozoan that grows mostly in two dimensions as flat disc), and an unidentified species of sponge belonging to the Microcionidae family (`r myCite('hart2009ecology')`; again a flat species that grows mostly in two dimensions) (Table 1).
-After 6 weeks in the field, sheets bearing colonies were returned to the University of Queensland where they were maintained in aerated seawater for up to 48 h prior to measurements.
+stolonifera* (arborescent bryozoans that have a three dimensional tree-like growth form), *Hippopodina iririkiensis* (an encrusting bryozoan that grows mostly in two dimensions as flat disc), and an unidentified species of sponge belonging to the Microcionidae family (`r myCite('hart2009ecology')`; again a flat species that grows mostly in two dimensions) (Table 1).
+After 6 weeks in the field (average SST 22 ˚C), sheets bearing colonies were returned to the University of Queensland where they were kept in aerated seawater tanks (18 ˚C) for up to 48 h prior to measurements.
 For encrusting colonies, sections of acetate bearing whole colonies were then cut from the sheets and fragmented (see e.g.
@@
 *2. Respiration measurements*  
-Individual metabolic rates (converted to $\mu$l $\textrm{O}_\textrm{2}$ h^-1^, *n* = `r nrow(metRates)`), $B_i$, were measured as rate of oxygen consumption, $\dot{V}\textrm{O}_\textrm{2}$ of colony fragments was measured at nominal temperatures of 10 or 25 ˚C using a 144-channel PreSens Sensor Dish Reader (SDR, AS-1 Scientific, Wellington, New Zealand) according to standard techniques `r myCite(c('koster2008meps', 'lighton2008book', 'white2011amnat'))`.
+Individual metabolic rates (converted to $\mu$l $\textrm{O}_\textrm{2}$ h^-1^, *n* = `r nrow(metRates)`), $B_i$, were measured as rate of oxygen consumption, $\dot{V}\textrm{O}_\textrm{2}$, of colony fragments at nominal temperatures of 10 or 25 ˚C.
+All individuals (regardless of size) and species were randomly assigned to test temperatures, and we used replicate vials and controlled temperature cabinets (see below).
+Measurements were conducted using a 144-channel PreSens Sensor Dish Reader (SDR, AS-1 Scientific, Wellington, New Zealand) according to standard techniques `r myCite(c('koster2008meps', 'lighton2008book', 'white2011amnat'))`.
 $\dot{V}\textrm{O}_\textrm{2}$ was measured by placing the colony or fragment in a glass vial containing 0.2$\mu$m filtered sea water and a non-consumptive $\textrm{O}_\textrm{2}$ sensor spot and calculated from the rate of change of $\textrm{O}_\textrm{2}$ saturation ($m_a$, \% h^‑1^) as:
@@
 
-where $m_b$ is the rate of change of $\textrm{O}_\textrm{2}$ saturation for blank vials containing no bryozoans (\% h^-1^), ${\beta}_{\textrm{O}_\textrm{2}}$ is the oxygen capacitance of air-saturated sea water at 10 ˚C (6.4 ml l^-1^) or 25 ˚C (4.74 ml l^-1^) `r myCite('cameron1986book')` and $V$ is water volume (chambers were 0.005 l in volume and water volume was calculated by subtracting the volume of acetate and animals).
+where $m_b$ is the rate of change of $\textrm{O}_\textrm{2}$ saturation for blank vials containing no bryozoans (\% h^-1^), ${\beta}_{\textrm{O}_\textrm{2}}$ is the oxygen capacitance of air-saturated sea water at 10 ˚C (6.4 ml l^-1^) or 25 ˚C (4.74 ml l^-1^) `r myCite('cameron1986book')` and $V$ is water volume (vials were 0.005 l in volume and water volume was calculated by subtracting the volume of acetate and animals).
 At least four blank vials were recorded simultaneously with each 24-well plate to account for microbial oxygen consumption and sensor spots were calibrated with air-saturated seawater (100\% A.S.) and water containing 2\% sodium sulfite (0\% A.S.).
+After a 1 h setting period (i.e.
+time necessary for the system to come into equilibrium), $\dot{V}\textrm{O}_\textrm{2}$ measurements lasted up to 20 h depending on individual mass (minute individuals require longer periods of time in order to detect $\dot{V}\textrm{O}_\textrm{2}$ signals).
 Following measurement of $\dot{V}\textrm{O}_\textrm{2}$, colonies or fragments were blotted dry and weighed to 0.1 mg (Sartorius A 200 S, Sartorius AG, Göttingen, Germany).
@@
 We also included date of experiment, $\Delta D$ as a random effect on the metabolic normalisation to help control for natural experimental variability.
-We then performed a Bayesian model selection procedure (see below) by comparing the above full model with a series of nested two- and one-way interaction models in order to formally test for mass-temperature interactions in metabolic rates among different species.
+We then performed a model selection procedure (see below) by comparing the above full model with a series of nested two- and one-way interaction models in order to formally test for mass-temperature interactions in metabolic rates among different species.
 
-All models were fitted in a Bayesian framework by calling *JAGS* version 3.4.0 from the R package *R2jags* version 0.05-03 `r myCite('r2jags')` in order to derive posterior distributions and associated 95\% credible intervals (CIs) for the fitted parameters.
+All models were initially fitted using maximum likelihood `r myCite('zuur2009book')` with the R package *lme4* version 1.1-8 `r myCite('lme4')`.
+Each nested model was compared against the full model using likelihood ratio tests, and models were considered significantly different if *p* < 0.05.
+We then fitted the best model in a Bayesian framework by calling *JAGS* version 3.4.0 from the R package *R2jags* version 0.05-03 `r myCite('r2jags')` in order to derive posterior distributions and associated 95\% credible intervals (CIs) for the fitted parameters.
 Random effects were assumed to be normally distributed, with means of 0.
@@
 locally uniform over the region supported by the likelihood) `r myCite('kruschke2014book')`.
-The posterior distributions of model parameters were estimated using Markov chain Monte Carlo (MCMC) methods by constructing five chains of 1,000,000 steps each, including 500,000-step burn-in periods.
-Chains were thinned using a 250-step interval, so a total of 10,000 steps were retained to estimate posterior distributions (i.e.
-5 $\times$ (1,000,000 - 500,000)/250 = 10,000).
-Each nested model was compared against the full model using leave-one-out cross-validation (LOO), which, similar to widely applicable information criterion (WAIC), is a fully Bayesian model selection procedure for estimating pointwise out-of-sample prediction accuracy using the log-likelihood evaluated at the posterior simulations of the parameter values `r myCite('hooten2015ecology')`.
-Model significance was assessed by calculating a *z*-score from the average difference $\pm$ SE between expected log pointwise predictive density of both models, and models were considered significantly different at $\alpha \le$ 0.05.
-Model comparison was implemented using the R package *loo* version 0.1.3 `r myCite('loo')`.
+The posterior distributions of model parameters were estimated using Markov chain Monte Carlo (MCMC) methods by constructing three chains of 500,000 steps each, including 250,000-step burn-in periods.
+Chains were then updated with 500,000 steps each until convergence was reached using 250-thinning step interval, so a total of 6,000 steps were retained to estimate posterior distributions (i.e.
+3 $\times$ 500,000/250 = 6,000).
 All data and code necessary to reproduce this paper, its analyses, tables and figures can be obtained on GitHub https://github.com/dbarneche/MTRBrEs.
@@
 
-Our model selection approach indicated that the three-way interaction model (equation 2) performed significantly better than all nested two- and one-way interaction models (Table 2), thus implying that different species present different interactions between mass-scaling and temperature dependence.
+The model selection indicated that the three-way interaction model (equation 2) performed significantly better than all nested two- and one-way interaction models (Table 2), thus implying that different species present different interactions between mass-scaling and temperature dependence.
 Particularly, the best model yields no interaction between mass scaling and temperature dependence for *Bugula stolonifera* and the Microcionidae sponge (large overlap between 95\% credible intervals for the mass-scaling slopes, $\alpha$) (Fig.
@@
 Still, while the mass-scaling slopes are statistically the same between temperatures for these species, there is evidence for higher metabolic normalisations (i.e.
-intercepts), $B_o$, at higher temperatures (though not statistically significant for Microcionidae).
+intercepts), $B_o$, at higher temperatures (though marginally significant for Microcionidae, i.e.
+slight overlap between 95\% credible intervals).
 For example, metabolic rates increase `r rounded(exp(averageEstimates$lnB25[averageEstimates$species == 'Stolonifera'])/exp(averageEstimates$lnB10[averageEstimates$species == 'Stolonifera']))`-fold between 10 ˚C and 25 ˚C for *B.
 stolonifera*, yielding a $Q_{10}$ of `r rounded(q10andErs$q10[q10andErs$species == 'Stolonifera'])`, which is equivalent to an activation energy of `r rounded(q10andErs$er[q10andErs$species == 'Stolonifera'], 2)` eV `r myCite(c('gillooly2001','yvondurocher2012a'))`.
-The best model indicates that the difference in metabolic rates between 25 ˚C and 10 ˚C for *Hippopodina* sp.
-decreases, on average, `r rounded(tempDeltas[1,'tempDelta'])`-fold moving from `r round(exp(2))` to `r round(exp(5))` mg.
+The best model indicates that the difference in metabolic rates between 25 ˚C and 10 ˚C for *Hippopodina iririkiensis* decreases, on average, `r rounded(tempDeltas[1,'tempDelta'])`-fold moving from `r round(exp(2))` to `r round(exp(5))` mg.
 For *Bugula neritina*, however, rates increase `r rounded(1/tempDeltas[3,'tempDelta'])`-fold along the same body-mass gradient.
@@
 Average posterior estimates of mass-scaling exponents are inconsistent among the studied species and between growth forms (2D and 3D).
-For example, for organisms that exhibit 2D growth, *Hippopodina* sp.
-shows exponents that encompass both 0.67 (Fractal geometry model prediction for 2D organisms; `r myCite('koontz2009amjbot')`) and 0.75 (Fractal geometry model prediction for 3D organisms; `r myCite(c('west1997science', 'west1999science'))`) at 10 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's10_97.5'], 2)`), but that are significantly shallower than 0.75 at 25 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's25_97.5'], 2)`).
+For example, for organisms that exhibit 2D growth, *Hippopodina iririkiensis* shows exponents that encompass both 0.67 (Fractal geometry model prediction for 2D organisms; `r myCite('koontz2009amjbot')`) and 0.75 (Fractal geometry model prediction for 3D organisms; `r myCite(c('west1997science', 'west1999science'))`) at 10 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's10_97.5'], 2)`), but that are significantly shallower than 0.75 at 25 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's25_97.5'], 2)`).
 Microcionidae presents exponents that only overlap 0.67 at 10 ˚C, but not at 25 ˚C (95\% CI at 10 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's10_97.5'], 2)`; 95\% CI at 25 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's25_97.5'], 2)`).
@@
 
-Here, we show experimentally that the mass-scaling of metabolic rates changes between two temperatures, and in different ways across different species of marine invertebrates.
-We build on recent empirical evidence documenting similar changes (e.g.
-`r myCite('killen2010ele')`), yielding further support to the need for new metabolic theories (or reformulation of current ones) that formally incorporate interactive effects between mass scaling and temperature dependence of biological rates.
-Our results, however, do not indicate consistent mass-temperature interactions across growth forms -- we found idiosyncratic effects of temperature on scaling exponents across closely related species with the same growth form.
-Possibly, another important axis involved, which has not been considered here, is the acclimation capacity of species upon acute temperature changes.
-Our experimental set up, however, acclimated organisms for periods (48 h) that seem to be long enough to buffer these effects `r myCite('rezende2014functecol')`.
+Here, we show experimentally that the mass-scaling of metabolic rates may change with temperature, and in different ways across different species of marine invertebrates and between growth forms.
+We thus add to recent empirical evidence documenting mass-temperature interactions in metabolic rates (e.g.
+`r myCite(c('killen2010ele', 'glazier2005biolrev'))`).
+Possibly, the ability of our studied species to acclimate upon acute temperature changes could have affected the results (see discussion in `r myCite('glazier2014systems')`).
+Our experimental set up, however, randomised species and individual sizes across temperatures such that systematic effects across sizes or among species are unlikely.
+Moreover, the temperature ranges span those experienced by these species in the field.
+Still, if mass-temperature interactions are proven to be frequent and consistent across different types of organisms, future studies will be necessary to mechanistically explain the interactive effects between mass scaling and temperature dependence of biological rates (but see the metabolic-level boundaries hypothesis in `r myCite(c('glazier2005biolrev', 'glazier2010biolrev', 'glazier2014systems'))`).
 
-That metabolic rate scales at an exponent of less than 1 across all four species (with the exception of *Hippopodina* sp.
-measured at 10 ˚C) has interesting implications for our understanding of the energetics of growth for different colonial animals.
+That metabolic rate scales at an exponent of less than 1 across all four species (with the exception of *Hippopodina iririkiensis* measured at 10 ˚C) has interesting implications for our understanding of the energetics of growth for different colonial animals.
 For instance, evidence suggest that water clearance and feeding rates scale either isometrically or allometrically with colony size in sponges `r myCite(c('riisgard1993meps', 'kowalke2000jembe', 'mcmurray2014aquatbiol'))`, and these rates may follow the same mass scaling of metabolic rates (see e.g.
@@
 Consequently, while every unit increase in size results in proportional increase in feeding rate, every unit increase in size results in a less-than-proportional increase in metabolic rate.
-Thus, as colonies increase in size, their capacity to capture food increases more quickly than the rate at which energy is expended.
+Thus, as colonies increase in size, their capacity to capture food increases more quickly than the rate at which they expend energy.
 
-This differential between ingestion rates and metabolic costs may explain the observed exponential growth and reproduction observed in some colonial marine invertebrates (e.g.
+The difference between the mass-scaling of ingestion rates and metabolic rates may explain the observed exponential growth and reproduction observed in some colonial marine invertebrates (e.g.
 `r myCite(c('marshall2013evolution', 'padilla1996limn', 'winston2010icb'))`) -- because the ratio of intake to expenditure increases with size, larger colonies can allocate proportionately more energy to growth and/or reproduction.
@@
 
-While the mass scaling exponent was unaffected by temperature for two species (*Bugula stolonifera* and Microcionidae), it decreased and increased for one encrusting (*Hippopodina* sp.) and one arborescent species (*Bugula neritina*), respectively.
+While the mass scaling exponent was unaffected by temperature for two species (*Bugula stolonifera* and Microcionidae), it decreased and increased for one encrusting (*Hippopodina iririkiensis*) and one arborescent species (*Bugula neritina*), respectively.
 These results imply that energy use may vary for the same species in different environments, possibly affecting their ultimate colonisation success and abundance.
 For example, the fact that *Bugula neritina* presented a shallower mass-scaling exponent at colder temperatures possibly indicates that large cold-environment individuals are more efficient in spending energy than counterparts in warm environments.
-On the other hand, our results indicate that *Hippopodina* sp.
-presents a reduction in metabolic benefits of attaining a larger size at colder temperatures.
+On the other hand, our results indicate that *Hippopodina iririkiensis* presents a reduction in metabolic benefits of attaining a larger size at colder temperatures.
 Importantly, though, we note that there are many extrinsic factors that may affect abundance and selection on body size (e.g.
 predation, competition).
-Yet, such differentials in the costs and benefits of attaining large size may help explain why some species living in cold temperatures present a slower pace of life (e.g.
+Nonetheless, such differentials in the costs and benefits of attaining large size may help explain why some species living in cold temperatures present a slower pace of life (e.g.
 `r myCite('rosa2010jmba')`).
-Moreover, they might improve current attempts to scale up energy from individuals to higher levels of organisation (e.g.
+Moreover, incorporation of differences in the size-dependence of the costs and benefits of size might improve current attempts to scale up energy from individuals to higher levels of organisation (e.g.
 `r myCite(c('allen2005','yvondurocher2012b','barneche2014ele','barneche2016procb'))`).
@@
 Our findings suggest that, for some species, size reductions mediated by predation or disturbance might modulate the energetic demands of individuals of similar body mass in different environments, thus affecting life-history trajectories.
-For instance, at 25 ˚C, a 400-gram individual of *Bugula stolonifera* will spend `r rounded(metabolicDiff$metabolicDiff[metabolicDiff$species == 'Stolonifera'], 2)`-fold more energy than a similar-sized individual at 10 ˚C.
+For instance, at 25 ˚C, a 400-gram individual of *Bugula stolonifera* will spend ~`r round(metabolicDiff$metabolicDiff[metabolicDiff$species == 'Stolonifera'])`-fold more energy than a similar-sized individual at 10 ˚C.
 Conversely, due to the estimated positive interaction between mass-scaling and temperature, a similar-sized individual of *B.
-neritina* will spend `r rounded(metabolicDiff$metabolicDiff[metabolicDiff$species == 'Bugula'], 2)`-fold more energy at 25 ˚C than at 10 ˚C.
+neritina* will spend ~`r round(metabolicDiff$metabolicDiff[metabolicDiff$species == 'Bugula'])`-fold more energy at 25 ˚C than at 10 ˚C.
 These differences, for instance, may help determine which species is a better competitor in different environments, as individual metabolic rates will also affect growth rates and time of reproduction onset `r myCite(c('sebens1987annurevecolsys', 'hart2009ecologyb'))`.
 
-At the population level, disturbance and predation can also create different size distributions.
+At the population level, disturbance and predation can also create different size distributions in colonial organisms.
 Our finding of allometric scaling of metabolic rate implies non-intuitive outcomes of such shifts in mean size of colonies in a population.
@@
 
-Here we have demonstrated that the effects of temperature interact in various ways with the mass-scaling of metabolic rates in different species of colonial marine invertebrates encompassing two distinct phyla.
-By adopting a size-manipulation approach, we achieved a more definitive test of mass-scaling of metabolic rates for cancelling the confounding effects of size and age.
+Here we have demonstrated that the effects of temperature can interact in different ways with the mass-scaling of metabolic rates in different species of colonial marine invertebrates encompassing two distinct phyla.
+By adopting a size-manipulation approach, we achieved a more definitive test of mass-scaling of metabolic rates by eliminating confounding effects of age and other functional traits that could also be correlated with size and/or metabolic rates.
 Mass-scaling exponents vary for different species, and are not consistent across growth forms nor phylogeny.
@@
 
-*Table 2.* The best candidate model was chosen based on a selection procedure with nested models where non-significant fixed effects were successively removed based on leave-one-out cross-validation.
-Model significance was assessed by calculating a *z*-score from the average difference $\pm$ SE between expected log pointwise predictive density of both models ($\widehat{elpd_{diff}}$).
+**Table 2.** The best model was constructed, using the R package *lme4*, by successively removing fixed effects based on likelihood ratio tests of significance `r myCite('zuur2009book')`.
+In the table, $\chi^2$ and d.f.
+refer to likelihood ratio test between the full model and nested model.
+The final best model, which includes all parameters, is indicated in bold.
 

### 2016-05-19 R/functions-figures.R
updated hippopodina status

@@ -193,7 +193,7 @@ fig1  <-  function() {
   colnames(tmat)  <-  names(coef(lmerModel1)$Run)
   species  <-  unique(metRates$Species)
   # species real names
-  spNames  <-  list(substitute(italic('Hippopodina')*' sp.'),
+  spNames  <-  list(substitute(italic('Hippopodina iririkiensis')),
                     'Microcionidae',
                     substitute(italic('Bugula neritina')),
                     substitute(italic('Bugula stolonifera')))

### 2016-05-19 text/MS.Rmd
changed abstract follwogin Ecology guidelines

@@
 
-1.
-Metabolic theories posit that body size and temperature are fundamental drivers of ecological patterns and processes, from population to ecosystem levels, because they determine metabolic rates at the individual level.
-Individual-level metabolic rates tend to scale sub-linearly with body mass and approximately exponentially with temperature, but whether these effects act independently remains a contentious issue.
-2.
-Most studies of intraspecific scaling of unitary organisms must rely on ontogenetic differences in size to examine its relationship with metabolic rate, thereby potentially confounding age, nutrition or growth rate with size, all of which can affect metabolic rate.
-In the present study, we use a manipulative approach to test for an association (or lack thereof) between metabolic mass scaling and temperature dependence in four species of colonial marine invertebrates belonging to two phyla.
-Experimental size manipulation in colonial organisms is of particular relevance to test how ecological processes such as predation and disturbance affect individual physiology and consequently energy flux at population and community levels.
-3.
-While increased temperature increased metabolic rate in the four species analysed, we found no consistent effect of temperature or growth form on the scaling of metabolic rate with body mass.
-Two species exhibited independent effects of mass and temperature on metabolic rate, whereas the other two species exhibited decreased and increased mass-scaling exponents with increasing temperature.
-Importantly, three out of four species consistently presented sub-linear allometric scaling of metabolic rates regardless of experimental temperature.
-We explore the ecological consequences of how these different metabolic scaling potentially affect energy fluxes at population and community levels.
-4.
-The allometric scaling of metabolic rate that we observe in the present study contrasts with the isometric scaling of water clearance and feeding rates observed in some colonial marine invertebrates.
-Our findings suggest that the often observed competitive superiority of colonial over unitary organisms may arise because energy availability for growth and or reproduction (the difference between energy intake and expenditure) increases more strongly with size in colonial organisms.
+Body size and temperature are fundamental drivers of ecological processes because they determine metabolic rates at the individual level.
+Whether these drivers act independently on individual-level metabolic rates remains uncertain.
+Most studies of intraspecific scaling of unitary organisms must rely on pre-existing differences in size to examine its relationship with metabolic rate, thereby potentially confounding size-correlated traits (e.g.
+age, nutrition) with size, which can affect metabolic rate.
+Here, we use a size manipulation approach to test whether metabolic mass scaling and temperature dependence interact in four species (two phyla) of colonial marine invertebrates.
+Size manipulation in colonial organisms allows to test how ecological processes (e.g.
+predation) affect individual physiology and consequently population- and community-level energy flux.
+Body mass and temperature interacted in two species, with one species exhibiting decreased and the other increased mass-scaling exponents with increasing temperature.
+The allometric scaling of metabolic rate that we observe in three species contrasts with the isometric scaling of ingestion rates observed in some colonial marine invertebrates.
+Thus, we suggest that the often observed competitive superiority of colonial over unitary organisms may arise because the difference between energy intake and expenditure increases more strongly with size in colonial organisms.
 

### 2016-05-19 paper.R
small fix

@@ -9,7 +9,7 @@ rm(list=ls())
 ###########
 knit('text/MS.Rmd', output=file.path(getwd(), 'text/MS.md'), quiet=TRUE, encoding = 'utf-8')
 # to word
-system('pandoc -o text/MS.docx text/MS.md -s -S --bibliography library.bib --csl functionalEcology.csl')
+system('pandoc -o text/MS.docx text/MS.md -s -S --bibliography library.bib --csl ecology.csl')
 
 ##############
 # COVER LETTER 

### 2016-05-19 text/MS.Rmd
more adjustments to language and text

@@
 
-Centre for Geometric Biology / School of Biological Sciences, Monash University, Clayton, Victoria, Australia 3800  
 \*Corresponding author: barnechedr@gmail.com
 
-**Summary**
+Affiliation:  
+Centre for Geometric Biology / School of Biological Sciences, Monash University, Clayton, Victoria, Australia 3800
+
+**Abstract**
 
@@
 
-Metabolic theories seek to unite hierarchical levels of biological organisation based on the scaling of energetic demands (i.e.
+Metabolic theories seek to unite hierarchical levels of biological organization based on the scaling of energetic demands (i.e.
 metabolic rates), from individuals to ecosystems `r myCite(c('nisbet2000janimecol', 'brown2004', 'vandermeer2006tree'))`.
 The role of individual body mass, $M_i$, as a determinant of individual-level metabolic rate, $B_i$, is key to this goal.
-Myriad studies have examined the relationship between $B_i$ and $M_i$ `r myCite(c('kleiber1932', 'kleiber1961', 'peters1983', 'schmidtnielsen1984', 'glazier2005biolrev', 'delong2010', 'glazier2010biolrev'))`, typically revealing a power function of the form $B_i = B_oM_i^{\alpha}$, where $B_o$ is a normalisation constant independent of mass and $\alpha$ is a dimensionless scaling exponent.
+Myriad studies have examined the relationship between $B_i$ and $M_i$ `r myCite(c('kleiber1932', 'kleiber1961', 'peters1983', 'schmidtnielsen1984', 'glazier2005biolrev', 'delong2010', 'glazier2010biolrev'))`, typically revealing a power function of the form $B_i = B_oM_i^{\alpha}$, where $B_o$ is a normalization constant independent of mass and $\alpha$ is a dimensionless scaling exponent.
 For most animals, $B_i$ scales sub-linearly (or negatively allometric, sensu `r myCite('huxley1932')`) with $M_i$, i.e.
@@
 For instance, it has been demonstrated that mimicking the effects of partial predation or physical disturbance through experimental size manipulation yields different responses in growth and reproductive onset in one arborescent (no effects on growth, delayed reproduction and reduced fecundity; `r myCite('bone2005jembe')`) and one encrusting bryozoan (reduced growth, with reproductive onset and fecundity dependent on the age of remaining zooids; `r myCite('hart2009ecologyb')`).
-On the other hand, size manipulations on one encrusting bryozoan `r myCite('white2011amnat')` and one colonial ascidian `r myCite('nakaya2005procb')` only affected the size of individuals, such that the scaling relationships between metabolic rate and colony size did not differ between size-manipulated and unmanipulated colonies.
+On the other hand, size manipulations on one encrusting bryozoan `r myCite('white2011amnat')` and one colonial ascidian `r myCite('nakaya2005procb')` only affected the size of individuals, such that the scaling relationships between metabolic rate and colony size did not differ between size-manipulated and non-manipulated colonies.
 Sub-linear allometric scaling, in turn, implies that mass-specific metabolic rates decrease with size, and therefore changes in size structure will influence total population- and community- level respiration flux (if total standing biomass is held fixed; `r myCite(c('allen2005', 'barneche2014ele'))`).
@@
 
-**Material and Methods**
+**Methods**
 
@@
 For encrusting colonies, sections of acetate bearing whole colonies were then cut from the sheets and fragmented (see e.g.
-`r myCite(c('nakaya2005procb', 'white2011amnat'))` for details and tests of potential experimental artefacts).
+`r myCite(c('nakaya2005procb', 'white2011amnat'))` for details and tests of potential experimental artifacts).
 Arborescent colonies were cut directly from the acetate sheets and fragmented.
@@
 
-where $m_b$ is the rate of change of $\textrm{O}_\textrm{2}$ saturation for blank vials containing no bryozoans (\% h^-1^), ${\beta}_{\textrm{O}_\textrm{2}}$ is the oxygen capacitance of air-saturated sea water at 10 ˚C (6.4 ml l^-1^) or 25 ˚C (4.74 ml l^-1^) `r myCite('cameron1986book')` and $V$ is water volume (vials were 0.005 l in volume and water volume was calculated by subtracting the volume of acetate and animals).
+where $m_b$ is the rate of change of $\textrm{O}_\textrm{2}$ saturation for blank vials containing no bryozoans (\% h^-1^), ${\beta}_{\textrm{O}_\textrm{2}}$ is the oxygen capacitance of air-saturated sea water at 10 ˚C (6.4 ml l^-1^) or 25 ˚C (4.74 ml l^-1^) `r myCite('cameron1986book')` and $V$ is water volume (vials were 5 ml in volume and water volume was calculated by subtracting the volume of acetate and animals).
 At least four blank vials were recorded simultaneously with each 24-well plate to account for microbial oxygen consumption and sensor spots were calibrated with air-saturated seawater (100\% A.S.) and water containing 2\% sodium sulfite (0\% A.S.).
 After a 1 h setting period (i.e.
-time necessary for the system to come into equilibrium), $\dot{V}\textrm{O}_\textrm{2}$ measurements lasted up to 20 h depending on individual mass (minute individuals require longer periods of time in order to detect $\dot{V}\textrm{O}_\textrm{2}$ signals).
+time necessary for the system to come into equilibrium), $\dot{V}\textrm{O}_\textrm{2}$ measurements lasted up to 20 h depending on individual mass (tiny individuals require much longer measurement periods in order to detect $\dot{V}\textrm{O}_\textrm{2}$).
 Following measurement of $\dot{V}\textrm{O}_\textrm{2}$, colonies or fragments were blotted dry and weighed to 0.1 mg (Sartorius A 200 S, Sartorius AG, Göttingen, Germany).
-All measurements of $\dot{V}\textrm{O}_\textrm{2}$ were made in one of four dark constant temperature cabinets, with temperatures within a run randomised among cabinets.
+All measurements of $\dot{V}\textrm{O}_\textrm{2}$ were made in one of four dark constant temperature cabinets, with temperatures within a run randomized among cabinets.
 
@@
 
-where mass (mg), $\textrm{ln}M_i$, is a continuous variable, while temperature (two levels: 10 ˚C and 25 ˚C) and species (four levels corresponding to the four species) are categorical variables.
-With this model, we allow each species to have a different mass-scaling slope that may change with temperature; we also account for species- and temperature-specific deviations on the metabolic normalisation (i.e.
+where mass, $\textrm{ln}M_i$ (mg), is a continuous variable, while temperature (two levels: 10 ˚C and 25 ˚C) and species (four levels corresponding to the four species) are categorical variables.
+With this model, we allow each species to have a different mass-scaling slope that may change with temperature; we also account for species- and temperature-specific deviations on the metabolic normalization (i.e.
 statistical model).
-We also included date of experiment, $\Delta D$ as a random effect on the metabolic normalisation to help control for natural experimental variability.
+We also included date of experiment, $\Delta D$ as a random effect on the metabolic normalization to help control for natural experimental variability.
 We then performed a model selection procedure (see below) by comparing the above full model with a series of nested two- and one-way interaction models in order to formally test for mass-temperature interactions in metabolic rates among different species.
@@
 1).
-Still, while the mass-scaling slopes are statistically the same between temperatures for these species, there is evidence for higher metabolic normalisations (i.e.
+Still, while the mass-scaling slopes are statistically the same between temperatures for these species, there is evidence for higher metabolic normalizations (i.e.
 intercepts), $B_o$, at higher temperatures (though marginally significant for Microcionidae, i.e.
-slight overlap between 95\% credible intervals).
+slight overlap between 95\% credible intervals; bottom-right quadrants in Fig.
+1).
 For example, metabolic rates increase `r rounded(exp(averageEstimates$lnB25[averageEstimates$species == 'Stolonifera'])/exp(averageEstimates$lnB10[averageEstimates$species == 'Stolonifera']))`-fold between 10 ˚C and 25 ˚C for *B.
@@
 
-Average posterior estimates of mass-scaling exponents are inconsistent among the studied species and between growth forms (2D and 3D).
+Average posterior estimates of mass-scaling exponents (top-left quadrants in Fig.
+1) are inconsistent among the studied species and between growth forms (2D and 3D).
 For example, for organisms that exhibit 2D growth, *Hippopodina iririkiensis* shows exponents that encompass both 0.67 (Fractal geometry model prediction for 2D organisms; `r myCite('koontz2009amjbot')`) and 0.75 (Fractal geometry model prediction for 3D organisms; `r myCite(c('west1997science', 'west1999science'))`) at 10 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's10_97.5'], 2)`), but that are significantly shallower than 0.75 at 25 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's25_97.5'], 2)`).
@@
 Possibly, the ability of our studied species to acclimate upon acute temperature changes could have affected the results (see discussion in `r myCite('glazier2014systems')`).
-Our experimental set up, however, randomised species and individual sizes across temperatures such that systematic effects across sizes or among species are unlikely.
+Our experimental set up, however, randomized species and individual sizes across temperatures such that systematic effects across sizes or among species are unlikely.
 Moreover, the temperature ranges span those experienced by these species in the field.
@@
 While the mass scaling exponent was unaffected by temperature for two species (*Bugula stolonifera* and Microcionidae), it decreased and increased for one encrusting (*Hippopodina iririkiensis*) and one arborescent species (*Bugula neritina*), respectively.
-These results imply that energy use may vary for the same species in different environments, possibly affecting their ultimate colonisation success and abundance.
+These results imply that energy use may vary for the same species in different environments, possibly affecting their ultimate colonization success and abundance.
 For example, the fact that *Bugula neritina* presented a shallower mass-scaling exponent at colder temperatures possibly indicates that large cold-environment individuals are more efficient in spending energy than counterparts in warm environments.
@@
 `r myCite('rosa2010jmba')`).
-Moreover, incorporation of differences in the size-dependence of the costs and benefits of size might improve current attempts to scale up energy from individuals to higher levels of organisation (e.g.
+Moreover, incorporation of differences in the size-dependence of the costs and benefits of size might improve current attempts to scale up energy from individuals to higher levels of organization (e.g.
 `r myCite(c('allen2005','yvondurocher2012b','barneche2014ele','barneche2016procb'))`).
@@
 
-**Table 2.** The best model was constructed, using the R package *lme4*, by successively removing fixed effects based on likelihood ratio tests of significance `r myCite('zuur2009book')`.
+*Table 2.* The best model was constructed, using the R package *lme4*, by successively removing fixed effects based on likelihood ratio tests of significance `r myCite('zuur2009book')`.
 In the table, $\chi^2$ and d.f.
@@
 
-**Figures**
+**Figure caption**
 
 *Figure 1.* Estimates from the three-way interaction model as obtained by Bayesian methods in *JAGS*.
-Results for each species are represented in groups of four plots, with blue and red colours representing actual data and parameter estimates at 10 ˚C and 25 ˚C, respectively.
-*Top-left* and *bottom-right*: posterior probabilities of mass-scaling exponents and metabolic normalisations (respectively), with bars over curves representing 95\% credible intervals, and numbers representing average estimates.
-*Bottom-left*: covariance between mass-scaling exponents and metabolic normalisations for each of the 10,000 MCMC iterations (translucent circles), with overlaid density heat-map.
+Results for each species are represented in groups of four plots, with blue and red colors representing actual data and parameter estimates at 10 ˚C and 25 ˚C, respectively.
+*Top-left* and *bottom-right*: posterior probabilities of mass-scaling exponents and metabolic normalizations (respectively), with bars over curves representing 95\% credible intervals, and numbers representing average estimates.
+*Bottom-left*: covariance between mass-scaling exponents and metabolic normalizations for each of the 10,000 MCMC iterations (translucent circles), with overlaid density heat-map.
 *Top-right*: regressions of individual metabolic rate, $\textrm{ln}B_i$, as functions of individual body mass, $\textrm{ln}M_i$.

### 2016-05-19 text/coverLetter.Rmd
edited cover letter

@@
-March 18 2016
+May 20 2016
 
-To the Editors of Functional Ecology,
+To the Editors of Ecology,
 
-Please find attached our manuscript entitled "Temperature effects on mass-scaling exponents in colonial animals: a manipulative test" for consideration as a Standard Paper in Functional Ecology.
+Please find attached our manuscript entitled "Temperature effects on mass-scaling exponents in colonial animals: a manipulative test" for consideration as an Article in Ecology.
 
 In this article, we use a manipulative approach to test for an association between metabolic mass scaling and temperature dependence in four species of colonial marine invertebrates belonging to two phyla.
-By adopting a size-manipulation approach, we achieved a more definitive test of mass-scaling of metabolic rates for cancelling the confounding effects of size and age, and by fragmenting colonies, a natural source of size variation in modular organisms in the wild.
+By adopting a size-manipulation approach, we achieved a more definitive test of mass-scaling of metabolic rates for canceling the confounding effects of size and age, and by fragmenting colonies, a natural source of size variation in modular organisms in the wild.
 Our results indicate that mass scaling and temperature dependence of metabolic rates interact in different ways across the studies species, and these interactions are not consistent across growth form nor phyla.
 
-Our study should be of general interest to readers of Functional Ecology because our findings suggest that different species might have different competitive advantages at different environmental temperatures.
+Our study should be of general interest to readers of Ecology because our findings suggest that different species might have different competitive advantages at different environmental temperatures.
 We suggest that the often observed competitive superiority of some colonial organisms over unitary ones may arise due to the differential scaling between metabolic and feeding rates -- energy availability for growth and or reproduction increases more strongly with size in colonial organisms compared to unitary ones.

### 2016-05-20 text/MS.Rmd
MS v 11 - submitted to ecology

@@
 Here, we use a size manipulation approach to test whether metabolic mass scaling and temperature dependence interact in four species (two phyla) of colonial marine invertebrates.
-Size manipulation in colonial organisms allows to test how ecological processes (e.g.
+Size manipulation in colonial organisms allows tests of how ecological processes (e.g.
 predation) affect individual physiology and consequently population- and community-level energy flux.
@@
 For example, the fractal geometry model `r myCite(c('west1997science', 'west1999science'))` predicts that $\alpha = n/(n + 1)$ for a $n$-dimensional organism, hence a value of $\alpha$ = 0.75 is predicted for three-dimensional animals while a value of $\alpha$ = 0.67 is predicted for two-dimensional animals (`r myCite(c('enquist2009nature', 'savage2008ploscompbiol', 'banavar2010pnas', 'kolokotrones2010nature', 'koontz2009amjbot'))`; but see `r myCite('price2007pnas')` for additional predictions of $\alpha$ upon modifications of canonical assumptions).
-Currently, however, it is uncertain how colonial organisms may conform to these predictions because the original fractal geometry model considers resource-distribution networks within unitary organisms (or individual zooids) and not colonies that might (or might not) distribute resources in fundamentally different ways.
+Currently, however, it is uncertain how colonial organisms may conform to these predictions because the original fractal geometry model considers resource-distribution networks within unitary organisms (or individual zooids) and not colonies that may (or may not) distribute resources in fundamentally different ways.
 Assessing the scaling of metabolic rate with body mass in colonial organisms is therefore a fundamental step in assessing how this mode of life conforms to general predictions, as well as providing an excellent opportunity for empirically manipulating mass.
@@
 On the other hand, size manipulations on one encrusting bryozoan `r myCite('white2011amnat')` and one colonial ascidian `r myCite('nakaya2005procb')` only affected the size of individuals, such that the scaling relationships between metabolic rate and colony size did not differ between size-manipulated and non-manipulated colonies.
-Sub-linear allometric scaling, in turn, implies that mass-specific metabolic rates decrease with size, and therefore changes in size structure will influence total population- and community- level respiration flux (if total standing biomass is held fixed; `r myCite(c('allen2005', 'barneche2014ele'))`).
+Sub-linear allometric scaling, in turn, implies that mass-specific metabolic rates decrease with size.
+Changes in size structure will therefore influence total population- and community- level respiration flux (if total standing biomass is held fixed; `r myCite(c('allen2005', 'barneche2014ele'))`).
 Meanwhile, resource ingestion rates appear to scale linearly with size, in some colonial organisms at least (e.g.
@@
 Possibly, the ability of our studied species to acclimate upon acute temperature changes could have affected the results (see discussion in `r myCite('glazier2014systems')`).
-Our experimental set up, however, randomized species and individual sizes across temperatures such that systematic effects across sizes or among species are unlikely.
+Our experimental procedure, however, randomized species and individual sizes across temperatures such that systematic effects across sizes or among species are unlikely.
 Moreover, the temperature ranges span those experienced by these species in the field.

### 2016-08-09 text/answers_to_reviewers.Rmd
first attempt to reply to reviewers comments

```{r, results="hide", warnings=FALSE, echo=FALSE, message=FALSE}
source('../R/functions-analyses.R')
cleanbib()
myRefs   <-  listBibs()
exporBibs(myRefs, erase=TRUE)
```
August 1, 2016

Dear Prof. Peter Raimondi,

Please find attached this resubmission version of our manuscript entitled "*Temperature effects on mass-scaling exponents in colonial animals: a manipulative test*" (original Manuscript ID: ECY16-0543) for consideration as an Article in Ecology. We would like to thank the reviewers for their insightful suggestions, most of which we have incorporated into this version of the MS, thereby increasing the clarity of the MS and the strength of its conclusions. In particular, we have improved considerably the description of our methodological set up (requested by both reviewers), and added new references and considerations regarding predation and density dependence of metabolic rates in the context of colonial organisms (requested by reviewer 2).  
Below you will find detailed responses to each one of the comments (in bold). Thank you for your consideration.

Sincerely,
Diego R. Barneche (on behalf of all coauthors) 
School of Biological Sciences 
Monash University
Melbourne, Australia 
barnechedr@gmail.com

Reviewer #1 (Comments to the Author):

This study presents some interesting results on scaling of oxygen consumption with mass in 4 taxa of colonial invertebrates and the effects of temperature on the scaling relationship. The conclusions arrived at and their implications are of general interest (particularly the identification of a potential advantage of colonial over unitary animals because their predicted energy availability for growth and reproduction increases more strongly with size). However, there are many undescribed aspects of the methods and material that are likely to influence how the results are interpreted and challenge the broad conclusions of the study. These issues are currently largely overlooked by the authors. Taken to the extreme, the confounding effects of these hidden issues may entirely discredit many of the current interpretations. In addition, the ecological significance of fragmentation is not sufficiently addressed (for example by demonstrating that fragmentation frequently occurs in natural populations). Below I outline the problems.

How was fragmentation of colonies done and what are the implications? Unfortunately, the reader is referred to seek details of fragmenting and tests of potential experimental artefacts as described in two studies: 1) Nakaya et al. (2015), and 2) White et al. (2011) (lines 112-114).  The former seems quite inappropriate as it is a study on an ascidian – techniques for fragmenting ascidians (which have extensive vascular connections amongst zooids) and the associated artefacts of such fragmentation will be very different than fragmenting encrusting bryozoans (and sponges and arborescent  bryozoans). White et al. (2011) worked on a different species of encrusting bryozoan than that studied by Barneche et al.  and it is not clear that fragmentation (which was presumably by cutting out circular discs from within colonies?) of one encrusting bryozoan species will entail similar artefacts in others.  Methods are meant to describe procedures so the study could be repeated but the specific details employed by Barneche et al. that would enable this to be done are almost entirely lacking. Was the sponge and arborescent bryozoan material cut in half and half again (serial fragmenting)? More generally, how many fragments were created per colony? And was the number of fragments of different sizes from different colonies always the same? If not, it would seem important to include genotype effects in all cases.  
**R: The reviewer makes a good point that not enough information was provided about the nature of fragmentation. The new version of our manuscript includes much more detail on this issue (page XX, lines XX), including a new schematic figure (new Figure 1). Briefly, for *Bugula neritina* and *Bugula stolonifera*, we reduced body size by cutting whole branches from the colony. This approach maintains the ratio of growing tips to core zooids and importantly mimics the kind of natural size reductions that *Bugula* spp. experience in the field when they are preyed upon by predatory nudibranchs `r myCite('bone2005jembe')`. For the encrusting species, we created 'pizza slice' style fragments to maintain the ratio of outer edge to inner colony as per the manipulation in White *et al.* `r myCite('white2011amnat')`. Importantly, the bryozoan used in this study has an extremely similar morphology and physiology to the one studied in White *et al.* `r myCite('white2011amnat')` so it seems that artifacts are unlikely given that White *et al.* `r myCite('white2011amnat')` comprehensively demonstrated no differences in metabolic rates between intact and experimentally fragmented colonies of similar body size. We maintained the same edge-to-inner-ratio approach by cutting pizza slices for sponges. As for the number of fragments per colony, for *Bugula* spp., this was strictly one per colony because we always cut whole branches to manipulate size, leaving the base intact. For the encrusting species, we used one to two fragments per colony and we did not record colony of origin so we cannot include genotype in our analyses. This potential caveat is now recognized in the methods (page XX, lines XX).**

Nakaya et al. (2015) measured oxygen consumption in an ascidian after recovery from fragmentation (waiting for take-over to occur and a new generation of daughter zooids to replace parental generation). No such recovery was provided by Barneche et al. The lack of a recovery period may pertain to fragmentation achieved by predation or storm damage but it is unclear how biologically relevant this is. Fragmentation and survival of branches of arborescent bryozoans may occur very rarely and it seems likely that the only fragments of arborescent colonies that will remain viable in the wild are the older attached portions. How does this influence interpretation of results (see also below)? Grazing, on the other hand, may more regularly fragment encrusting colonies, and such partial predation may often leave viable portions that are still attached to the substratum. The study would benefit from an attempt to reveal the frequency and extent of such fragmentation events in natural populations of the species studied - particularly because this is the only scenario for which the results of their manipulative study can really be interpreted (see below). This would bring an ecological context to what is currently an entirely laboratory-based investigation.  
**R: As we now note point out in the methods (page XX, lines XX), our fragmentation of *Bugula* spp. mimics predation in the field -- we only used older attached portions -- so we used the most ecologically relevant fragmentation approach for arborescent bryozoans. Although we have no formal estimates of the frequency and extent of fragmentation in our encrusting study species, there is very good evidence from a range of other encrusting species that such fragmentation events are common. Also, anecdotally, we commonly observed fragments of our study species during our study, indeed, it was difficult to obtain fully intact colonies in our experience because fragmentation is always so common.**

What was the life history status of the material?  The material used in the study was collected after 6 weeks in the field. By this time it seems likely that at least some if not all of the bryozoans or sponges were investing in sexual reproduction. This is a critical issue because ignoring it entails two problems for interpreting the results and their implications. First, fragments created from mature colonies that are investing in sexual reproduction may well not be equivalent to similarly-sized immature colonies that are only allocating energy to growth – nevertheless, much of the interpretations in the discussion section assume this to be the case. This means that ‘age’ potentially remains a confounding factor in the analysis, although it is claimed not to be (line 280). Second, investment in defensive polymorphs (in Bugula stolonifera and Hippopodina iririkiensis) may also increase with age and will thus entail similar problems in interpreting the results.  
**R: We focused on non-reproductive colonies to avoid the potentially complicating effects that reproduction might bring. The nature of our fragmentation approach meant that the age profile of the fragments was identical regardless of size, by cutting whole branches or pizza slices for arborescent and encrusting species respectively. As such, age is not a confounding factor in the analysis as all colonies were approximately the same age, regardless of size.**

NOTE: All the bryozoans that were studied produce ovicells in which larvae are brooded and their presence would be an indication of investment in sexual reproduction. Perhaps the material was saved and could be sampled for such activity? (Allocation to defensive polymorphs could also be assessed.) Sexual activity will be harder to determine for the sponge.  
**R: We are well practiced in detecting ovicells in bryozoans and we did not use reproductive colonies (at that time of year 6 weeks was not sufficient time for colonies to show reproduction). For the sponge, we could find no evidence of reproduction, though we  cannot rule out sperm production. We now make a note of this in the manuscript (page XX, lines XX). Importantly, however, our fragmentation approach maintained identical age profiles across all fragment sizes, and therefore age is not confounding our results.**

Related and other issues:

White et al. (2011) found that VO2 differed between intact colonies and colony fragments. This implies that it may be inappropriate to use both whole colonies and fragments in the same experiment. However, line 124 describes measuring VO2 by placing ‘the colony or fragment’ in a glass vial. Line 139 also implies both whole colonies and fragments were used. If this is the case then the effects of fragmentation vs. no fragmentation adds an additional complexity to the study that is not addressed.  
**R: We only used fragments, the use of the term colony was an unfortunate error based on using text from the earlier paper to describe the methods. This has now been corrected throughout the methods (page XX, lines XX, page XX, lines XX). Importantly, as we point out in our methods (page XX, lines XX, page XX, lines XX, page XX, lines XX), our earlier paper `r myCite('white2011amnat')` found no difference in scaling between intact and fragments of colonies, such that our results about scaling (which is the key issue in the discussion) almost certainly apply to intact colonies of different sizes as well.**

Line 130: presumably blank vials also contained no sponges?  
**R: The reviewer is correct. We have clarified this sentence replacing 'bryozoans' for 'animals' (page XX, lines XX).**

Reviewer #2 (Comments to the Author):

In this manuscript the authors test for interactive effects of body mass and temperature on the metabolic rates of four species of colonial marine invertebrates. By collecting individuals that colonized retrievable sheets, they were able to take apart the colonies and then reassemble them into colonies of different size, thus adding a size-manipulation component to their study. They then measured oxygen consumption and evaluated the effects of species, size, and temperature on oxygen consumption across experimental colonies.

I like this study because it 1) addresses group-level metabolic rates, 2) varies colonies experimentally, 3) tests for an effect of temperature on the size scaling, 4) does so for 4 species that grow either in two or three dimensions, and 5) analyzes the data all together. It strikes me as a solid study and an interesting contribution to the metabolic ecology literature, with implications for community ecology through competitive dominance and growth.  
**R: We really appreciate the encouraging words of Dr. DeLong, and that he is appreciative of our experimental approach.**

I would offer that the set-up could use a bit of clarification. The expectations for scaling slopes for different growth forms (2D versus 3D) are for individuals not colonies, so I’m not sure the expectations should really apply. You might just clarify why they still make sense anyway.  
**R: We strongly agree, and we would like to point out that we already mentioned this uncertainty regarding mass scaling exponents in colonial organisms in the introduction (page XX, lines XX): *"Currently, however, it is uncertain how colonial organisms may conform to these predictions because the original fractal geometry model considers resource-distribution networks within unitary organisms (or individual zooids) and not colonies that may (or may not) distribute resources in fundamentally different ways. Assessing the scaling of metabolic rate with body mass in colonial organisms is therefore a fundamental step in assessing how this mode of life conforms to general predictions, as well as providing an excellent opportunity for empirically manipulating mass."***

Also, because the measurements deal with colonies, you might touch on the role of population size/density in setting metabolic rates, since many organisms have lower per capita metabolic rates as population density increases, including colonial organisms (DeLong et al. 2014). The sublinear scaling exponents you find provide further evidence that population size is linked to metabolic rate, because the larger colonies have more individuals. This suggests some role for intra-specific processes to influence metabolic rates. It’s not clear what those should be, but perhaps this study could suggest some possibilities.  
**R: The reviewer makes an interesting point, but there is an important difference. We now point out in the discussion (including the recommended citation) this potential role of density-dependence energetic requirements at the colony level (page XX, lines XX): "*This new predation-mediated density dependence effect will increase total population energy flux, which is opposite to what is observed in unitary organisms that form groups or colonies, where group- or population-level metabolic rates decrease often due to increasing intraspecific competition (e.g. DeLong et al. 2014). We note, however, that a ‘colony’ here is a collection of physiologically connected interdependent zooids, which is different from a colony of independent unitary individuals, hence different mechanisms are likely to be operating as density of individuals changes.*".**

Should also probably mention the literature on scaling of metabolic rate in colonial insects. While not directly addressing an interaction between temperature and mass, a few papers have addressed scaling of individual and colony-level metabolic rate, with implications for competitive advantages and evolution of colonies. Check out (DeLong 2011, Shik et al. 2014).  
**R: We appreciate the suggestion. We have now incorporated DeLong 2011 in light of our findings in the discussion (page XX, lines XX).**

Little things

Line 55 – Should probably also mention Ginzburg and Damuth (2008) here, as they clearly raised this issue in their dimensionality paper.  
**R: Done (page XX, lines XX).**

Line 69 – In this list of ecological factors should go resource supply or density dependence, which has been shown to influence metabolic rates of individuals in groups (DeLong et al. 2014). In addition, you might want to look at (Glazier et al. 2011), who found an effect of predation risk on metabolic scaling exponents for amphipods.  
**R: Both studies are now incorporated in our discussion (page XX, lines XX).**

Line 227 – This might explain why group- and population-level energy use often increase with body mass, rather than showing energetic equivalence.  
**R: Point taken, and we now mention this potential in our discussion (page XX, lines XX).**

Line 260 – Seems like the Glazier et al 2011 result may be important here, as adding predation may change the scaling slope.  
**R: The study is now incorporated in the discussion (page XX, lines XX).**

Line 271 – Potentially also an effect of density here (DeLong et al. 2014).  
**R: The study is now incorporated in the discussion (page XX, lines XX).**

Line 465 – I can't tell which model is bolded.  
**R: Fixed (page XX, lines XX).**

Signed, John DeLong  
DeLong, J. P. 2011. Energetic inequivalence in eusocial insect colonies. Biology Letters 7:611–614.  
DeLong, J. P., T. C. Hanley, and D. A. Vasseur. 2014. Competition and the density dependence of metabolic rates. The Journal of animal ecology 83:51–58.  
Ginzburg, L., and J. Damuth. 2008. The space-lifetime hypothesis: viewing organisms in four dimensions, literally. The American Naturalist 171:125–131.  
Glazier, D. S., E. M. Butler, S. A. Lombardi, T. J. Deptola, A. J. Reese, and E. V. Satterthwaite. 2011. Ecological effects on metabolic scaling: amphipod responses to fish predators in freshwater springs. Ecological Monographs 81:599–618.  
Shik, J. Z., J. C. Santos, J. N. Seal, A. Kay, Mueller Ulrich G., and M. Kaspari. 2014. Metabolism and the rise of fungus cultivation by ants. The American Naturalist 184:364–373.

### 2016-08-09 paper.R

@@ -17,3 +17,10 @@ system('pandoc -o text/MS.docx text/MS.md -s -S --bibliography library.bib --csl
 knit('text/coverLetter.Rmd', output=file.path(getwd(), 'text/coverLetter.md'), quiet=TRUE, encoding = 'utf-8')
 # to word
 system('pandoc -o text/coverLetter.docx text/coverLetter.md')
+
+####################
+# ANSWER TO REFEREES 
+####################
+knit('text/answers_to_reviewers.Rmd', output=file.path(getwd(), 'text/answers_to_reviewers.md'), quiet=TRUE, encoding = 'utf-8')
+# to word
+system('pandoc -o text/answers_to_reviewers.docx text/answers_to_reviewers.md -s -S --bibliography library.bib --csl ecology.csl')

### 2016-08-16 text/answers_to_reviewers.Rmd
MS v 13 - resubmmitted to Ecology

@@
 myRefs   <-  listBibs()
-exporBibs(myRefs, erase=TRUE)
+exporBibs(myRefs, erase=FALSE)
 ```
-August 1, 2016
+August 16, 2016
 
@@
 For the encrusting species, we used one to two fragments per colony and we did not record colony of origin so we cannot include genotype in our analyses.
-This potential caveat is now recognized in the methods (page XX, lines XX).**
+However, as we now point out in the methods (page XX, lines XX), our goal was not to partition variance in metabolic rate in a Genotpye x Environment framework, instead, we were interested in how the phenotype of size influences metabolic rate under different temperatures.**
 
@@
 Second, investment in defensive polymorphs (in Bugula stolonifera and Hippopodina iririkiensis) may also increase with age and will thus entail similar problems in interpreting the results.
-**R: We focused on non-reproductive colonies to avoid the potentially complicating effects that reproduction might bring.
+**R: As we now note point out in the methods (page XX, lines XX), we focused on non-reproductive colonies to avoid the potentially complicating effects that reproduction might bring.
 The nature of our fragmentation approach meant that the age profile of the fragments was identical regardless of size, by cutting whole branches or pizza slices for arborescent and encrusting species respectively.
-As such, age is not a confounding factor in the analysis as all colonies were approximately the same age, regardless of size.**
+As such, age is not a confounding factor in the analysis as all colonies were approximately the same age, regardless of size (page XX, lines XX).**
 
@@
 We now make a note of this in the manuscript (page XX, lines XX).
-Importantly, however, our fragmentation approach maintained identical age profiles across all fragment sizes, and therefore age is not confounding our results.**
+Importantly, however, our fragmentation approach maintained identical age profiles across all fragment sizes, and therefore age is not confounding our results (page XX, lines XX).**
 
@@
 **R: We only used fragments, the use of the term colony was an unfortunate error based on using text from the earlier paper to describe the methods.
-This has now been corrected throughout the methods (page XX, lines XX, page XX, lines XX).
-Importantly, as we point out in our methods (page XX, lines XX, page XX, lines XX, page XX, lines XX), our earlier paper `r myCite('white2011amnat')` found no difference in scaling between intact and fragments of colonies, such that our results about scaling (which is the key issue in the discussion) almost certainly apply to intact colonies of different sizes as well.**
+This has now been corrected throughout the methods (page XX, line XX and XX).
+Importantly, as we point out in our methods (page XX, lines XX), `r myCite('white2011amnat')` found no difference in scaling between intact and fragments of colonies, such that our results about scaling (which is the key issue in the discussion) almost certainly apply to intact colonies of different sizes as well.**
 
@@
 2014).
-**R: We appreciate the suggestion. We have now incorporated DeLong 2011 in light of our findings in the discussion (page XX, lines XX).**
+**R: We appreciate the suggestion. We have now incorporated DeLong 2011 in light of our findings in the discussion (page XX, line XX).**
 
@@
 2011), who found an effect of predation risk on metabolic scaling exponents for amphipods.
-**R: Both studies are now incorporated in our discussion (page XX, lines XX).**
+**R: Both studies are now incorporated in our discussion (page XX, line XX; page XX, lines XX).**
 
@@
 Shik, J. Z., J. C. Santos, J. N. Seal, A. Kay, Mueller Ulrich G., and M. Kaspari. 2014. Metabolism and the rise of fungus cultivation by ants. The American Naturalist 184:364–373.
+
+**References**

### 2016-08-16 text/MS.Rmd
MS v 13 - resubmmitted to Ecology

@@
 Importantly, $\alpha$ estimates can also vary considerably among taxa and with many biotic and abiotic variables (e.g.
-`r myCite(c('lovegrove2000amnat', 'moses2008amnat', 'white2009evolution', 'delong2010', 'killen2010ele', 'kolokotrones2010nature', 'vaca2010biollett', 'glazier2011ecolmongr', 'ketola2012jlinsoc', 'white2012procb', 'white2013jcompphysb', 'barneche2014ele', 'hirst2014ele'))`).
+`r myCite(c('lovegrove2000amnat', 'moses2008amnat', 'white2009evolution', 'delong2010', 'killen2010ele', 'kolokotrones2010nature', 'vaca2010biollett', 'glazier2011ecolmonogr', 'ketola2012jlinsoc', 'white2012procb', 'white2013jcompphysb', 'barneche2014ele', 'hirst2014ele'))`).
 While these studies demonstrate clearly that the value of $\alpha$ is not fixed, most share a common limitation: they rely on pre-existing differences in mass.
@@
 `r myCite(c('pratt2005meps', 'hart2009ecologyb', 'white2011amnat'))`).
-Furthermore, different colonial species may grow in different dimensions, and the mass-scaling exponent $\alpha$ is explicitly predicted to change with number of body dimensions `r myCite(c('nakaya2005procb', 'white2011amnat', 'kearney2012amnat'))`.
+Furthermore, different colonial species may grow in different dimensions, and the mass-scaling exponent $\alpha$ is explicitly predicted to change with number of body dimensions `r myCite(c('nakaya2005procb', 'ginzburg2008amnat', 'white2011amnat', 'kearney2012amnat'))`.
 For example, the fractal geometry model `r myCite(c('west1997science', 'west1999science'))` predicts that $\alpha = n/(n + 1)$ for a $n$-dimensional organism, hence a value of $\alpha$ = 0.75 is predicted for three-dimensional animals while a value of $\alpha$ = 0.67 is predicted for two-dimensional animals (`r myCite(c('enquist2009nature', 'savage2008ploscompbiol', 'banavar2010pnas', 'kolokotrones2010nature', 'koontz2009amjbot'))`; but see `r myCite('price2007pnas')` for additional predictions of $\alpha$ upon modifications of canonical assumptions).
@@
 For instance, it has been demonstrated that mimicking the effects of partial predation or physical disturbance through experimental size manipulation yields different responses in growth and reproductive onset in one arborescent (no effects on growth, delayed reproduction and reduced fecundity; `r myCite('bone2005jembe')`) and one encrusting bryozoan (reduced growth, with reproductive onset and fecundity dependent on the age of remaining zooids; `r myCite('hart2009ecologyb')`).
-On the other hand, size manipulations on one encrusting bryozoan `r myCite('white2011amnat')` and one colonial ascidian `r myCite('nakaya2005procb')` only affected the size of individuals, such that the scaling relationships between metabolic rate and colony size did not differ between size-manipulated and non-manipulated colonies.
+On the other hand, size manipulations on one encrusting bryozoan `r myCite('white2011amnat')` and one colonial ascidian `r myCite('nakaya2005procb')` only affected the size of colonies, such that the scaling relationships between metabolic rate and colony size did not differ between size-manipulated and non-manipulated colonies.
 Sub-linear allometric scaling, in turn, implies that mass-specific metabolic rates decrease with size.
@@
 Arborescent colonies were cut directly from the acetate sheets and fragmented.
+We focused on non-reproductive colonies of bryozoans (at that time of year 6 weeks was not sufficient time for colonies to show reproduction) to avoid the potentially complicating effects that reproduction might bring.
+For the sponge, we found no evidence of reproduction, though we cannot rule out sperm production.
 
-*2. Respiration measurements*  
+*2. Fragmentation procedures*  
+For *Bugula neritina* and *B.
+stolonifera*, we reduced body size by cutting whole branches from the colony.
+This approach maintains the ratio of growing tips to core zooids (except for the base stolon which is present in all remaining fragments) and, importantly, mimics the kind of natural size reductions that *Bugula* spp.
+experience in the field when they are preyed upon by predatory nudibranchs `r myCite('bone2005jembe')`.
+For these arborescent species, we only used one fragment per colony because we always cut whole branches to manipulate size, leaving the base stolon intact (Fig.
+1, top).
+For the encrusting species, we created 'pizza slice' style fragments to maintain the ratio of outer growing edge to inner colony area (Fig.
+1, bottom), which is important for growth of fragment colonies `r myCite(c('hart2009ecologyb', 'white2011amnat'))`.
+Importantly, the encrusting bryozoan used in this study has an extremely similar morphology and physiology to the one studied in White et al.
+`r myCite('white2011amnat')`.
+Thus, it seems that artifacts are unlikely given that White et al.
+`r myCite('white2011amnat')` comprehensively demonstrated no differences in metabolic rates between intact and experimentally fragmented colonies of similar body size.
+We maintained the same edge-to-inner-ratio approach by cutting pizza slices for sponges.
+For the encrusting species, we used one to two fragments per colony.
+
+We note that we did not record colony of origin so we cannot include 'genotype' in our analyses.
+Importantly, our goal was not to partition variance in metabolic rate in a Genotype $\times$ Environment framework; instead, we were interested in how the phenotype of size influences metabolic rate under different temperatures.
+The nature of our fragmentation approach meant that the age profile of zooids within the fragments we created was identical, regardless of size, by cutting whole branches or pizza slices for arborescent and encrusting species respectively (Fig.
+1).
+As such, age was not a confounding factor in the analysis as all colonies were approximately the same age, regardless of size.
+
+*3. Respiration measurements*  
 Individual metabolic rates (converted to $\mu$l $\textrm{O}_\textrm{2}$ h^-1^, *n* = `r nrow(metRates)`), $B_i$, were measured as rate of oxygen consumption, $\dot{V}\textrm{O}_\textrm{2}$, of colony fragments at nominal temperatures of 10 or 25 ˚C.
@@
 Measurements were conducted using a 144-channel PreSens Sensor Dish Reader (SDR, AS-1 Scientific, Wellington, New Zealand) according to standard techniques `r myCite(c('koster2008meps', 'lighton2008book', 'white2011amnat'))`.
-$\dot{V}\textrm{O}_\textrm{2}$ was measured by placing the colony or fragment in a glass vial containing 0.2$\mu$m filtered sea water and a non-consumptive $\textrm{O}_\textrm{2}$ sensor spot and calculated from the rate of change of $\textrm{O}_\textrm{2}$ saturation ($m_a$, \% h^‑1^) as:
+$\dot{V}\textrm{O}_\textrm{2}$ was measured by placing the colony fragment in a glass vial containing 0.2$\mu$m filtered sea water and a non-consumptive $\textrm{O}_\textrm{2}$ sensor spot and calculated from the rate of change of $\textrm{O}_\textrm{2}$ saturation ($m_a$, \% h^‑1^) as:
 
@@
 
-where $m_b$ is the rate of change of $\textrm{O}_\textrm{2}$ saturation for blank vials containing no bryozoans (\% h^-1^), ${\beta}_{\textrm{O}_\textrm{2}}$ is the oxygen capacitance of air-saturated sea water at 10 ˚C (6.4 ml l^-1^) or 25 ˚C (4.74 ml l^-1^) `r myCite('cameron1986book')` and $V$ is water volume (vials were 5 ml in volume and water volume was calculated by subtracting the volume of acetate and animals).
+where $m_b$ is the rate of change of $\textrm{O}_\textrm{2}$ saturation for blank vials containing no animals (\% h^-1^), ${\beta}_{\textrm{O}_\textrm{2}}$ is the oxygen capacitance of air-saturated sea water at 10 ˚C (6.4 ml l^-1^) or 25 ˚C (4.74 ml l^-1^) `r myCite('cameron1986book')` and $V$ is water volume (vials were 5 ml in volume and water volume was calculated by subtracting the volume of acetate and animals).
 At least four blank vials were recorded simultaneously with each 24-well plate to account for microbial oxygen consumption and sensor spots were calibrated with air-saturated seawater (100\% A.S.) and water containing 2\% sodium sulfite (0\% A.S.).
-After a 1 h setting period (i.e.
+After a 1 h settling-in period (i.e.
 time necessary for the system to come into equilibrium), $\dot{V}\textrm{O}_\textrm{2}$ measurements lasted up to 20 h depending on individual mass (tiny individuals require much longer measurement periods in order to detect $\dot{V}\textrm{O}_\textrm{2}$).
-Following measurement of $\dot{V}\textrm{O}_\textrm{2}$, colonies or fragments were blotted dry and weighed to 0.1 mg (Sartorius A 200 S, Sartorius AG, Göttingen, Germany).
+Following measurement of $\dot{V}\textrm{O}_\textrm{2}$, colony fragments were blotted dry and weighed to 0.1 mg (Sartorius A 200 S, Sartorius AG, Göttingen, Germany).
 All measurements of $\dot{V}\textrm{O}_\textrm{2}$ were made in one of four dark constant temperature cabinets, with temperatures within a run randomized among cabinets.
 
-*3. Statistical analyses*  
+*4. Statistical analyses*  
 To test for the interacting effects of body mass and temperature on metabolic rate across species, we fitted a three-way interaction linear model using log-transformed data
@@
 Particularly, the best model yields no interaction between mass scaling and temperature dependence for *Bugula stolonifera* and the Microcionidae sponge (large overlap between 95\% credible intervals for the mass-scaling slopes, $\alpha$) (Fig.
-1).
+2).
 Still, while the mass-scaling slopes are statistically the same between temperatures for these species, there is evidence for higher metabolic normalizations (i.e.
@@
 slight overlap between 95\% credible intervals; bottom-right quadrants in Fig.
-1).
+2).
 For example, metabolic rates increase `r rounded(exp(averageEstimates$lnB25[averageEstimates$species == 'Stolonifera'])/exp(averageEstimates$lnB10[averageEstimates$species == 'Stolonifera']))`-fold between 10 ˚C and 25 ˚C for *B.
@@
 Average posterior estimates of mass-scaling exponents (top-left quadrants in Fig.
-1) are inconsistent among the studied species and between growth forms (2D and 3D).
+2) are inconsistent among the studied species and between growth forms (2D and 3D).
 For example, for organisms that exhibit 2D growth, *Hippopodina iririkiensis* shows exponents that encompass both 0.67 (Fractal geometry model prediction for 2D organisms; `r myCite('koontz2009amjbot')`) and 0.75 (Fractal geometry model prediction for 3D organisms; `r myCite(c('west1997science', 'west1999science'))`) at 10 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's10_97.5'], 2)`), but that are significantly shallower than 0.75 at 25 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's25_97.5'], 2)`).
@@
 
-The difference between the mass-scaling of ingestion rates and metabolic rates may explain the observed exponential growth and reproduction observed in some colonial marine invertebrates (e.g.
+The difference between the mass-scaling of ingestion rates and metabolic rates may explain the exponential growth and reproduction observed in some colonial marine invertebrates (e.g.
 `r myCite(c('marshall2013evolution', 'padilla1996limn', 'winston2010icb'))`) -- because the ratio of intake to expenditure increases with size, larger colonies can allocate proportionately more energy to growth and/or reproduction.
@@
 `r myCite(c('kooijman2009book', 'kearney2012amnat'))`) in a range of systems `r myCite(c('buss1980pnas', 'buss1990tree'))`.
+It also suggests that perhaps colony- and population-level energy use increase with individual body mass, rather than showing energetic equivalence `r myCite('damuth1987')`.
+Importantly, if that is the case, this scenario would add a new mechanism to explain deviations from energetic equivalence, which are thought, for example, to be a consequence of size-dependent mortality rates in eusocial insects `r myCite('delong2011biollett')` and energetic subsidies in reef fishes `r myCite('barneche2016procb')`.
 
@@
 
-In the wild, partial predation and physical disturbance are common sources of partial mortality in colonial organisms `r myCite('sebens1987annurevecolsys')`.
-Our findings suggest that, for some species, size reductions mediated by predation or disturbance might modulate the energetic demands of individuals of similar body mass in different environments, thus affecting life-history trajectories.
+In the wild, partial predation and physical disturbance are common sources of depression of metabolic rates in unitary organisms (e.g.
+`r myCite('glazier2011ecolmonogr')`) and partial mortality in colonial organisms `r myCite('sebens1987annurevecolsys')`.
+Our findings suggest that, for some colonial species, size reductions mediated by predation or disturbance might modulate the energetic demands of individuals of similar body mass in different environments, thus affecting life-history trajectories.
 For instance, at 25 ˚C, a 400-gram individual of *Bugula stolonifera* will spend ~`r round(metabolicDiff$metabolicDiff[metabolicDiff$species == 'Stolonifera'])`-fold more energy than a similar-sized individual at 10 ˚C.
@@
 For example, at 25 ˚C a population of 10 colonies of Microcionidae that each weigh 400 mg would have a cumulative metabolic demand of ~`r rounded(10 * exp(averageEstimates$lnB25[averageEstimates$species == 'Sponge'] + log(400)*averageEstimates$s25[averageEstimates$species == 'Sponge']), 0)` $\mu$l $\textrm{O}_\textrm{2}$ h^-1^.
-If a disturbance event halved each colony with minimal loss of biomass, then a population of those (now) 20 colonies, each weighing 200 mg, would have cumulative metabolic demand that is `r rounded((round(20 * exp(averageEstimates$lnB25[averageEstimates$species == 'Sponge'] + log(200)*averageEstimates$s25[averageEstimates$species == 'Sponge']))) / (round(10 * exp(averageEstimates$lnB25[averageEstimates$species == 'Sponge'] + log(400)*averageEstimates$s25[averageEstimates$species == 'Sponge']))), 1)`-fold higher (i.e.
+If a disturbance event halved each colony with minimal loss of biomass, then a population of those (now) 20 colonies (i.e.
+density twice as high), each weighing 200 mg, would have cumulative metabolic demand that is `r rounded((round(20 * exp(averageEstimates$lnB25[averageEstimates$species == 'Sponge'] + log(200)*averageEstimates$s25[averageEstimates$species == 'Sponge']))) / (round(10 * exp(averageEstimates$lnB25[averageEstimates$species == 'Sponge'] + log(400)*averageEstimates$s25[averageEstimates$species == 'Sponge']))), 1)`-fold higher (i.e.
 ~`r rounded(20 * exp(averageEstimates$lnB25[averageEstimates$species == 'Sponge'] + log(200)*averageEstimates$s25[averageEstimates$species == 'Sponge']), 0)` $\mu$l $\textrm{O}_\textrm{2}$ h^-1^), according to their new size.
-Thus, disturbance events change not only the size distribution of colonies but also the energy requirements of the population even if total standing biomass remains unchanged `r myCite('barneche2014ele')`.
+Thus, disturbance events change not only the size distribution of colonies but also increase the energy requirements of the population even if total standing biomass remains unchanged `r myCite('barneche2014ele')`.
+This new predation-mediated density dependence effect will increase total population energy flux, which is opposite to what is observed in unitary organisms that form groups or colonies, where group- or population-level metabolic rates often decrease due to increasing intraspecific competition (e.g.
+`r myCite('delong2014janimecol')`).
+We note, however, that a 'colony' here is a collection of physiologically connected interdependent zooids, which is different from a colony of independent unitary individuals, hence different mechanisms are likely to be operating as density of individuals changes.
 
@@
 
-We thank the Australian Research Council for financial support (grants DP0987626, DP110101776, FT130101493 to C.
+We thank John DeLong and three anonymous referees for insightful comments on earlier versions of this manuscript.
+We also thank the Australian Research Council for financial support (grants DP0987626, DP110101776, FT130101493 to C.
 R.
@@
 
-*Figure 1.* Estimates from the three-way interaction model as obtained by Bayesian methods in *JAGS*.
+*Figure 1.* Schematic illustration of our fragmentation approach.
+Top: Demonstration of how whole branches were cut (read and blue lines) in arborescent 3D bryozoans (*Bugula neritina* and *B.
+stolonifera*), and the resulting fragments (red and blue colonies) used for respirometry.
+Importantly, this approach maintains the ratio of growing tips to core zooids and importantly mimics the kind of natural size reductions that *Bugula* spp.
+experience in the field Bottom: Approach adopted with encrusting 2D organisms (the bryozoan *Hippopodina iririkiensis* and the Microcionidae sponge), where, based on previous studies, pizza slices were used in order to maintain a constant ratio between individual growing edge (G) and area (A).
+
+*Figure 2.* Estimates from the three-way interaction model as obtained by Bayesian methods in *JAGS*.
 Results for each species are represented in groups of four plots, with blue and red colors representing actual data and parameter estimates at 10 ˚C and 25 ˚C, respectively.

### 2016-10-05 text/MS.Rmd
version accepted in Ecology; figure 1 made from code now; updated readme

@@
 Sub-linear allometric scaling, in turn, implies that mass-specific metabolic rates decrease with size.
-Changes in size structure will therefore influence total population- and community- level respiration flux (if total standing biomass is held fixed; `r myCite(c('allen2005', 'barneche2014ele'))`).
+Changes in size structure will therefore influence total population- and community-level respiration flux (if total standing biomass is held fixed; `r myCite(c('allen2005', 'barneche2014ele'))`).
 Meanwhile, resource ingestion rates appear to scale linearly with size, in some colonial organisms at least (e.g.
@@
 With this model, we allow each species to have a different mass-scaling slope that may change with temperature; we also account for species- and temperature-specific deviations on the metabolic normalization (i.e.
-statistical model).
+model intercept).
 We also included date of experiment, $\Delta D$ as a random effect on the metabolic normalization to help control for natural experimental variability.
@@
 *Figure 1.* Schematic illustration of our fragmentation approach.
-Top: Demonstration of how whole branches were cut (read and blue lines) in arborescent 3D bryozoans (*Bugula neritina* and *B.
-stolonifera*), and the resulting fragments (red and blue colonies) used for respirometry.
+Top: Demonstration of how whole branches were cut (dark- and light-grey lines) in arborescent 3D bryozoans (*Bugula neritina* and *B.
+stolonifera*), and the resulting fragments (dark- and light-grey colonies) used for respirometry.
 Importantly, this approach maintains the ratio of growing tips to core zooids and importantly mimics the kind of natural size reductions that *Bugula* spp.
@@
 *Figure 2.* Estimates from the three-way interaction model as obtained by Bayesian methods in *JAGS*.
-Results for each species are represented in groups of four plots, with blue and red colors representing actual data and parameter estimates at 10 ˚C and 25 ˚C, respectively.
+Results for each species are represented in groups of four plots, with grey and black colors representing actual data and parameter estimates at 10 ˚C and 25 ˚C, respectively.
 *Top-left* and *bottom-right*: posterior probabilities of mass-scaling exponents and metabolic normalizations (respectively), with bars over curves representing 95\% credible intervals, and numbers representing average estimates.

### 2016-10-05 R/functions-figures.R

@@ -2,219 +2,305 @@
 # AUXILLIARY FUNCTIONS
 ######################
 toPdf <- function(expr, filename, ...) {
-  toDev(expr, pdf, filename, ...)
+    toDev(expr, pdf, filename, ...)
 }
 
 toDev <- function(expr, dev, filename, ..., verbose=TRUE) {
-  if ( verbose )
-    cat(sprintf('Creating %s\n', filename))
-  dev(filename, family='CM Roman', ...)
-  on.exit(dev.off())
-  eval.parent(substitute(expr))
+    if ( verbose )
+        cat(sprintf('Creating %s\n', filename))
+    dev(filename, family='CM Roman', ...)
+    on.exit(dev.off())
+    eval.parent(substitute(expr))
 }
 
 rounded  <-  function(value, precision=1, change=FALSE) {
-  if(change) {
-    value  <-  value * -1
-  }
-  sprintf(paste0('%.', precision, 'f'), round(value, precision))
+    if(change) {
+        value  <-  value * -1
+    }
+    sprintf(paste0('%.', precision, 'f'), round(value, precision))
+}
+
+circle  <-  function(x, y, radius, start=0, end=pi, nsteps=100,...){  
+    pinRatio  <-  par('pin')[2]/par('pin')[1]
+    rs        <-  seq(start, end, len=nsteps) 
+    xc        <-  x + radius * cos(rs)
+    yc        <-  (y + radius * sin(rs)) / pinRatio
+    lines(xc,yc,...) 
+}
+
+proportionalPng  <-  function(logo, px, py, log = FALSE, ...) {
+    if(!is.numeric(px) | !is.numeric(py) | length(px) != 2) {
+        stop('wrong position coordinates [0,1]')
+    }
+    usr  <-  par('usr')
+    pin  <-  par('pin')
+    # first get proportions of coordinates right
+    pxRg    <-  px[2] - px[1]
+    xProp   <-  usr[1] + px * (usr[2] - usr[1]) # x range from relative px 
+    yProp   <-  usr[3] + py * (usr[4] - usr[3]) # minimum y from relative py
[…]

### 2016-10-05 figures.R

@@ -1,15 +1,18 @@
 library(lme4)
 library(LoLinR)
+library(png)
 library(MASS)
 library(RColorBrewer)
 library(extrafont)
 library(fontcm)
-loadfonts()
+loadfonts(quiet=TRUE)
 
 rm(list=ls())
 source('paths.R')
 source('R/functions-figures.R')
 load('output/RDatafiles/analyses.RData')
 
-toPdf(fig1(), 'output/figures/fig1.pdf', width=8, height=7)
+toPdf(fig1(), 'output/figures/fig1.pdf', width=8, height=5)
 embed_fonts('output/figures/fig1.pdf')
+toPdf(fig2(), 'output/figures/fig2.pdf', width=8, height=7)
+embed_fonts('output/figures/fig2.pdf')

### 2016-10-08 text/MS.Rmd
attempt 1 to restructure code using remake

@@
 ```{r, results="hide", warnings=FALSE, echo=FALSE, message=FALSE}
-load('../output/RDatafiles/analyses.RData')
-load('../output/RDatafiles/numbers.RData')
-source('../R/functions-analyses.R')
-source('../R/functions-figures.R')
 cleanbib()
@@
 
+```{r}
+ci95Estimates     <-  numbersList$ci95Estimates
+metabolicDiff     <-  numbersList$metabolicDiff
+averageEstimates  <-  numbersList$averageEstimates
+q10andErs         <-  numbersList$q10andErs
+tempDeltas        <-  numbersList$tempDeltas
+```
+
 The model selection indicated that the three-way interaction model (equation 2) performed significantly better than all nested two- and one-way interaction models (Table 2), thus implying that different species present different interactions between mass-scaling and temperature dependence.

### 2016-10-08 R/functions-analyses.R

@@ -1,198 +1,276 @@
-#########
-# GENERAL
-#########
-readFile  <-  function(path, ...) {
-	read.csv(path, header=TRUE, stringsAsFactors=FALSE, ...)
+######
+# DATA
+######
+readAndCleanData  <-  function(dataPath) {
+    metRates  <-  read.csv(dataPath, header=TRUE, stringsAsFactors=FALSE)
+    metRates  <-  metRates[complete.cases(metRates), ]
+    
+    # correct log values from log 10 to ln
+    metRates$lnMass  <-  log(10^metRates$logM)
+    metRates$lnRate  <-  log(10^metRates$logVo2)
+    
+    # create inverse temperature column
+    metRates$TempK  <-  metRates$Temp + 273.15
+    metRates$invKT  <-  1 / 8.62e-5 * (1 / 293.15 - 1 / metRates$TempK)
+    
+    # factor temperature column for ancova
+    metRates$Temp         <-  as.factor(metRates$Temp)
+    metRates
 }
 
-rmFcts  <-  function(filename) {
-  source(filename, local=TRUE)
-  fct_files  <-  ls()
-  rm(list=ls(.GlobalEnv)[ls(.GlobalEnv) %in% fct_files], envir=.GlobalEnv)
+########
+# TABLES
+########
+makeTables  <-  function(dataTable1, dataTable2) {
+    makeTable1(dataTable1)
+    makeTable2(dataTable2)
+}
+
+makeTable1  <-  function(data) {
+    tab1  <-  ddply(data, .(Species, Temp), function(x) {
+        data.frame(minMass = rounded(min(exp(x$lnMass))),
+                   maxMass = rounded(max(exp(x$lnMass))),
+                   n         = nrow(x), stringsAsFactors=FALSE)
+    })
+    write.csv(tab1, 'output/data/table1.csv', row.names=FALSE)
+}
+
+makeTable2  <-  function(data) {
+    tab2  <-  data.frame()
+    for(j in 2:9) {
+        refModel  <-  data[[paste0('lmerModel', j)]]
+        aovTab    <-  data.frame(anova(data$lmerModel1, refModel))['lmerModel1',]
+        tab2      <-  rbind(tab2, data.frame('d.f.'=aovTab[,'Chi.Df'], 'Chisq'=aovTab[,'Chisq'], 'P'=aovTab[,'Pr..Chisq.'], stringsAsFactors=FALSE))
+    }
[…]

### 2016-10-08 analyses.R

@@ -1,31 +1,0 @@
-library(lme4)
-library(R2jags)
-
-rm(list=ls())
-source('paths.R')
-source('dataManipulation.R')
-source('R/functions-analyses.R')
-
-# compare random effects among models
-lmerModela  <-  lmer(lnRate ~ Species + lnMass + Temp + Species:lnMass + Species:Temp + Temp:lnMass + Temp:lnMass:Species + (1|Run) + (1|Plate), data=metRates, REML=FALSE)
-lmerModelb  <-  lmer(lnRate ~ Species + lnMass + Temp + Species:lnMass + Species:Temp + Temp:lnMass + Temp:lnMass:Species + (1|Plate), data=metRates, REML=FALSE)
-lmerModelc  <-  lmer(lnRate ~ Species + lnMass + Temp + Species:lnMass + Species:Temp + Temp:lnMass + Temp:lnMass:Species + (1|Run), data=metRates, REML=FALSE)
-modelSelectionRandomTab  <-  anova(lmerModela, lmerModelb, lmerModelc)
-
-# lmerModelc most significant. Now run models to asses significance of fixed effects
-lmerModel1  <-  lmer(lnRate ~ Species + lnMass + Temp + Species:lnMass + Species:Temp + Temp:lnMass + Temp:lnMass:Species + (1|Run), data=metRates, REML=FALSE)
-lmerModel2  <-  lmer(lnRate ~ Species + lnMass + Temp + Species:lnMass + Species:Temp + Temp:lnMass + (1|Run), data=metRates, REML=FALSE)
-lmerModel3  <-  lmer(lnRate ~ Species + lnMass + Temp + Species:lnMass + Species:Temp + (1|Run), data=metRates, REML=FALSE)
-lmerModel4  <-  lmer(lnRate ~ Species + lnMass + Temp + Species:lnMass + Temp:lnMass + (1|Run), data=metRates, REML=FALSE)
-lmerModel5  <-  lmer(lnRate ~ Species + lnMass + Temp + Species:Temp + Temp:lnMass + (1|Run), data=metRates, REML=FALSE)
-lmerModel6  <-  lmer(lnRate ~ Species + lnMass + Temp + Species:lnMass + (1|Run), data=metRates, REML=FALSE)
-lmerModel7  <-  lmer(lnRate ~ Species + lnMass + Temp + Species:Temp + (1|Run), data=metRates, REML=FALSE)
[…]

### 2016-10-08 dataManipulation.R

@@ -1,13 +1,0 @@
-metRates  <-  read.csv('data/data.csv', header=TRUE, stringsAsFactors=FALSE)
-metRates  <-  metRates[complete.cases(metRates), ]
-
-# correct log values from log 10 to ln
-metRates$lnMass  <-  log(10^metRates$logM)
-metRates$lnRate  <-  log(10^metRates$logVo2)
-
-# create inverse temperature column
-metRates$TempK  <-  metRates$Temp + 273.15
-metRates$invKT  <-  1 / 8.62e-5 * (1 / 293.15 - 1 / metRates$TempK)
-
-# factor temperature column for ancova
-metRates$Temp         <-  as.factor(metRates$Temp)

### 2016-10-08 figures.R

@@ -1,17 +1,4 @@
-library(lme4)
-library(LoLinR)
-library(png)
-library(MASS)
-library(RColorBrewer)
-library(extrafont)
-library(fontcm)
 loadfonts(quiet=TRUE)
-
-rm(list=ls())
-source('paths.R')
-source('R/functions-figures.R')
-load('output/RDatafiles/analyses.RData')
-
 toPdf(fig1(), 'output/figures/fig1.pdf', width=8, height=5)
 embed_fonts('output/figures/fig1.pdf')
 toPdf(fig2(), 'output/figures/fig2.pdf', width=8, height=7)

### 2016-10-08 numbers.R

@@ -1,24 +1,0 @@
-library(plyr)
-library(lme4)
-
-rm(list=ls())
-source('paths.R')
-source('R/functions-analyses.R')
-load('output/RDatafiles/analyses.RData')
-
-# get average estimates and 95% CI for each parameter
-species           <-  unique(metRates$Species)
-averageEstimates  <-  ldply(species, getAverageEstimates)
-ci95Estimates     <-  ldply(species, get95CIEstimates)
-
-# compare deltas between temperatures for each species
-tempDeltas  <-  ddply(averageEstimates, .(species), getTempDeltas)
-
-# metabolic difference (in n-fold) between different temperatures for similar-sized animals
-metabolicDiff  <-  ddply(averageEstimates, .(species), getMetabolicDiff, lnMass=log(400))
-
-# q10-equivalents [ q10 == (r25 / r10) ^ (10/(25-10)) ] for Microcionidae and Bugula stolonifera
-q10andErs  <-  ddply(averageEstimates[averageEstimates$species %in% c('Sponge', 'Stolonifera'), ], .(species), getq10andEr)
-
-rm(list=ls()[!(ls() %in% c('species', 'averageEstimates', 'ci95Estimates', 'tempDeltas', 'metabolicDiff', 'q10andErs'))])
-save.image('output/RDatafiles/numbers.RData')

### 2016-10-08 paper.R

@@ -1,9 +1,3 @@
-library(knitr)
-library(bibtex)
-library(knitcitations)
-library(plyr)
-
-rm(list=ls())
 ###########
 # MAIN TEXT 
 ###########

### 2016-10-08 tables.R

@@ -1,25 +1,0 @@
-library(lme4)
-library(plyr)
-rm(list=ls())
-source('paths.R')
-source('R/functions-analyses.R')
-source('R/functions-figures.R')
-load('output/RDatafiles/analyses.RData')
-
-# table 1
-table1  <-  ddply(metRates, .(Species, Temp), function(x) {
-	data.frame(minMass = rounded(min(exp(x$lnMass))),
-			   maxMass = rounded(max(exp(x$lnMass))),
-			   n         = nrow(x), stringsAsFactors=FALSE)
-})
-write.csv(table1, 'output/data/table1.csv', row.names=FALSE)
-
-# table 2
-table2  <-  data.frame()
-for(j in 2:9) {
-	refModel  <-  get(paste0('lmerModel', j))
-	aovTab    <-  data.frame(anova(lmerModel1, refModel))['lmerModel1',]
-	table2    <-  rbind(table2, data.frame('d.f.'=aovTab[,'Chi.Df'], 'Chisq'=aovTab[,'Chisq'], 'P'=aovTab[,'Pr..Chisq.'], stringsAsFactors=FALSE))
-}
-table2$P  <-  sapply(table2$P, cleanPvals)
-write.csv(table2, 'output/data/table2.csv', row.names=FALSE)

### 2016-10-09 text/MS.Rmd
reshaped code to work with remake

--- 2016-10-05 text/MS.Rmd
@@
-```{r, results="hide", warnings=FALSE, echo=FALSE, message=FALSE}
-load('../output/RDatafiles/analyses.RData')
-load('../output/RDatafiles/numbers.RData')
-source('../R/functions-analyses.R')
-source('../R/functions-figures.R')
-cleanbib()
-myRefs   <-  listBibs()
-exporBibs(myRefs, erase=TRUE)
-```
 **Temperature effects on mass-scaling exponents in colonial animals: a manipulative test**
@@
 *3. Respiration measurements*  
-Individual metabolic rates (converted to $\mu$l $\textrm{O}_\textrm{2}$ h^-1^, *n* = `r nrow(metRates)`), $B_i$, were measured as rate of oxygen consumption, $\dot{V}\textrm{O}_\textrm{2}$, of colony fragments at nominal temperatures of 10 or 25 ˚C.
+Individual metabolic rates (converted to $\mu$l $\textrm{O}_\textrm{2}$ h^-1^, *n* = `r nrow(numbersList$metRates)`), $B_i$, were measured as rate of oxygen consumption, $\dot{V}\textrm{O}_\textrm{2}$, of colony fragments at nominal temperatures of 10 or 25 ˚C.
 All individuals (regardless of size) and species were randomly assigned to test temperatures, and we used replicate vials and controlled temperature cabinets (see below).
@@
 **Results**
+
+```{r, echo=FALSE}
+ci95Estimates     <-  numbersList$ci95Estimates
+metabolicDiff     <-  numbersList$metabolicDiff
+averageEstimates  <-  numbersList$averageEstimates
+q10andErs         <-  numbersList$q10andErs
+tempDeltas        <-  numbersList$tempDeltas
+```
 

### 2016-10-09 R/functions-analyses.R

@@ -21,9 +21,10 @@ readAndCleanData  <-  function(dataPath) {
 ########
 # TABLES
 ########
-makeTables  <-  function(dataTable1, dataTable2) {
+makeTables  <-  function(dataTable1, pathToDataTable2) {
     makeTable1(dataTable1)
-    makeTable2(dataTable2)
+    load(pathToDataTable2)
+    makeTable2(output)
 }
 
 makeTable1  <-  function(data) {
@@ -33,17 +34,19 @@ makeTable1  <-  function(data) {
                    n         = nrow(x), stringsAsFactors=FALSE)
     })
     write.csv(tab1, 'output/data/table1.csv', row.names=FALSE)
+    cat('Table 1 written to output/data/table1.csv \n')
 }
 
 makeTable2  <-  function(data) {
     tab2  <-  data.frame()
     for(j in 2:9) {
         refModel  <-  data[[paste0('lmerModel', j)]]
-        aovTab    <-  data.frame(anova(data$lmerModel1, refModel))['lmerModel1',]
+        aovTab    <-  data.frame(anova(data$lmerModel1, refModel))['data$lmerModel1',]
         tab2      <-  rbind(tab2, data.frame('d.f.'=aovTab[,'Chi.Df'], 'Chisq'=aovTab[,'Chisq'], 'P'=aovTab[,'Pr..Chisq.'], stringsAsFactors=FALSE))
     }
     tab2$P  <-  sapply(tab2$P, cleanPvals)
     write.csv(tab2, 'output/data/table2.csv', row.names=FALSE)
+    cat('Table 2 written to output/data/table2.csv \n')
 }
 
 cleanPvals  <-  function(x) {
@@ -114,7 +117,7 @@ jagsModel  <-  function(run=TRUE, outputFileName) {
     write(jmod, outputFileName)
 }
 
-fitJags  <-  function(lmModelMatrix, run=TRUE, data, outputFileName='mod.bug') {
+fitJags  <-  function(lmModelMatrix, data, run=TRUE, outputFileName='mod.bug') {
     # fitting an interaction model
     # first extract model matrix from lm model
     tjagsModelMatrix  <-  model.matrix(lmModelMatrix)
@@ -138,13 +141,15 @@ fitJags  <-  function(lmModelMatrix, run=TRUE, data, outputFileName='mod.bug') {
 ###############
 # PAPER NUMBERS
 ###############
-extractNumbersList  <-  function() {
+extractNumbersList  <-  function(metRates, outputPath) {
+    load(outputPath)
[…]

### 2016-10-09 R/functions-figures.R

@@ -75,6 +75,16 @@ whiteGrid  <-  function (...) {
         col = 'grey70', lwd = 0.3, ...)
 }
 
+makeFigure1  <-  function() {
+    toPdf(fig1(), 'output/figures/fig1.pdf', width=8, height=5)
+    embed_fonts('output/figures/fig1.pdf')
+}
+
+makeFigure2  <-  function(...) {
+    toPdf(fig2(...), 'output/figures/fig2.pdf', width=8, height=7)
+    embed_fonts('output/figures/fig2.pdf')
+}
+
 ###############
 # PAPER FIGURES
 ###############
@@ -94,7 +104,7 @@ getProb  <-  function(densObjt) {
     densObjt
 }
 
-plotTriad  <-  function(species=species, spNames=spNames, tmat, x1=TRUE, y1=TRUE, x2=TRUE, y2=TRUE, x3=TRUE, y3=TRUE, x4=TRUE, y4=TRUE) {
+plotTriad  <-  function(species=species, spNames=spNames, tmat, dat, x1=TRUE, y1=TRUE, x2=TRUE, y2=TRUE, x3=TRUE, y3=TRUE, x4=TRUE, y4=TRUE) {
     col10     <-  colorRampPalette(rev(brewer.pal(9, 'Greys')))
     col10     <-  col10(32)
     col10[1]  <-  '#FFFFFF'
@@ -185,7 +195,6 @@ plotTriad  <-  function(species=species, spNames=spNames, tmat, x1=TRUE, y1=TRUE
     proportionalLabel(figureProp(quants), rep(0.86, 2), text=FALSE, type='l', col='black', lwd=1.2)
     proportionalLabel(0.97, 0.86, rounded(mean(s25), 2), adj=c(1, 0.5), col='black', cex=0.8)
     
-    dat         <-  metRates[metRates$Species == species, ]
     dat$colors  <-  c('grey50', 'black')[match(dat$TempK, c(283.15, 298.15))]
     plot(NA, xlab='', ylab='', xlim=c(-1, 7), ylim=c(-6, 6), axes=FALSE, type='n')
     usr  <-  par('usr')
@@ -273,11 +282,12 @@ fig1  <-  function() {
     text(0.5, -0.14/pinRatio, 'G / 4', adj=c(0, 0.5), col='grey60', cex=1.3, font=3)
 }
 
[…]

### 2016-10-09 figures.R

@@ -1,5 +1,0 @@
-loadfonts(quiet=TRUE)
-toPdf(fig1(), 'output/figures/fig1.pdf', width=8, height=5)
-embed_fonts('output/figures/fig1.pdf')
-toPdf(fig2(), 'output/figures/fig2.pdf', width=8, height=7)
-embed_fonts('output/figures/fig2.pdf')

### 2016-10-09 outputRequirements.R

@@ -1,3 +1,4 @@
+extrafont::loadfonts(quiet=TRUE)
 dir.create('output/data',  recursive=TRUE, showWarnings=FALSE)
 dir.create('output/figures',  recursive=TRUE, showWarnings=FALSE)
 dir.create('output/RDatafiles', recursive=TRUE, showWarnings=FALSE)

### 2016-10-09 paper.R

@@ -1,9 +1,10 @@
+library(knitr)
+
 ###########
 # MAIN TEXT 
 ###########
-knit('text/MS.Rmd', output=file.path(getwd(), 'text/MS.md'), quiet=TRUE, encoding = 'utf-8')
 # to word
-system('pandoc -o text/MS.docx text/MS.md -s -S --bibliography library.bib --csl ecology.csl')
+system('pandoc -o text/MS.docx text/MS.md -s -S --bibliography references.bib --csl ecology.csl')
 
 ##############
 # COVER LETTER 
@@ -17,4 +18,4 @@ system('pandoc -o text/coverLetter.docx text/coverLetter.md')
 ####################
 knit('text/answers_to_reviewers.Rmd', output=file.path(getwd(), 'text/answers_to_reviewers.md'), quiet=TRUE, encoding = 'utf-8')
 # to word
-system('pandoc -o text/answers_to_reviewers.docx text/answers_to_reviewers.md -s -S --bibliography library.bib --csl ecology.csl')
+system('pandoc -o text/answers_to_reviewers.docx text/answers_to_reviewers.md -s -S --bibliography references.bib --csl ecology.csl')

### 2016-10-09 sh/cleanCSV.sh
removed unncessary files and renamed paper to wordFiles

@@ -1,7 +1,0 @@
-#to run type bash sh/cleanCSV.sh
-cd data/
-for file in *.csv
-do
-   perl -pi -e 's/\r/\n/g' $file
-done
-cd ..

### 2016-10-09 text/answers_to_reviewers.Rmd
Some suggestions to simplify workflow
- use rmarkdown to call pandoc via remake
- don't need to store output in Rdata
- simplify readme, add instructions for obtaining JAGS
- stash extra files in text/

@@
-```{r, results="hide", warnings=FALSE, echo=FALSE, message=FALSE}
+<!-- ```{r, results="hide", warnings=FALSE, echo=FALSE, message=FALSE}
 source('../R/functions-analyses.R')
@@
 ```
+ -->
 August 16, 2016

### 2016-10-09 R/functions-analyses.R

@@ -21,23 +21,16 @@ readAndCleanData  <-  function(dataPath) {
 ########
 # TABLES
 ########
-makeTables  <-  function(dataTable1, pathToDataTable2) {
-    makeTable1(dataTable1)
-    load(pathToDataTable2)
-    makeTable2(output)
-}
-
-makeTable1  <-  function(data) {
+makeTable1  <-  function(data, dest) {
     tab1  <-  ddply(data, .(Species, Temp), function(x) {
         data.frame(minMass = rounded(min(exp(x$lnMass))),
                    maxMass = rounded(max(exp(x$lnMass))),
                    n         = nrow(x), stringsAsFactors=FALSE)
     })
-    write.csv(tab1, 'output/data/table1.csv', row.names=FALSE)
-    cat('Table 1 written to output/data/table1.csv \n')
+    write.csv(tab1, dest, row.names=FALSE)
 }
 
-makeTable2  <-  function(data) {
+makeTable2  <-  function(data, dest) {
     tab2  <-  data.frame()
     for(j in 2:9) {
         refModel  <-  data[[paste0('lmerModel', j)]]
@@ -45,8 +38,7 @@ makeTable2  <-  function(data) {
         tab2      <-  rbind(tab2, data.frame('d.f.'=aovTab[,'Chi.Df'], 'Chisq'=aovTab[,'Chisq'], 'P'=aovTab[,'Pr..Chisq.'], stringsAsFactors=FALSE))
     }
     tab2$P  <-  sapply(tab2$P, cleanPvals)
-    write.csv(tab2, 'output/data/table2.csv', row.names=FALSE)
-    cat('Table 2 written to output/data/table2.csv \n')
+    write.csv(tab2, dest, row.names=FALSE)
 }
 
 cleanPvals  <-  function(x) {
@@ -141,8 +133,8 @@ fitJags  <-  function(lmModelMatrix, data, run=TRUE, outputFileName='mod.bug') {
 ###############
 # PAPER NUMBERS
 ###############
-extractNumbersList  <-  function(metRates, outputPath) {
-    load(outputPath)
+extractNumbersList  <-  function(metRates, output) {
+
     # get average estimates and 95% CI for each parameter
     species           <-  unique(metRates$Species)
     averageEstimates  <-  ldply(species, getAverageEstimates, jagsSummary=output$model$BUGSoutput$summary, modelMatrix=output$lmerModel1)

### 2016-10-09 R/functions-figures.R

@@ -1,6 +1,9 @@
 ######################
 # AUXILLIARY FUNCTIONS
 ######################
+
+extrafont::loadfonts(quiet=TRUE)
+
 toPdf <- function(expr, filename, ...) {
     toDev(expr, pdf, filename, ...)
 }
@@ -282,8 +285,7 @@ fig1  <-  function() {
     text(0.5, -0.14/pinRatio, 'G / 4', adj=c(0, 0.5), col='grey60', cex=1.3, font=3)
 }
 
-fig2  <-  function(metRates, outputPath) {
-    load(outputPath)
+fig2  <-  function(metRates, output) {
     tmat  <-  output$model$BUGSoutput$sims.matrix
     tmat  <-  tmat[,paste0('beta[', seq_len(ncol(coef(output$lmerModel1)$Run)), ']')]
     colnames(tmat)  <-  names(coef(output$lmerModel1)$Run)

### 2016-10-09 wordFiles.R

@@ -1,21 +1,0 @@
-library(knitr)
-
-###########
-# MAIN TEXT 
-###########
-# to word
-system('pandoc -o text/MS.docx text/MS.md -s -S --bibliography references.bib --csl ecology.csl')
-
-##############
-# COVER LETTER 
-##############
-knit('text/coverLetter.Rmd', output=file.path(getwd(), 'text/coverLetter.md'), quiet=TRUE, encoding = 'utf-8')
-# to word
-system('pandoc -o text/coverLetter.docx text/coverLetter.md')
-
-####################
-# ANSWER TO REFEREES 
-####################
-knit('text/answers_to_reviewers.Rmd', output=file.path(getwd(), 'text/answers_to_reviewers.md'), quiet=TRUE, encoding = 'utf-8')
-# to word
-system('pandoc -o text/answers_to_reviewers.docx text/answers_to_reviewers.md -s -S --bibliography references.bib --csl ecology.csl')

### 2016-10-10 R/functions-analyses.R
changes to improve remake workflow

@@ -134,7 +134,6 @@ fitJags  <-  function(lmModelMatrix, data, run=TRUE, outputFileName='mod.bug') {
 # PAPER NUMBERS
 ###############
 extractNumbersList  <-  function(metRates, output) {
-
     # get average estimates and 95% CI for each parameter
     species           <-  unique(metRates$Species)
     averageEstimates  <-  ldply(species, getAverageEstimates, jagsSummary=output$model$BUGSoutput$summary, modelMatrix=output$lmerModel1)

### 2016-10-10 R/functions-figures.R

@@ -78,14 +78,14 @@ whiteGrid  <-  function (...) {
         col = 'grey70', lwd = 0.3, ...)
 }
 
-makeFigure1  <-  function() {
-    toPdf(fig1(), 'output/figures/fig1.pdf', width=8, height=5)
-    embed_fonts('output/figures/fig1.pdf')
+makeFigure1  <-  function(dest) {
+    toPdf(fig1(), dest, width=8, height=5)
+    embed_fonts(dest)
 }
 
-makeFigure2  <-  function(...) {
-    toPdf(fig2(...), 'output/figures/fig2.pdf', width=8, height=7)
-    embed_fonts('output/figures/fig2.pdf')
+makeFigure2  <-  function(dest, ...) {
+    toPdf(fig2(...), dest, width=8, height=7)
+    embed_fonts(dest)
 }
 
 ###############

### 2016-10-11 text/MS.Rmd
added github doi to text + few typos

@@
 stolonifera* (arborescent bryozoans that have a three dimensional tree-like growth form), *Hippopodina iririkiensis* (an encrusting bryozoan that grows mostly in two dimensions as flat disc), and an unidentified species of sponge belonging to the Microcionidae family (`r myCite('hart2009ecology')`; again a flat species that grows mostly in two dimensions) (Table 1).
-After 6 weeks in the field (average SST 22 ˚C), sheets bearing colonies were returned to the University of Queensland where they were kept in aerated seawater tanks (18 ˚C) for up to 48 h prior to measurements.
+After 6 weeks in the field (average SST 22˚C), sheets bearing colonies were returned to the University of Queensland where they were kept in aerated seawater tanks (18˚C) for up to 48 h prior to measurements.
 For encrusting colonies, sections of acetate bearing whole colonies were then cut from the sheets and fragmented (see e.g.
@@
 *3. Respiration measurements*  
-Individual metabolic rates (converted to $\mu$l $\textrm{O}_\textrm{2}$ h^-1^, *n* = `r nrow(numbersList$metRates)`), $B_i$, were measured as rate of oxygen consumption, $\dot{V}\textrm{O}_\textrm{2}$, of colony fragments at nominal temperatures of 10 or 25 ˚C.
+Individual metabolic rates (converted to $\mu$l $\textrm{O}_\textrm{2}$ h^-1^, *n* = `r nrow(numbersList$metRates)`), $B_i$, were measured as rate of oxygen consumption, $\dot{V}\textrm{O}_\textrm{2}$, of colony fragments at nominal temperatures of 10 or 25˚C.
 All individuals (regardless of size) and species were randomly assigned to test temperatures, and we used replicate vials and controlled temperature cabinets (see below).
@@
 
-where $m_b$ is the rate of change of $\textrm{O}_\textrm{2}$ saturation for blank vials containing no animals (\% h^-1^), ${\beta}_{\textrm{O}_\textrm{2}}$ is the oxygen capacitance of air-saturated sea water at 10 ˚C (6.4 ml l^-1^) or 25 ˚C (4.74 ml l^-1^) `r myCite('cameron1986book')` and $V$ is water volume (vials were 5 ml in volume and water volume was calculated by subtracting the volume of acetate and animals).
+where $m_b$ is the rate of change of $\textrm{O}_\textrm{2}$ saturation for blank vials containing no animals (\% h^-1^), ${\beta}_{\textrm{O}_\textrm{2}}$ is the oxygen capacitance of air-saturated sea water at 10˚C (6.4 ml l^-1^) or 25 ˚C (4.74 ml l^-1^) `r myCite('cameron1986book')` and $V$ is water volume (vials were 5 ml in volume and water volume was calculated by subtracting the volume of acetate and animals).
 At least four blank vials were recorded simultaneously with each 24-well plate to account for microbial oxygen consumption and sensor spots were calibrated with air-saturated seawater (100\% A.S.) and water containing 2\% sodium sulfite (0\% A.S.).
@@
 
-where mass, $\textrm{ln}M_i$ (mg), is a continuous variable, while temperature (two levels: 10 ˚C and 25 ˚C) and species (four levels corresponding to the four species) are categorical variables.
+where mass, $\textrm{ln}M_i$ (mg), is a continuous variable, while temperature (two levels: 10˚C and 25˚C) and species (four levels corresponding to the four species) are categorical variables.
 With this model, we allow each species to have a different mass-scaling slope that may change with temperature; we also account for species- and temperature-specific deviations on the metabolic normalization (i.e.
@@
 3 $\times$ 500,000/250 = 6,000).
-All data and code necessary to reproduce this paper, its analyses, tables and figures can be obtained on GitHub https://github.com/dbarneche/MTRBrEs.
+All data and code necessary to reproduce this paper, its analyses, tables and figures can be obtained on GitHub https://github.com/dbarneche/MTRBrEs `r myCite('barneche2016ecologyData')`.
 
@@
 2).
-For example, metabolic rates increase `r rounded(exp(averageEstimates$lnB25[averageEstimates$species == 'Stolonifera'])/exp(averageEstimates$lnB10[averageEstimates$species == 'Stolonifera']))`-fold between 10 ˚C and 25 ˚C for *B.
+For example, metabolic rates increase `r rounded(exp(averageEstimates$lnB25[averageEstimates$species == 'Stolonifera'])/exp(averageEstimates$lnB10[averageEstimates$species == 'Stolonifera']))`-fold between 10˚C and 25˚C for *B.
 stolonifera*, yielding a $Q_{10}$ of `r rounded(q10andErs$q10[q10andErs$species == 'Stolonifera'])`, which is equivalent to an activation energy of `r rounded(q10andErs$er[q10andErs$species == 'Stolonifera'], 2)` eV `r myCite(c('gillooly2001','yvondurocher2012a'))`.
-The best model indicates that the difference in metabolic rates between 25 ˚C and 10 ˚C for *Hippopodina iririkiensis* decreases, on average, `r rounded(tempDeltas[1,'tempDelta'])`-fold moving from `r round(exp(2))` to `r round(exp(5))` mg.
+The best model indicates that the difference in metabolic rates between 25˚C and 10˚C for *Hippopodina iririkiensis* decreases, on average, `r rounded(tempDeltas[1,'tempDelta'])`-fold moving from `r round(exp(2))` to `r round(exp(5))` mg.
 For *Bugula neritina*, however, rates increase `r rounded(1/tempDeltas[3,'tempDelta'])`-fold along the same body-mass gradient.
@@
 2) are inconsistent among the studied species and between growth forms (2D and 3D).
-For example, for organisms that exhibit 2D growth, *Hippopodina iririkiensis* shows exponents that encompass both 0.67 (Fractal geometry model prediction for 2D organisms; `r myCite('koontz2009amjbot')`) and 0.75 (Fractal geometry model prediction for 3D organisms; `r myCite(c('west1997science', 'west1999science'))`) at 10 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's10_97.5'], 2)`), but that are significantly shallower than 0.75 at 25 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's25_97.5'], 2)`).
-Microcionidae presents exponents that only overlap 0.67 at 10 ˚C, but not at 25 ˚C (95\% CI at 10 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's10_97.5'], 2)`; 95\% CI at 25 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's25_97.5'], 2)`).
-On the other hand, for 3D organisms, *Bugula neritina* presents scaling exponents that overlap both 0.67 and 0.75 at 25 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's25_97.5'], 2)`), but are significantly shallower at 10 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's10_97.5'], 2)`).
-Finally, *Bugula stolonifera* presents values that overlap both 0.67 and 0.75 at both temperatures (95\% CI at 10 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's10_97.5'], 2)`; 95\% CI at 25 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's25_97.5'], 2)`).
+For example, for organisms that exhibit 2D growth, *Hippopodina iririkiensis* shows exponents that encompass both 0.67 (Fractal geometry model prediction for 2D organisms; `r myCite('koontz2009amjbot')`) and 0.75 (Fractal geometry model prediction for 3D organisms; `r myCite(c('west1997science', 'west1999science'))`) at 10˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's10_97.5'], 2)`), but that are significantly shallower than 0.75 at 25˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's25_97.5'], 2)`).
+Microcionidae presents exponents that only overlap 0.67 at 10˚C, but not at 25˚C (95\% CI at 10˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's10_97.5'], 2)`; 95\% CI at 25˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's25_97.5'], 2)`).
+On the other hand, for 3D organisms, *Bugula neritina* presents scaling exponents that overlap both 0.67 and 0.75 at 25˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's25_97.5'], 2)`), but are significantly shallower at 10˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's10_97.5'], 2)`).
+Finally, *Bugula stolonifera* presents values that overlap both 0.67 and 0.75 at both temperatures (95\% CI at 10˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's10_97.5'], 2)`; 95\% CI at 25˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's25_97.5'], 2)`).
 
@@
 
-That metabolic rate scales at an exponent of less than 1 across all four species (with the exception of *Hippopodina iririkiensis* measured at 10 ˚C) has interesting implications for our understanding of the energetics of growth for different colonial animals.
+That metabolic rate scales at an exponent of less than 1 across all four species (with the exception of *Hippopodina iririkiensis* measured at 10˚C) has interesting implications for our understanding of the energetics of growth for different colonial animals.
 For instance, evidence suggest that water clearance and feeding rates scale either isometrically or allometrically with colony size in sponges `r myCite(c('riisgard1993meps', 'kowalke2000jembe', 'mcmurray2014aquatbiol'))`, and these rates may follow the same mass scaling of metabolic rates (see e.g.
@@
 Our findings suggest that, for some colonial species, size reductions mediated by predation or disturbance might modulate the energetic demands of individuals of similar body mass in different environments, thus affecting life-history trajectories.
-For instance, at 25 ˚C, a 400-gram individual of *Bugula stolonifera* will spend ~`r round(metabolicDiff$metabolicDiff[metabolicDiff$species == 'Stolonifera'])`-fold more energy than a similar-sized individual at 10 ˚C.
+For instance, at 25˚C, a 400-gram individual of *Bugula stolonifera* will spend ~`r round(metabolicDiff$metabolicDiff[metabolicDiff$species == 'Stolonifera'])`-fold more energy than a similar-sized individual at 10˚C.
 Conversely, due to the estimated positive interaction between mass-scaling and temperature, a similar-sized individual of *B.
-neritina* will spend ~`r round(metabolicDiff$metabolicDiff[metabolicDiff$species == 'Bugula'])`-fold more energy at 25 ˚C than at 10 ˚C.
+neritina* will spend ~`r round(metabolicDiff$metabolicDiff[metabolicDiff$species == 'Bugula'])`-fold more energy at 25˚C than at 10˚C.
 These differences, for instance, may help determine which species is a better competitor in different environments, as individual metabolic rates will also affect growth rates and time of reproduction onset `r myCite(c('sebens1987annurevecolsys', 'hart2009ecologyb'))`.
@@
 Our finding of allometric scaling of metabolic rate implies non-intuitive outcomes of such shifts in mean size of colonies in a population.
-For example, at 25 ˚C a population of 10 colonies of Microcionidae that each weigh 400 mg would have a cumulative metabolic demand of ~`r rounded(10 * exp(averageEstimates$lnB25[averageEstimates$species == 'Sponge'] + log(400)*averageEstimates$s25[averageEstimates$species == 'Sponge']), 0)` $\mu$l $\textrm{O}_\textrm{2}$ h^-1^.
+For example, at 25˚C a population of 10 colonies of Microcionidae that each weigh 400 mg would have a cumulative metabolic demand of ~`r rounded(10 * exp(averageEstimates$lnB25[averageEstimates$species == 'Sponge'] + log(400)*averageEstimates$s25[averageEstimates$species == 'Sponge']), 0)` $\mu$l $\textrm{O}_\textrm{2}$ h^-1^.
 If a disturbance event halved each colony with minimal loss of biomass, then a population of those (now) 20 colonies (i.e.
@@
 
-The dataset associated with this manuscript will be archived on DRYAD should the manuscript be accepted.
-All data and code necessary to reproduce this paper, its analyses, tables and figures can also be obtained on GitHub https://github.com/dbarneche/MTRBrEs.
+All data and code necessary to reproduce this paper, its analyses, tables and figures can be obtained on GitHub https://github.com/dbarneche/MTRBrEs (`r myCite('barneche2016ecologyData')`, doi: 10.5281/zenodo.159736).
 
@@
 *Figure 2.* Estimates from the three-way interaction model as obtained by Bayesian methods in *JAGS*.
-Results for each species are represented in groups of four plots, with grey and black colors representing actual data and parameter estimates at 10 ˚C and 25 ˚C, respectively.
+Results for each species are represented in groups of four plots, with grey and black colors representing actual data and parameter estimates at 10˚C and 25˚C, respectively.
 *Top-left* and *bottom-right*: posterior probabilities of mass-scaling exponents and metabolic normalizations (respectively), with bars over curves representing 95\% credible intervals, and numbers representing average estimates.
